### 2023-03-02 scripts/orv_sys_rev_data_analysis.Rmd

---
title: "Analysis of the included studies"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
    keep_tex: yes
    fig_crop: no
---

```{r echo=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('mdthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')
library('Hmisc')
library('ggwordcloud')

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'jpeg')
knitr::opts_chunk$set(out.height = "\\textheight",  out.width = "\\textwidth")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

#create a folder for storing the manuscript figures

if(!dir.exists('../figs_manuscript'))
    {
    dir.create('../figs_manuscript', 
               showWarnings = FALSE)
    }

#save figures in this folder
save_fig_here <- '../figs_manuscript'

save_thesis_fig_here <- "C:/Users/JAMESAZAM/Dropbox/My Academic Repository/_SACEMA/Academic/_PhD/__PhD_Thesis/_Thesis_LaTeX/figs/chapter_3"
```

```{r}
# Load the cleaned data 
long_data <- read_rds('../data/final_data/cleaned_data/review_data_long_cleaned.rds')

#The cleaned data without citations
compact_data_init <- read_rds('../data/final_data/cleaned_data/review_data_compact_cleaned.rds')

#a vector of shorter replacements for the original parametrization descriptions
parametrization_str_replacements <- c('values_from_literature' = 'literature', 
                  'values_from_guess_and_expert_opion' = 'expert_opinion', 
                  'values_fitted_to_ts' = 'fitted',
                  ' ' = ' and ' #replace space with 'and' to make it more readable
                  )

#a vector of shorter replacements for the original validation descriptions
validation_str_replacements <- c('compared_with_data' = 'data',
                                 'compared_with_other_model_output' = 'another_model',
                                 'model_not_validated' = 'none',
                                 ' ' = '_and_'
                                 )

# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds'))

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_combination_with_others', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

#replace the original descriptions of parametrization and validation with shorter strings
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  mutate(parametrization = str_replace_all(parametrization, 
                                           parametrization_str_replacements
                                           )
         ) %>% 
  mutate(validation = str_replace_all(validation, 
                                      validation_str_replacements
                                      )
         ) %>% 
  mutate(model_structure = str_replace(model_structure, 'deterministic stochastic', 'both'))

#Rearrange the columns
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  relocate(c(title, bibtexkey), .before = publication_year) %>% 
  relocate(publication_year, .before = bibtexkey) %>% 
  relocate(c(x0_reviewer, category, author:language), 
           .after = collab_type
           )

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd')

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% 
  filter(disease != 'fmd')

#Filter the FMD data
fmd_data <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```

# Human vaccine-preventable diseases

```{r eval=FALSE}
library(PRISMAstatement)

prisma_chart <- prisma(found = 12986,
       found_other = 0,
       no_dupes = 5012, 
       extra_dupes_box = TRUE,
       screened = 5012, 
       screen_exclusions = 4211, 
       full_text = 801,
       full_text_exclusions = 548, 
       qualitative = 253, 
       quantitative = 0,
       width = 800, height = 800,
       font_size = 12)

prisma_chart
```

<!-- ## Publications per year -->

<!-- ```{r} -->

<!-- publications_per_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(publication_year)  -->

<!-- publications_per_year_plot <- ggplot(data = publications_per_year) +  -->
<!--   geom_bar(aes(x = publication_year,  -->
<!--                y = n -->
<!--                ),  -->
<!--            stat = 'identity',  -->
<!--            color = 'black', -->
<!--            fill = 'tomato3' -->
<!--            ) + -->
<!--   scale_y_continuous(breaks = seq(0, 30, 2), -->
<!--                      labels = seq(0, 30, 2) -->
<!--                      ) + -->
<!--   labs(x = '**Year of publication**', y = '**Number of papers**') + -->
<!--   md_theme_minimal(base_size = 12) -->

<!-- plot(publications_per_year_plot) -->

<!-- ggsave(filename = 'publications_per_year.tiff',  -->
<!--        plot = publications_per_year_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'publications_per_year.tiff',  -->
<!--        plot = publications_per_year_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- ## Collaboration patterns in time -->

<!-- ### Number of studies per combinations of author affiliation types -->

<!-- ```{r echo=FALSE} -->
<!-- #' Count by collaboration types and make a bar plot -->
<!-- collab_types_count <- compact_data_with_citation_keys_no_fmd %>% -->
<!--     count(author_affiliation_type, sort = TRUE) %>%  -->
<!--     arrange(desc(n)) %>%  -->
<!--     adorn_totals() %>%  -->
<!--     adorn_percentages(denominator = 'col') %>%  -->
<!--     adorn_pct_formatting() %>%  -->
<!--     adorn_ns() -->

<!-- knitr::kable(collab_types_count,  -->
<!--              format = "latex", booktabs = TRUE,  -->
<!--              caption = 'Number of publications per author affiliation type combination') %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))  -->
<!-- ``` -->

<!-- ### Proportions of the total publications per year by collaboration type -->

<!-- ```{r echo=FALSE, fig.width=8} -->
<!-- #' Count the number of each author affiliation type per year (including FMD) -->
<!-- collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(year_aggreg,  -->
<!--         collab_type,  -->
<!--         name = 'total_publications' -->
<!--         ) -->

<!-- #The plot -->
<!-- # Modify the x-axis labels to account for the lumping of the years from 2005 and below -->
<!-- x_axis_labels <- as.character(seq(2005, 2020, 2)) -->

<!-- x_axis_labels[1] <- '1970-2005' -->

<!-- collab_proportions_plot <- ggplot(data = collap_types_by_year) +  -->
<!--   geom_bar(aes(x = year_aggreg, -->
<!--                y = total_publications, -->
<!--                fill = collab_type -->
<!--                ), -->
<!--            color = 'black', -->
<!--            stat = 'identity', -->
<!--            position = position_fill(reverse = TRUE)#, -->
<!--            #show.legend = FALSE -->
<!--              ) + -->
<!--   scale_fill_manual(values = c('tan3', 'turquoise2')) + -->
<!--   scale_x_continuous(breaks = seq(2005, 2020, 2), -->
<!--                      labels = x_axis_labels, -->
<!--                      expand = c(0, 0) -->
<!--                      ) + -->
<!--   #coord_flip() + -->
<!--   labs(#title = 'Publications per collaboration type over time', -->
<!--        x = '**Year**', -->
<!--        y = '**Proportion of total publications**', -->
<!--        fill = '**Collaboration type**'  -->
<!--          ) + -->
<!--   theme(legend.position = 'none') + -->
<!--  # md_theme_minimal(base_size = 16) #For Powerpoint -->
<!--   md_theme_minimal(base_size = 12) -->

<!-- plot(collab_proportions_plot) -->

<!-- ggsave(filename = 'collab_proportions.tiff',  -->
<!--        plot = collab_proportions_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'collab_proportions.tiff',  -->
<!--        plot = collab_proportions_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- ### Absolute number of publications per year by collaboration type -->

<!-- ```{r} -->
<!-- #' Count the number of each author affiliation type per year -->
<!-- collabs_per_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(year_aggreg,  -->
<!--         collab_type,  -->
<!--         name = 'total_publications' -->
<!--         )  -->

<!-- collabs_per_year_plot <- ggplot(data = collabs_per_year,  -->
<!--                                     aes(x = year_aggreg,  -->
<!--                                         y = total_publications -->
<!--                                         ),  -->
<!--                                     color = 'black') +  -->
<!--   geom_bar(aes(fill = collab_type),  -->
<!--            stat = 'identity',  -->
<!--            position = 'dodge' -->
<!--            ) +  -->
<!--   scale_fill_manual(values = c('tan3', 'turquoise2')) +  -->
<!--   scale_x_continuous(breaks = seq(2005, 2020, 2), -->
<!--                      labels = x_axis_labels, -->
<!--                      expand = c(0, 0) -->
<!--                      ) + -->
<!--   labs(x = '**Year**',  -->
<!--        y = '**Total publications**',  -->
<!--        fill = 'Collaboration type' -->
<!--        ) + -->
<!--   md_theme_minimal(base_size = 12) + -->
<!--   #md_theme_minimal(base_size = 16) + -->
<!--   theme(legend.position = 'bottom')  -->

<!-- plot(collabs_per_year_plot) -->

<!-- ggsave(filename = 'total_collabs_per_year_plot.tiff', -->
<!--        plot = collabs_per_year_plot,  -->
<!--        device = 'tiff', -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'total_collabs_per_year_plot.tiff', -->
<!--        plot = collabs_per_year_plot,  -->
<!--        device = 'tiff', -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- Even though non-academic collaborations increased over time, the share of total publications per year over time decreased. -->

<!-- ## Collaboration patterns in geographic space -->

## Collaboration types (aggregated)
```{r}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- compact_data_with_citation_keys_no_fmd %>%
    count(collab_type, sort = TRUE) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(studies_per_collab_type, caption = 'Number of studies per collaboration type (Human VPDs)',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Locations studied

### Real versus hypothetical

```{r}
location_types_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    mutate(location_type = ifelse(country_studied == 'none', 'none', 'real')) %>% 
    count(location_type) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(location_types_tally, 
             caption = 'Number of studies per location type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

### Larger than one country

```{r}
larger_geographies_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    filter(country_studied %in% c('west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
    count(country_studied) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(larger_geographies_tally, caption = 'Number of studies per location larger than one country', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

### Countries studied

```{r}
countries_studied_by_count <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>% 
    count(collab_type, country_studied, name = 'num_of_studies', sort = T)

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  group_by(country_studied) %>%
  summarise(num_of_studies = sum(num_of_studies),
            .groups = 'drop'
            ) %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  ungroup()

#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             caption = 'Number of studies per country', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'), font_size = 7)
```

### Continents studied

```{r}
continents_studied_by_count <- countries_studied_by_count %>% 
  filter(country_studied %nin% c('none', 'northern_hemisphere', 'global')) %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 'asia') == T ~ 'asia', 
                                         str_detect(continent_studied, 'africa') == T ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            .groups = 'drop'
            ) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

#Print the results
knitr::kable(continents_studied_by_count, 
             caption = 'Number of studies per continent', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Geographic connection of authors to the studied locations

```{r}
author_connectedness_top_5_countries <- compact_data_with_citation_keys_no_fmd %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
separate_rows(country_studied,  sep = ' ') %>% 
  group_by(collab_type) %>% 
  count(country_studied, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  arrange(collab_type, desc(yes)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no')) %>% 
  ungroup()

knitr::kable(author_connectedness_top_5_countries, caption = 'Studies with at least one author affiliation in the location studied', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position', 'repeat_header'), font_size = 8)
```

### Author affiliation in the studied location (overall)

```{r}
author_connectedness_to_location_aggreg <- compact_data_with_citation_keys_no_fmd %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('other', 'none', 'northern_hemisphere', 'west_africa', 'global')) %>% 
  count(collab_type, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(author_connectedness_to_location_aggreg, caption = 'At least one author with an affiliation in the studied location', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Diseases studied per collab type -->

<!-- ```{r echo=FALSE} -->
<!-- #count the number of unique diseases studied -->
<!-- disease_studied_by_collab_type <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   separate_rows(disease,  sep = ' ') %>%  -->
<!--   distinct(title, disease, .keep_all = TRUE) %>%  -->
<!--   count(collab_type, disease, sort = TRUE) %>% -->
<!--   pivot_wider(names_from = collab_type, values_from = n) %>%  -->
<!--   replace_na(list(purely_academic = 0, mixed = 0)) %>%  -->
<!--   adorn_totals(where = c('row', 'col')) %>%  -->
<!--   adorn_percentages(... = c('purely_academic', 'mixed')) %>%  -->
<!--   adorn_pct_formatting(... = c('purely_academic', 'mixed')) %>%  -->
<!--   adorn_ns(... = c('purely_academic', 'mixed')) %>%  -->
<!--   mutate(disease = str_to_sentence(str_replace(disease, '_', ' '))) %>%  -->
<!--   arrange(desc(Total)) -->

<!-- knitr::kable(disease_studied_by_collab_type, caption = 'Number of studies per disease and collaboration type', format = "latex", booktabs = TRUE) %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position')) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #tally the diseases by collaboration type -->
<!-- disease_by_collab_type_long <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   separate_rows(disease,  sep = ' ') %>%  -->
<!--   count(collab_type, disease, sort = TRUE)  -->

<!-- disease_by_collab_type_plot <- ggplot(disease_by_collab_type_long) +  -->
<!--   geom_bar(aes(x = reorder(disease, n),  -->
<!--                y = n,  -->
<!--                fill = collab_type -->
<!--                ),  -->
<!--            color = 'black', -->
<!--            stat = 'identity', -->
<!--            position = position_fill(reverse = TRUE) -->
<!--            ) +  -->
<!--   labs(x = '**Disease**', y = '**Number of studies**') +  -->
<!--   coord_flip() + -->
<!--   md_theme_minimal(base_size = 12) -->

<!-- print(disease_by_collab_type_plot) -->
<!-- ``` -->

## Interventions

### Types of interventions

```{r}
intervention_categories <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, intervention_type, sort = TRUE) %>% 
  pivot_wider(names_from = intervention_type, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no_vax', 'vax_combination_with_others', 'vax_single')) %>% 
  adorn_pct_formatting(... = c('no_vax', 'vax_combination_with_others', 'vax_single')) %>% 
  adorn_ns(... = c('no_vax', 'vax_combination_with_others', 'vax_single'))

knitr::kable(intervention_categories, caption = 'Number of studies per intervention categories', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Impact of vaccination

```{r}
vax_impact_conclusions <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_single') %>% 
  count(collab_type, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
    replace_na(list(yes = 0, n0 = 0, the_outcomes_were_inconclusive = 0)) %>% 
  rename('inconclusive' = 'the_outcomes_were_inconclusive') %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(vax_impact_conclusions, 
             caption = 'Conclusions about impact of vaccination', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position')
                                  )
```

```{r eval=FALSE}
no_vax_modelled <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(disease,  sep = ' ') %>% 
    filter(vax_modelled == 'no') %>% 
    count(disease, sort = TRUE, name = 'num_of_studies') %>% 
     adorn_totals(where = 'row') %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(no_vax_modelled, caption = 'Number of studies per disease for which vaccination was not modelled', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
vax_impact_by_disease <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_single') %>% 
  count(disease, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  rename('inconclusive' = 'the_outcomes_were_inconclusive') %>% 
    replace_na(list(yes = 0, no = 0, inconclusive = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(vax_impact_by_disease, 
             caption = 'Conclusions about impact of vaccination per disease', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
no_vax_impact_by_disease_citation_table <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_single', is_vax_effective %in% c('no', 'the_outcomes_were_inconclusive')) %>% 
  select(title, disease, bibtexkey, publication_year, intervention_modelled, is_vax_effective)

#Print the results
# knitr::kable(no_vax_impact_by_disease_citation_table, 
#              caption = 'Studies that did not find vaccination as most impactful'
#              )

write.xlsx(no_vax_impact_by_disease_citation_table, file = '../citation_tables/no_vax_impact_by_disease_citation_table.xlsx')
```

```{r eval=FALSE}
vax_as_single_intervention_studies <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_single') %>% 
  select(title, disease, bibtexkey, publication_year, intervention_modelled)

#Print the results
# knitr::kable(no_vax_impact_by_disease_citation_table, 
#              caption = 'Studies that did not find vaccination as most impactful'
#              )

write.xlsx(vax_as_single_intervention_studies, file = '../citation_tables/vax_as_single_intervention_studies.xlsx')
```

## Modelling objectives 
<!-- We grouped studies into whether they assessed the impact of a future intervention strategy (prospective), a past intervention, or both (dual).  -->

<!-- In general, there were more (73.2%; 167/228) studies about prospective intervention impact assessments than retrospective assessments (24.6%; 56/228), and the dual (2.2%; 5/228). A similar pattern was observed across the two collaboration types except that purely academic collaborations were more likely to perform the dual assessments.   -->

<!-- In terms of the outbreak types, there was an approximately equal number of papers about real and hypothetical outbreaks overall, and there was no clear difference between the two collaborations.   -->

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_count <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('both', 'future', 'past')) %>% 
  adorn_pct_formatting(... = c('both', 'future', 'past')) %>% 
  adorn_ns(... = c('both', 'future', 'past'))

#Print a table of the results
knitr::kable(modelling_objectives_count,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Outbreak types

```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_pct_formatting(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_ns(... = c('hypothetical_outbreak', 'real_outbreak'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type,
             caption = 'outbreak types by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
objectives_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  count(objectives, collab_type, name = 'num_of_studies', sort = TRUE) %>% 
  pivot_wider(names_from = objectives, 
              values_from = num_of_studies
              ) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('future', 'past', 'both')) %>% 
  adorn_pct_formatting(... = c('future', 'past', 'both')) %>% 
  adorn_ns(... = c('future', 'past', 'both'))

knitr::kable(objectives_by_collab_type, 
             caption = 'Number of studies per objective by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Model characteristics  -->

## Individual heterogeneity: agent-based versus compartmental models

```{r}
individual_representation <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, individual_representation, sort = TRUE) %>% 
  pivot_wider(names_from = individual_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('compartments', 'agents', 'individuals_representation_other')) %>% 
  adorn_pct_formatting(... = c('compartments', 'agents', 'individuals_representation_other')) %>% 
  adorn_ns(... = c('compartments', 'agents', 'individuals_representation_other'))

knitr::kable(individual_representation, caption = 'How individuals were represented', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```
<!-- ### Individual representation by model structure -->
<!-- ```{r} -->
<!-- mod_structure_vs_individual_representation <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   group_by(collab_type) %>%  -->
<!--   count(model_structure, individual_representation, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = model_structure, values_from = n) %>%  -->
<!--   adorn_totals(where = c('col', 'row')) -->

<!-- knitr::kable(mod_structure_vs_individual_representation, caption = 'Model structure versus individual heterogeneity') -->
<!-- ``` -->

## Spatial heterogeneity

```{r}
space_representation <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, space_representation, sort = TRUE) %>% 
  pivot_wider(names_from = space_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(space_representation, caption = 'Was space explicitly represented in the model',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ### Spatial models by model structure -->
<!-- ```{r} -->
<!-- mod_structure_vs_space_representation <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   group_by(collab_type) %>%  -->
<!--   count(model_structure, space_representation, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = model_structure, values_from = n) %>%  -->
<!--   adorn_totals(where = c('col', 'row')) -->

<!-- knitr::kable(mod_structure_vs_space_representation, caption = 'Model structure versus spatial heterogeneity') -->
<!-- ``` -->

## Model dynamics: deterministic vs stochastic

```{r}
stochastic_vs_deterministic <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, model_structure) %>% 
  pivot_wider(names_from = model_structure, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('both', 'deterministic', 'stochastic')) %>% 
  adorn_pct_formatting(... = c('both', 'deterministic', 'stochastic')) %>% 
  adorn_ns(... = c('both', 'deterministic', 'stochastic'))

knitr::kable(stochastic_vs_deterministic, caption = 'Model dynamics (deterministic versus stochastic)', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Modelling methods -->

## Outcomes measured

<!-- We tallied the overall top five outcomes most used between the two collaboration types.  -->

<!-- The top five most commonly used outcomes, irrespective of collaboration type, were: final epidemic size, attack rate, timing of epidemic peak, cases averted, and outbreak duration. -->

<!-- There was no difference in the two collaboration types as they were commonly used by both groups. -->

```{r}
outcomes_overall <- compact_data_with_citation_keys_no_fmd %>% 
mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(collab_type, outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  filter(outcome_measured %nin% c('outcome_other', '', 's')) %>% 
  count(collab_type, outcome_measured, name = 'num_of_studies') %>% 
  ungroup(outcome_measured) %>% 
  arrange(desc(num_of_studies), .by_group = TRUE) %>% 
  ungroup() %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(outcomes_overall, caption = 'Model outcomes by collaboration type', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position', 'repeat_header'), font_size = 8)
```

```{r eval=FALSE}
outcomes_tally_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  ungroup(outcome_measured) %>% 
  filter(outcome_measured %nin% c('outcome_other', '')) %>% 
  group_by(collab_type) %>% 
  count(outcome_measured) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  # slice(1:6) %>% 
  ungroup()

# outcomes_tally_plot <- ggplot(data = outcomes_tally_by_collab_type, 
#                               aes(x = reorder(outcome_measured, n), 
#                                   y = n
#                                   )
#                               ) + 
#   geom_bar(stat = 'identity') + 
#   labs(x = '**Outcome**', y = '**Number of times used**') +
#   coord_flip() +
#   facet_wrap('collab_type') +
#   md_theme_minimal()

collab_type_labels <- c('purely_academic' = 'Purely academic', 
                        'mixed' = 'Mixed'
                        )

outcomes_tally_plot <- ggplot(data = outcomes_tally_by_collab_type, 
                              aes(label = outcome_measured, 
                                  size = n
                                  )
                              ) + 
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 10) + 
  facet_wrap(~ collab_type, 
             labeller = as_labeller(collab_type_labels),
             strip.position = 'top'
             ) +
  #theme_minimal() + 
  theme(strip.text = element_text(size = 15, face = 'bold'))

print(outcomes_tally_plot)

ggsave(filename = 'outcomes_tally_plot.tiff',
       plot = outcomes_tally_plot, 
       device = 'tiff',
       path = save_fig_here
       )

ggsave(filename = 'outcomes_tally_plot.tiff',
       plot = outcomes_tally_plot, 
       device = 'tiff',
       path = save_thesis_fig_here
       )
```

<!-- ### Parametrization methods -->

<!-- ```{r} -->
<!-- #wide data for table -->
<!-- model_parametrization_wide <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(collab_type, parametrization, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = parametrization, values_from = n) %>%  -->
<!--   adorn_totals(where = c('row', 'col')) %>%  -->
<!--   adorn_percentages(... = -c('Total', 'collab_type')) %>%  -->
<!--   adorn_pct_formatting(... = -c('Total', 'collab_type')) %>%  -->
<!--   adorn_ns(... = -c('Total', 'collab_type'))  -->

<!-- #table -->
<!-- knitr::kable(model_parametrization_wide, caption = 'How model parameters are obtained',  -->
<!--              format = "latex", booktabs = TRUE) %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'scale_down', 'HOLD_position')) -->
<!-- ``` -->

<!-- ```{r eval=FALSE} -->
<!-- #long data for plotting -->
<!-- model_parametrization_long <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   group_by(collab_type) %>%  -->
<!--   count(parametrization, sort = TRUE) -->

<!-- #plotting -->
<!-- ggplot(data = model_parametrization_long,  -->
<!--        aes(fill = collab_type,  -->
<!--            area = n,  -->
<!--            label = parametrization, -->
<!--            subgroup = parametrization -->
<!--            ) -->
<!--        ) +  -->
<!--   geom_treemap() + -->
<!--   geom_treemap_text(colour = "white", -->
<!--                     place = "centre", -->
<!--                     size = 15, -->
<!--                     grow = TRUE -->
<!--                     ) + -->
<!--   md_theme_minimal(base_size = 12)  -->

<!-- ``` -->

<!-- ### Validation methods  -->

<!-- ```{r} -->
<!-- #wide data for table -->
<!-- model_validation_wide <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(collab_type, validation, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = validation, values_from = n) %>%  -->
<!--   replace_na(list(none = 0,  -->
<!--                   data = 0,  -->
<!--                   another_model = 0,  -->
<!--                   none_and_data = 0,  -->
<!--                   data_and_another_model = 0 -->
<!--                   ) -->
<!--              ) %>%  -->
<!--   adorn_totals(where = c('row', 'col')) %>%  -->
<!--   adorn_percentages(... = -c('Total', 'collab_type')) %>%  -->
<!--   adorn_pct_formatting(... = -c('Total', 'collab_type')) %>%  -->
<!--   adorn_ns(... = -c('Total', 'collab_type'))  -->

<!-- #table -->
<!-- knitr::kable(model_validation_wide, caption = 'How the model\'s performance is assessed',  -->
<!--              format = "latex", booktabs = TRUE) %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'scale_down', 'HOLD_position')) -->
<!-- ``` -->

## Sensitivity analysis

<!-- Less than half (46.5%; 106/228) of the total studies performed sensitivity analyses. There was no clear difference between the collaborations.  -->

```{r eval=FALSE}
mod_structure_vs_sensitivity_analysis <- compact_data_with_citation_keys_no_fmd %>% 
  count(model_structure, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(mod_structure_vs_sensitivity_analysis, 'Sensitivity analysis in relation to model structure', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

```{r}
sensitivity_analysis <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(sensitivity_analysis, caption = 'Sensitivity analysis', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Data use and data availability

```{r}
data_use_and_availability <- compact_data_with_citation_keys_no_fmd %>% 
  filter(data_used == 'yes') %>%
  count(collab_type, data_available, sort = TRUE) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(data_use_and_availability, 
             caption = 'Data accessibility', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Code availability

```{r}
code_availability <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, simulation_code_available) %>% 
  pivot_wider(names_from = simulation_code_available, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(code_availability, caption = 'Code availability', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
papers_with_code_available <- compact_data_with_citation_keys_no_fmd %>% 
    filter(simulation_code_available == 'yes') %>% 
    as.data.frame()

openxlsx::write.xlsx(papers_with_code_available, file = '../citation_tables/papers_with_code_available.xlsx')

```

# Foot and mouth disease (FMD)

<!-- ## Collaboration patterns in time -->

## Unique combinations of author affiliation types

```{r}
#' Count by collaboration types and make a bar plot
author_affiliation_type_combos <- fmd_data %>%
    count(author_affiliation_type, sort = TRUE) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type)) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(author_affiliation_type_combos, caption = 'Number of studies by author affiliation type combination',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Collaboration types (aggregated)

```{r}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- fmd_data %>%
    count(collab_type, sort = TRUE) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(studies_per_collab_type, caption = 'Number of studies per collaboration type',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Proportions of the total publications per year by collaboration type

```{r echo=FALSE, fig.width=8}
#' Count the number of each author affiliation type per year (including FMD)
fmd_collab_types_per_year <- fmd_data %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'

fmd_collab_proportions_plot <- ggplot(data = fmd_collab_types_per_year) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)#,
           #show.legend = FALSE
             ) +
  scale_fill_manual(values = c('tan3', 'turquoise2')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  #coord_flip() +
  labs(#title = 'Publications per collaboration type over time',
       x = '**Year**',
       y = '**Proportion of total publications**',
       fill = '**Collaboration type**' 
         ) +
  theme(legend.position = 'none') +
 # md_theme_minimal(base_size = 16) #For Powerpoint
  md_theme_minimal(base_size = 12)

plot(fmd_collab_proportions_plot)

```

### Absolute number of publications per year by collaboration type

```{r}
#' Count the number of each author affiliation type per year
fmd_collabs_per_year <- fmd_data %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

collabs_per_year_plot <- ggplot(data = fmd_collabs_per_year, 
                                    aes(x = year_aggreg, 
                                        y = total_publications
                                        ), 
                                    color = 'black') + 
  geom_bar(aes(fill = collab_type), 
           stat = 'identity', 
           position = 'dodge',
           width = 0.5
           ) + 
  scale_fill_manual(values = c('tan3', 'turquoise2')) + 
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  labs(x = '**Year**', 
       y = '**Total publications**', 
       fill = 'Collaboration type'
       ) +
  md_theme_minimal(base_size = 12) +
  #md_theme_minimal(base_size = 16) +
  theme(legend.position = 'bottom')

plot(collabs_per_year_plot)

```

<!-- ## Collaboration patterns in geographic space -->

## Geographic connection of authors to the studied locations 
```{r}
author_connectedness_to_location_fmd <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
  group_by(collab_type) %>% 
  count(country_studied, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  arrange(collab_type, desc(yes)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no')) %>% 
  ungroup()

knitr::kable(author_connectedness_to_location_fmd, caption = 'Studies with at least one author affiliation in the location studied', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Aggregated

```{r}
author_connectedness_to_location_aggreg <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('other', 'none', 'northern_hemisphere', 'west_africa', 'global')) %>% 
  count(collab_type, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(author_connectedness_to_location_aggreg, 
             caption = 'Studies with at least one author affiliation in the location studied', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Interventions

### Types of interventions

```{r}
intervention_categories_fmd <- fmd_data %>% 
  count(collab_type, intervention_type, sort = TRUE) %>% 
  pivot_wider(names_from = intervention_type, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no_vax', 'vax_combination_with_others', 'vax_single')) %>% 
  adorn_pct_formatting(... = c('no_vax', 'vax_combination_with_others', 'vax_single')) %>% 
  adorn_ns(... = c('no_vax', 'vax_combination_with_others', 'vax_single'))

knitr::kable(intervention_categories_fmd, caption = 'Number of studies per intervention type', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Impact of vaccination

```{r}
fmd_vax_impact_conclusions <- fmd_data %>% 
  filter(intervention_type == 'vax_single') %>% 
  count(collab_type, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  replace_na(list(no = 0, inconclusive = 0, yes = 0)) %>% 
  rename('inconclusive' = 'the_outcomes_were_inconclusive') %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(fmd_vax_impact_conclusions, 
             caption = 'Conclusions about impact of vaccination', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Modelling objectives

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_fmd <- fmd_data %>% 
  count(collab_type, objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('future', 'past')) %>% 
  adorn_pct_formatting(... = c('future', 'past')) %>% 
  adorn_ns(... = c('future', 'past'))

#Print a table of the results
knitr::kable(modelling_objectives_fmd,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ### Modelling objectives and outbreak type by collaboration types -->

<!-- ```{r echo=FALSE} -->
<!-- objectives_and_outbreak_types_fmd <- fmd_data %>%  -->
<!--   count(objectives, outbreak_type, name = 'num_of_studies', sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = objectives,  -->
<!--               values_from = num_of_studies -->
<!--               ) %>%  -->
<!--   replace_na(list(future = 0, past = 0, both = 0)) %>%  -->
<!--   adorn_totals(where = c('row', 'col')) %>%  -->
<!--   adorn_percentages(... = c('future', 'past')) %>%  -->
<!--   adorn_pct_formatting(... = c('future', 'past')) %>%  -->
<!--   adorn_ns(... = c('future', 'past'))  -->

<!-- knitr::kable(objectives_and_outbreak_types_fmd,  -->
<!--              caption = 'Number of studies per objective and outbreak type',  -->
<!--              format = "latex", booktabs = TRUE) %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position')) -->
<!-- ``` -->

## Outbreak types

```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type_fmd <- fmd_data %>% 
  count(collab_type, outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_pct_formatting(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_ns(... = c('hypothetical_outbreak', 'real_outbreak'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type_fmd,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Model characteristics  -->

## Individual heterogeneity: agent-based versus compartmental models

```{r}
individual_representation_fmd <- fmd_data %>% 
  count(collab_type, individual_representation, sort = TRUE) %>% 
  pivot_wider(names_from = individual_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('compartments', 'agents')) %>% 
  adorn_pct_formatting(... = c('compartments', 'agents')) %>% 
  adorn_ns(... = c('compartments', 'agents'))

knitr::kable(individual_representation_fmd, 
             caption = 'How individuals are represented', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Spatial heterogeneity

```{r}
space_representation_fmd <- fmd_data %>% 
  count(collab_type, space_representation, sort = TRUE) %>% 
  pivot_wider(names_from = space_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(space_representation_fmd, caption = 'Models with spatial structure', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Model dynamics: deterministic vs stochastic

```{r}
stochastic_vs_deterministic_fmd <- fmd_data %>% 
  count(collab_type, model_structure) %>% 
  pivot_wider(names_from = model_structure, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('deterministic', 'stochastic')) %>% 
  adorn_pct_formatting(... = c('deterministic', 'stochastic')) %>% 
  adorn_ns(... = c('deterministic', 'stochastic'))

knitr::kable(stochastic_vs_deterministic_fmd, 
             caption = 'Model dynamics (deterministic versus stochastic)', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

<!-- ## Modelling methods -->

## Outcomes measured
```{r}
outcomes_tally_by_collab_type <- fmd_data %>% 
  mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(collab_type, outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  filter(outcome_measured %nin% c('outcome_other', '')) %>% 
  count(collab_type, outcome_measured, name = 'num_of_studies') %>% 
  ungroup(outcome_measured) %>% 
  arrange(desc(num_of_studies), .by_group = TRUE) %>% 
  ungroup() %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(outcomes_tally_by_collab_type, 
             caption = 'Model outcomes by collaboration type', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position', 'repeat_header'))
```

```{r eval=FALSE}
collab_type_labels <- c('purely_academic' = 'Purely academic', 
                        'mixed' = 'Mixed'
                        )

outcomes_tally_plot <- ggplot(data = outcomes_tally_by_collab_type, 
                              aes(label = outcome_measured, 
                                  size = n
                                  )
                              ) + 
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 7) + 
  facet_wrap(~ collab_type, 
             labeller = as_labeller(collab_type_labels),
             strip.position = 'top'
             ) +
  #theme_minimal() + 
  theme(strip.text = element_text(size = 15, face = 'bold'))

#save the plot
ggsave(filename = 'outcomes_tally_fmd_plot.tiff',
       plot = outcomes_tally_plot, 
       device = 'tiff',
       path = save_fig_here
       )
```

## Parametrization methods

```{r}
#wide data for table
model_parametrization_fmd_wide <- fmd_data %>% 
  count(collab_type, parametrization, sort = TRUE) %>% 
  pivot_wider(names_from = parametrization, values_from = n) %>% 
  replace_na(list('expert_opinion and fitted' = 0, 'literature and fitted' = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type')) 
#table
knitr::kable(model_parametrization_fmd_wide, 
             caption = 'Model parametrization methods', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'scale_down',  'HOLD_position'))
```

## Validation methods

```{r}
#wide data for table
model_validation_fmd_wide <- fmd_data %>% 
  count(collab_type, validation, sort = TRUE) %>% 
  pivot_wider(names_from = validation, values_from = n) %>% 
  replace_na(list(none = 0, 
                  data = 0, 
                  another_model = 0, 
                  data_and_another_model = 0
                  )
             ) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type'))

#table
knitr::kable(model_validation_fmd_wide, 
             caption = 'Model validation methods', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped',  'HOLD_position'))
```

## Sensitivity analysis

```{r}
sensitivity_analysis_fmd <- fmd_data %>% 
  count(collab_type, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(sensitivity_analysis_fmd, caption = 'Sensitivity analysis', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped',  'HOLD_position'))

```

## Data use and data availability

```{r}
data_use_and_availability_fmd <- fmd_data %>% 
  filter(data_used == 'yes') %>%
  count(collab_type, data_available, sort = TRUE) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(data_use_and_availability_fmd, 
             caption = 'Data accessibility', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Code availability

```{r}

code_availability_fmd <- fmd_data %>% 
  count(collab_type, simulation_code_available) %>% 
  pivot_wider(names_from = simulation_code_available, values_from = n) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(code_availability_fmd, caption = 'Code availability', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```