### 2023-03-02 scripts/orv_sys_rev_data_analysis_old.Rmd

---
title: "Systematic review of outbreak response models"
author: "James Azam"
date: "01/06/2021"
output:
  word_document:
    keep_md: yes
#  html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Sets the working directory for subsequent chunks, but not this chunk
#knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
```

```{r echo=FALSE, message=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```

```{r echo=FALSE, include=FALSE, eval=FALSE}
# Source the script that cleans the raw data
# Load the script that cleans the data and saves it; 
# this should be run each time, just in case something has changed in the data cleaning process;
# The script clears the environment after cleaning the data
#source('./data_cleaning/orv_models_review_data_cleaning.R')
```

```{r echo=FALSE, include=FALSE}
# Load the cleaned data 
long_data <- read_rds('../data/final_data/cleaned_data/review_data_long_cleaned.rds')

#The cleaned data without citations
compact_data_init <- read_rds('../data/final_data/cleaned_data/review_data_compact_cleaned.rds')

compact_data <- compact_data_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  relocate(vax_modelled, intervention_type, .after = 'intervention_modelled')

# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds'))

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
  mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  relocate(vax_modelled, intervention_type, .after = 'intervention_modelled')

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd')

compact_data_no_fmd <- compact_data %>% filter(disease != 'fmd')

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% filter(disease != 'fmd') 
```

# Exploratory Data Analysis
1. Publications per year
2. Distribution of collaborations across types (aggregated)
3. Countries/settings being studied (author affiliation vs country studied)

# What science is being inferred from models?

1. Impact of vaccination (human diseases vs FMD); analyse with and without Influenza
2. What diseases are being studied (to understand vaccine impact)
3. Which interventions are being studied?

```{r echo=FALSE}

pubs_per_year <- compact_data_no_fmd %>% 
  mutate(publication_year = as.numeric(publication_year)) %>% 
  count(publication_year)

pubs_per_year_bar_plot <- ggplot(data = pubs_per_year) + 
  geom_bar(aes(x = publication_year, 
               y = n
               ), 
           stat = 'identity', 
           color = 'black',
           fill = 'tomato3') +
  scale_y_continuous(breaks = seq(0, 30, 2),
                     labels = seq(0, 30, 2)
                     ) +
  labs(x = 'Year of publication', y = 'Number of papers') +
  theme_minimal()

plot(pubs_per_year_bar_plot)

```

# Modelling objectives

```{r echo=FALSE}
#Find the uniques entries
modelling_objectives_unique_entries <- long_data_no_fmd %>% 
  select(title, objectives) %>% 
  distinct()

#count the number of studies per objective type
modelling_objectives_count <- modelling_objectives_unique_entries %>% 
  count(objectives, name = 'num_of_studies')

#Print a table of the results
knitr::kable(modelling_objectives_count,
             col.names = c('Objective type',
                           'Number of studies'
                           ),
             caption = 'Number of studies per objective type'
             )
```

```{r echo=FALSE}
#Find the unique entries 
outbreak_types_unique_entries <- long_data_no_fmd %>% 
  select(title, outbreak_type) %>% 
  distinct()

#Count the number of studies per outbreak type  
outbreak_types_count <- outbreak_types_unique_entries %>% 
  count(outbreak_type, name = 'num_of_studies')

#Print a table of the results
knitr::kable(outbreak_types_count,
             col.names = c('Outbreak type',
                           'Number of studies'
                           ),
             caption = 'Number of studies per outbreak type'
             )
```

```{r echo=FALSE}
#Find the number of unique entries per objective type and outbreak type combination  
modelling_objectives_by_outbreak_type_unique_entries <- long_data_no_fmd %>% 
  select(title, objectives, outbreak_type) %>% 
  group_by(title) %>% 
  distinct() %>% 
  ungroup()

modelling_objectives_by_outbreak_type_count <- modelling_objectives_by_outbreak_type_unique_entries %>% 
  count(objectives, outbreak_type, name = 'num_of_studies') %>% 
  arrange(desc(num_of_studies)) %>% 
  pivot_wider(names_from = outbreak_type, 
              values_from = num_of_studies
              )

knitr::kable(modelling_objectives_by_outbreak_type_count, 
             # col.names = c('Objective type', 
             #               'Outbreak type', 
             #               'Number of studies'
             #               ),
             caption = 'Number of studies per objective and outbreak type'
             ) 
```

```{r echo=FALSE} 
#Year with the highest publications
max_pub_year <- compact_data_no_fmd %>% 
    count(publication_year, sort = T) %>% 
    filter(n == max(n))

print(max_pub_year)
```

## Number of studies per disease 
```{r echo=FALSE}
disease_studied_count <- long_data_no_fmd %>% 
  group_by(disease) %>% 
  distinct(title) %>% 
  count(disease, name = 'num_of_studies', sort = T) %>%
  ungroup()

knitr::kable(disease_studied_count, col.names = c('Disease', 'Number of studies'))
```

### Disease most studied
```{r echo=FALSE}
disease_studied_most <- disease_studied_count %>% 
  filter(num_of_studies == max(num_of_studies))

knitr::kable(disease_studied_most)
```

## Number of studies per country and continent

```{r echo=FALSE, warning=FALSE, message=FALSE}
#FIXME: I have to cross tabulate the countries studied with number of studies with at least one author located in the country being studied.

#Tally of how many times and percentage of times each COUNTRY has been studied
countries_studied_by_count <- long_data_no_fmd %>% 
    group_by(country_studied) %>%
    distinct(title) %>% 
    count(country_studied, name = 'num_of_studies', sort = T) %>% 
    ungroup(country_studied) %>% 
    mutate(country_studied = as_factor(country_studied), 
           percentage = round(num_of_studies/sum(num_of_studies)*100, 1)
           )

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  mutate(country_studied_aggreg = if_else(num_of_studies < 9, 
                                          'other', 
                                          as.character(country_studied)
                                          )
         ) %>% 
  group_by(country_studied_aggreg) %>%
  summarise(num_of_studies = sum(num_of_studies),
            percentage = sum(percentage),
            .groups = 'drop'
            ) %>% 
  arrange(desc(percentage)) %>% 
  mutate(country_studied_aggreg = as_factor(country_studied_aggreg))

#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             col.names = c('Country', 
                           'Number of studies', 
                           'Percentage'
                           ),
             caption = 'Number/percentage of studies per country'
             )
```

## Number and percentage of studies per continent
```{r echo=FALSE}
continents_studied_by_count <- countries_studied_by_count %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 'asia') == T ~ 'asia', 
                                         str_detect(continent_studied, 'africa') == T ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            percentage = sum(percentage),
            .groups = 'drop'
            ) %>% 
  arrange(desc(percentage))

#Print the results
knitr::kable(continents_studied_by_count, 
             col.names = c('Continent', 
                           'Number of studies', 
                           'Percentage'
                           ),
             caption = 'Number/percentage of studies per continent'
             )
```

### Country most studied 
```{r echo=FALSE}
countries_studied_most <- countries_studied_by_count %>% 
    filter(country_studied != 'none') %>%
    filter(num_of_studies == max(num_of_studies))

#Print the result as a table    
knitr::kable(countries_studied_most, 
             col.names = c('Country', 
                           'Number of studies', 
                           'Percentage'
                           ),
             caption = 'Most studied country'
             )
```

### Continent most studied
```{r echo=FALSE}

continent_studied_most <- continents_studied_by_count %>% 
    dplyr::filter(continent_studied != 'none') %>% 
    filter(num_of_studies == max(num_of_studies))

#Print the result as a table      
knitr::kable(continent_studied_most,
             col.names = c('Continent', 
                           'Number of studies',
                           'Percentage'
                           ),
             caption = 'Most studied continent'
             )
```

## Number of studies per continent
```{r echo=FALSE}
continents_studied <- continents_studied_by_count %>% 
  filter(continent_studied != 'none') %>% 
  filter(continent_studied != 'global') %>% 
  filter(continent_studied != 'northern_hemisphere') %>% 
  group_by(continent_studied) %>% 
  summarise(num_of_studies = sum(num_of_studies )) %>% 
  ungroup() %>% 
  mutate(percentage = round(num_of_studies/sum(num_of_studies)*100, 1))

#Print table of results
knitr::kable(continents_studied_by_count %>% arrange(desc(percentage)),
             col.names = c('Continent',
                           'Number of papers',
                           'Percentage'
                           ),
             caption = 'Number/percent of studies by continent'
             )

```

### Highest number of studies per year and disease

```{r echo=FALSE}
knitr::kable(tibble(max_pub_year = max_pub_year$publication_year, 
                    disease_with_max_pubs = disease_studied_most$disease,
                    country_most_studied = countries_studied_most$country_studied
                    )
             )
```

# Primary objectives

## 1.A. Is modelling being used to inform outbreak response policy-making?

### I. How are the results distributed according to collaboration type?

Note that we do not include FMD in all the primary analyses. We only use it for comparisons, where necessary.

```{r echo=FALSE}
# 1B. Create the unique combinations of collaboration types and reword the string NA
collab_type_data <- compact_data_no_fmd %>%
  select(title, publication_year, author_affiliation_type, author_in_country_studied) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    )

#' Count by collaboration types and make a bar plot
collab_types_count <- collab_type_data %>%
    count(author_affiliation_type) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type))

knitr::kable(collab_types_count %>% arrange(desc(n)))
```

### II. What has been the trend in collaboration over the years?

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}

collap_types_count_barplot <- collab_types_count %>%
    ggplot() + geom_bar(aes(x = reorder(author_affiliation_type, n),
                            y = n
                            ),
                        stat = 'identity',
                        color = 'black'
                        ) +
    scale_y_continuous(breaks = seq(0, 140, 20),
                       labels = seq(0, 140, 20)
                       ) +
    labs(title = 'Number of studies per collaboration type', 
         x = '', 
         y = 'Number of studies') + 
  coord_flip() +
  theme_minimal() +
  theme(legend.position = 'none')

plot(collap_types_count_barplot)

knitr::kable(collab_types_count, 
             col.names = c('Collaboration type', 'Number of publications'),
             caption = 'Number of publications per collaboration type'
             )

```

### III. Proportions of the total publications per year shared by collaboration type (purely academic versus mixed).

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'

collab_type_by_year_plot <- ggplot(data = collap_types_by_year) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)  
             ) +
  scale_fill_manual(values = c('tan1', 'turquoise1')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels
                     ) +
  labs(title = 'Publications per collaboration type over time',
       x = 'Year',
       y = 'Proportion of total publications in the year',
       fill = 'Collaboration type' 
         ) +
  theme_minimal(base_size = 12)

plot(collab_type_by_year_plot)

```
### IV. Number of studies by collaboration type (purely academic versus mixed).

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        ) %>% 
  pivot_wider(names_from = collab_type, values_from = total_publications)

knitr::kable(collap_types_by_year %>% clean_names() %>% adorn_totals(... = c(mixed, purely_academic)))

```

```{r echo=FALSE}
#' Count the number of each author affiliation type per year
collap_types_by_year <- compact_data_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

collap_types_by_year_plot <- ggplot(data = collap_types_by_year, 
                                    aes(x = year_aggreg, 
                                        y = total_publications
                                        ), 
                                    color = 'black') + 
  geom_bar(aes(fill = collab_type), 
           stat = 'identity', 
           position = 'dodge'
           ) +
  labs(x = 'Year', y = 'Total publications')

plot(collap_types_by_year_plot)

```

Even though non-academic collaborations increased over time, the share of total publications per year over time decreased.

### V. Academic collaborations versus other non-academic collaborations (excluding purely academic collaborations).

```{r echo=FALSE}

academic_collabs_versus_others <- compact_data_no_fmd %>% 
  filter(author_affiliation_type != 'academic_institutions') %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(str_detect(author_affiliation_type, 'academic_institutions'), 
                                         'Academic', 
                                         'Other'
                                         )
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

#plot the result
academic_collabs_trend <- ggplot(data = academic_collabs_versus_others) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)  
             ) +
  scale_fill_manual(values = c('tan1', 'turquoise1')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels
                     ) +
  labs(title = 'Academic and other collaborations over time',
       x = 'Year',
       y = 'Proportion of publications',
       fill = 'Collaboration type' 
         ) +
  theme_minimal(base_size = 12)

plot(academic_collabs_trend)

#Number of each collaboration type

```

# B. How often are papers authored by locals?
```{r}
author_location_count_by_disease_wide <- compact_data_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  mutate(author_in_country_studied = as_factor(author_in_country_studied), 
         disease = as_factor(disease)
         ) %>% 
  group_by(disease) %>%
  distinct(title, .keep_all = TRUE) %>% 
  count(author_in_country_studied) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0, not_applicable = 0)) %>% 
  mutate(total = not_applicable + no + yes) %>% 
  arrange(desc(total))

knitr::kable(author_location_count_by_disease_wide)
```

```{r}
author_location_count_by_disease_long <- author_location_count_by_disease_wide %>% 
  select(-c(not_applicable, total)) %>% 
  pivot_longer(cols = c(yes, no), 
               names_to = 'author_in_country_studied',
               values_to = 'n'
               )

author_location_count_by_disease_plot <- ggplot(data = author_location_count_by_disease_long, 
       aes(x = reorder(disease, n), 
           y = n
           )
       ) + geom_bar(aes(fill = author_in_country_studied),
                    stat = 'identity',
                    position = position_fill(reverse = TRUE)
                    ) +
  labs(x = 'Disease', y = 'Proportion of total studies') +
  coord_flip()

plot(author_location_count_by_disease_plot)
```

# 2. Is vaccination always the most impactful single intervention?

## Classification of interventions by inclusion/exclusion of vaccination

### Human diseases
```{r echo=FALSE}
intervention_categories_human_diseases <- compact_data_no_fmd %>% 
  group_by(intervention_type, vax_modelled_with_non_vax) %>% 
  count(is_vax_effective, sort = TRUE) %>% 
  pivot_wider(names_from = vax_modelled_with_non_vax, values_from = n) %>% 
  replace_na(list(no = 0, yes = 0))

knitr::kable(intervention_categories_human_diseases)
```
### FMD

```{r echo=FALSE}
intervention_categories_fmd <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  group_by(intervention_type, vax_modelled_with_non_vax) %>% 
  count(is_vax_effective, sort = TRUE) %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  adorn_totals(where = 'col')

knitr::kable(intervention_categories_fmd)
```

## Classification of interventions by inclusion of vaccination 
```{r echo=FALSE}
study_categories_by_vax_single <- intervention_categories_human_diseases %>% 
  filter(intervention_type == 'vax_single')

knitr::kable(study_categories_by_vax_single)
```

## Citation table of studies that modelled vaccination as a single intervention
```{r echo=FALSE}
intervention_categories_citation_tbl <- compact_data_with_citation_keys_no_fmd %>%
  filter(intervention_type == 'vax_single') %>% 
  select(title, publication_year, bibtexkey, disease, intervention_modelled, vax_modelled, vax_modelled_with_non_vax, is_vax_effective)

knitr::kable(intervention_categories_citation_tbl %>% arrange(is_vax_effective
, disease))

```

# Is vaccination the most effective when compared to other single interventions?

## Vaccine impact categories (human diseases)
```{r echo=FALSE}
vax_impact_categories_no_fmd <- compact_data_no_fmd %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        )

#Print the results
knitr::kable(vax_impact_categories_no_fmd %>% adorn_totals, 
             col.names = c('Vaccination is most impactful', 
                           'Number of studies'
                           ),
             caption = 'Studies\' conclusions about impact of vaccination (Human diseases)'
             )
```

## Vaccine impact categories (FMD)
```{r echo=FALSE}
vax_impact_categories_fmd <- compact_data %>%
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        )

#Print the results
knitr::kable(vax_impact_categories_fmd %>% adorn_totals, 
             col.names = c('Vaccination is most impactful', 
                           'Number of studies'
                           ),
             caption = 'Studies\' conclusions about impact of vaccination (FMD)'
             )
```

## Citation table for human diseases
```{r echo=FALSE}
vax_impact_categories_citations_no_fmd <- compact_data_with_citation_keys %>% 
  filter(disease != 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  select(title, publication_year, disease, objectives, is_vax_effective, bibtexkey) %>% 
  arrange(is_vax_effective)

#Print the results
knitr::kable(vax_impact_categories_citations_no_fmd,
             caption = 'Studies\' conclusions about impact of vaccination (citations for human diseases)'
             )
```

## Citation table for FMD
```{r echo=FALSE}
vax_impact_categories_citations_fmd <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  select(title, publication_year, disease, objectives, is_vax_effective, bibtexkey) %>% 
  arrange(is_vax_effective)

#Print the results
knitr::kable(vax_impact_categories_citations_fmd,
             caption = 'Studies\' conclusions about impact of vaccination (citations for human diseases)'
             )

```

```{r echo=FALSE, eval=FALSE}
vax_impact_plot <- ggplot(data = vax_impact_categories) + 
  geom_bar(
    aes(x = is_vax_effective, 
        y = n,
        fill = is_vax_effective
        ),
    stat = 'identity',
    width = 0.25
    ) +
  scale_y_continuous(breaks = seq(0, 
                                  max(vax_impact_categories$n),
                                  2
                                  ),
                     labels = seq(0, 
                                max(vax_impact_categories$n),
                                2
                                )
                     ) +
  scale_x_discrete(breaks = c('no', 
                              'yes', 
                              'the_outcomes_were_inconclusive'
                              ), 
                   labels = c('no' = 'No', 
                              'yes' = 'Yes', 
                              'the_outcomes_were_inconclusive' = 'Inconclusive'
                              )
                   ) +
  scale_fill_discrete(name = 'Vaccine impact',
                      labels = c('No', 
                                 'Inconclusive',
                                 'Yes'
                      )
                      ) +
  labs(x = 'Vaccination is most impactful single intervention',
       y = 'Number of studies'
       ) +
  theme_minimal(base_size = 12)

plot(vax_impact_plot)

```

# Number of studies per disease

## By collaboration type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_collab_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         author_affiliation_type = as_factor(str_replace(author_affiliation_type, ' ', '+'))
         ) %>% 
    mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, collab_type, sort = TRUE) %>%
  pivot_wider(names_from = collab_type, values_from = n) %>% 
  replace_na(list(purely_academic = 0, mixed = 0)) %>% 
  mutate(total = purely_academic + mixed) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_collab_type)
```

## By outbreak type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_outbreak_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         outbreak_type = as_factor(str_replace(outbreak_type, ' ', '+'))
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, outbreak_type, sort = TRUE) %>%
  pivot_wider(names_from = outbreak_type, values_from = n) %>% 
  replace_na(list(hypothetical_outbreak = 0, real_outbreak = 0)) %>% 
  mutate(total = hypothetical_outbreak + real_outbreak) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_outbreak_type)
```

## By objective type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_objective_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         objectives = as_factor(objectives)
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, objectives) %>%
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  mutate(total = future + past + both) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_objective_type)
```

## FMD by collaboration type
```{r echo=FALSE}
fmd_studies_by_collab_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  mutate(disease = as_factor(disease), 
         author_affiliation_type = as_factor(str_replace(author_affiliation_type, ' ', '+'))
         ) %>% 
  mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         ) %>% 
  count(collab_type, sort = TRUE)

knitr::kable(fmd_studies_by_collab_type)
```
## FMD by objective type
```{r echo=FALSE}
fmd_studies_by_objective_and_outbreak_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  count(objectives, outbreak_type) %>% 
  pivot_wider(names_from = outbreak_type, values_from = n)

knitr::kable(fmd_studies_by_objective_and_outbreak_type)
```

# Number of studies per disease and intervention.

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup()

#interventions are disease specific
interventions_per_disease <- interventions_unique_entries %>% 
  mutate(intervention_modelled = str_trim(intervention_modelled)) %>% 
  filter(intervention_modelled != 'e.g.') %>% #remove false rows
  filter(intervention_modelled != 'various') %>% #remove false rows
  group_by(disease) %>% 
  count(intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  ungroup()

knitr::kable(interventions_per_disease)

```

# Number of studies per outcome

```{r echo=FALSE}
#count the number of unique outcomes measured
outcomes_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(outcome_measured, .keep_all = TRUE) %>% 
  ungroup()

#Count the number of studies 
outcomes_count <- outcomes_unique_entries %>% 
  group_by(outcome_measured) %>% 
  count(outcome_measured, name = 'num_of_studies', sort = T) %>% 
  ungroup() %>% 
  filter(outcome_measured != 's')

knitr::kable(outcomes_count)

#Plot the result
outcome_count_plot <- ggplot(data = outcomes_count) + 
  geom_point(aes(x = reorder(outcome_measured, 
                             num_of_studies), 
               y = num_of_studies
               ), 
           stat = 'identity',
           color = 'darkslategray4'
           ) +
  coord_flip() +
  theme_minimal()

plot(outcome_count_plot)

```

# Number of studies per model structure type

```{r echo=FALSE}
#Keep only the unique entries 
model_structure_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(model_structure) %>% 
  ungroup()

#Count the number of studies
model_structure_unique_count <- model_structure_unique_entries %>%
    group_by(model_structure) %>% 
    count(model_structure, name = 'num_of_studies')

#Print the table
knitr::kable(model_structure_unique_count)
```

# Number of studies by the model structure, parametrization, and validation types

```{r echo=FALSE}
#Cross tally the model structure with parametrization and validation categories
#FIXME: Check that studies are either not being under counted or double-counted. What do I want to count here? Is it the unique studies per method or combined method because most studies used a combination of parametrization methods especially
parametrization_and_validation_by_model_structure <- compact_data_no_fmd %>%
  mutate(parametrization = str_replace(parametrization, ' ', ' + '),
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(model_structure, parametrization, validation) %>% 
  pivot_wider(names_from = model_structure, values_from = n)

#Print the table of results
knitr::kable(parametrization_and_validation_by_model_structure,
             caption = 'Parametrization and validation types per model structure')

```

# Model parametrization & validation

```{r echo=FALSE}
#Keep only the unique entries
parametrization_and_validation_count <- compact_data_no_fmd %>% 
  mutate(parametrization = str_replace(parametrization, ' ', ' + '), 
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(parametrization, 
        validation, 
        name = 'num_of_studies'
        ) %>% 
  group_by(parametrization) %>% 
  arrange(desc(num_of_studies), .by_group = T)

#Print the table
knitr::kable(parametrization_and_validation_count)

```

# How is sensitivity analysis distributed by model structure?

```{r echo=FALSE}
#Keep only the unique entries
sensitivity_analysis_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(sensitivity_analysis, .keep_all = TRUE) %>% 
  ungroup()

#Count the number of studies
sensitivity_analysis_count <- sensitivity_analysis_unique_entries %>% 
  count(model_structure, sensitivity_analysis, name = 'num_of_studies')

#Print the table
knitr::kable(sensitivity_analysis_count)

```

# Data use and availability

```{r echo=FALSE}
#Keep only the unique entries
code_and_data_availability_count <- compact_data_no_fmd %>% 
  count(data_used, data_available) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(not_applicable = 0, 
                  yes = 0, 
                  no = 0
                  )
             )

#Print the results
knitr::kable(code_and_data_availability_count, caption = 'Data use and availability')

```
# Code availability

```{r echo=FALSE}
#Keep only the unique entries
code_availability_count <- compact_data_no_fmd %>% 
  count(simulation_code_available)

#Print the results
knitr::kable(code_availability_count, caption = 'Code availability')

```

# Secondary summaries
## Studies before 2006 (low frequency period) 
```{r echo=FALSE}
#Number of studies before 2005 (low frequency period)
n_studies_before_2006 <- compact_data_no_fmd %>% 
  filter(publication_year < 2006) %>% 
  summarise(n = n())

print(paste('Number of studies before 2006:', n_studies_before_2006))
```
## Number of studies between 2006-2019 (high frequency period)
```{r echo=FALSE}
n_studies_btwn_2006_2019 <- compact_data_no_fmd %>% 
  filter(publication_year >= 2006, publication_year <= 2019) %>% 
  summarise(n = n())

print(paste('Number of studies between 2006 and 2019:', n_studies_btwn_2006_2019))
```

# Exploration of the peak periods (2009-2011, 2015-2016, 2019)

## Citations of studies during 2009-2011
```{r echo=FALSE}
#Citation table of studies during 2009-2011 (Peak period)
studies_in_2009_2011_df <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  filter(publication_year >= 2009, publication_year <= 2011) %>%
  arrange(disease) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_in_2009_2011_df, caption = 'Studies published during 2009-2011 (Peak period)')
```
### Number of studies during 2009-2011 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_2009_2011_df <- studies_in_2009_2011_df %>%
  count(disease, objectives) %>% 
  arrange(disease, n) %>% 
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(both = 0, future = 0, past = 0))

knitr::kable(n_studies_2009_2011_df, caption = 'Studies published during 2009-2011 (Peak period)')
```

## Citation table of studies during 2015-2016 per disease and objective type (Peak period)
```{r echo=FALSE}
studies_between_2015_2016_df <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  filter(publication_year >= 2015, publication_year <= 2016) %>%
  arrange(disease, publication_year) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_between_2015_2016_df, caption = 'Studies published in 2015-2016 (Peak period)')
```

### Number of studies from 2015-2016 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_between_2015_2016_df <- studies_between_2015_2016_df %>% 
  count(disease, objectives) %>% 
  arrange(desc(n))  %>% 
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(both = 0, future = 0, past = 0))

knitr::kable(n_studies_between_2015_2016_df, caption = 'Papers published in 2015-2016 (Peak period)')
```

## Citation table of studies in 2019 (Peak period)
```{r echo=FALSE}
studies_in_2019_df <- compact_data_with_citation_keys_no_fmd %>%
  filter(publication_year == 2019) %>%
  arrange(disease) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_in_2019_df, caption = 'Studies published in 2019 (Peak year)')
```
### Number of studies in 2019 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_in_2019_df <- studies_in_2019_df %>% 
  count(disease, objectives) %>% 
  arrange(n)

knitr::kable(n_studies_in_2019_df, caption = 'Number of studies published in 2019 per disease and objective type (Peak year)')
```

### Total number of studies in 2019 (Peak period)
```{r echo=FALSE}
n_studies_in_2019 <- n_studies_in_2019_df %>%  
  summarise(n = n())

print(paste('Number of studies in 2019 (Peak year):', 
            n_studies_in_2019
            )
      )
```

## Number of diseases in the peak periods
```{r echo=FALSE}
peak_period_df <- bind_rows(studies_in_2009_2011_df, 
                            studies_between_2015_2016_df, 
                            studies_in_2019_df
                            ) %>% 
  arrange(disease, publication_year)

knitr::kable(peak_period_df , 
             caption = 'Studies published in peak period'
             )
```

```{r echo=FALSE}
n_studies_by_disease_and_objectives_peak_periods <- peak_period_df %>% 
  count(disease, objectives)

knitr::kable(n_studies_by_disease_and_objectives_peak_periods, 
             caption = 'Number of studies per disease and objective in peak periods'
             )
```

```{r echo=FALSE}
total_studies_by_disease_peak_period_df <- peak_period_df %>% 
  count(disease) %>% 
  arrange(n)

knitr::kable(total_studies_by_disease_peak_period_df, 
             caption = 'Number of studies per disease in peak periods'
             )
```

## Influenza during 2009 - 2011 peaks
```{r echo=FALSE}
n_Influenza_studies_by_objectives_2009_2011 <- peak_period_df %>% 
  filter(publication_year >= 2009, publication_year <= 2011) %>%
  filter(disease == 'influenza') %>% 
  count(objectives)

knitr::kable(n_Influenza_studies_by_objectives_2009_2011, 
             caption = 'Number of Influenza studies per objective type (2009 & 2011)'
             )
```
## Ebola in 2015-2016
```{r echo=FALSE}
n_ebola_studies_by_objectives_2015_2016 <- studies_between_2015_2016_df %>% 
  filter(disease == 'ebola') %>% 
  count(objectives)

knitr::kable(n_ebola_studies_by_objectives_2015_2016, 
             caption = 'Number of Ebola studies per objective type (2015-2016)'
             )
```
### Studies that used both retrospective and prospective assessments in peak periods
```{r echo=FALSE}
knitr::kable(peak_period_df %>% filter(objectives == 'both'), 
             caption = 'Studies with retrospective and prospective assessments'
             )

```