### 2023-05-15 analyses/scripts/archive/orv_sys_rev_data_analysis_FMD.Rmd
Moved redundant scripts into archive directory

---
title: "Outbreak response models of foot and mouth disease (1970-2019) - A primary analysis"
author: "James Azam"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    keep_md: yes
    reference_docx: word-style.docx
    toc: yes
---

```{r echo=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('mdthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')
library('Hmisc')
library('treemapify')
library('ggwordcloud')

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'jpeg')
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
knitr::opts_chunk$set(message = FALSE)

#create a folder for storing the manuscript figures

if(!dir.exists('../figs_manuscript'))
    {
    dir.create('../figs_manuscript', 
               showWarnings = FALSE)
    }

#save figures in this folder
save_fig_here <- '../figs_manuscript'
```

```{r}
# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds'))

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

#a vector of shorter replacements for the original parametrization descriptions
parametrization_str_replacements <- c('values_from_literature' = 'literature', 
                  'values_from_guess_and_expert_opion' = 'expert_opinion', 
                  'values_fitted_to_ts' = 'fitted',
                  ' ' = ' and ' #replace space with 'and' to make it more readable
                  )

#a vector of shorter replacements for the original validation descriptions
validation_str_replacements <- c('compared_with_data' = 'data',
                                 'compared_with_other_model_output' = 'another_model',
                                 'model_not_validated' = 'none',
                                 ' ' = '_and_'
                                 )

#replace the original descriptions of parametrization and validation with shorter strings
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  mutate(parametrization = str_replace_all(parametrization, 
                                           parametrization_str_replacements
                                           )
         ) %>% 
  mutate(validation = str_replace_all(validation, 
                                      validation_str_replacements
                                      )
         ) %>% 
  mutate(model_structure = str_replace(model_structure, 'deterministic stochastic', 'both'))

#Rearrange the columns
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  relocate(c(title, bibtexkey), .before = publication_year) %>% 
  relocate(publication_year, .before = bibtexkey) %>% 
  relocate(c(x0_reviewer, category, author:language), 
           .after = collab_type
           )

#Filter the FMD data
fmd_data <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```

# Primary objectives of review

## Collaboration patterns in time

### Number of FMD studies per collaboration types

```{r echo=FALSE}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- fmd_data %>%
    count(author_affiliation_type, sort = TRUE) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type)) %>% 
  arrange(desc(n))

knitr::kable(studies_per_collab_type)
```

### Proportions of the total publications per year by collaboration type

```{r echo=FALSE, fig.width=8}
#' Count the number of each author affiliation type per year (including FMD)
fmd_collab_types_per_year <- fmd_data %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'

fmd_collab_proportions_plot <- ggplot(data = fmd_collab_types_per_year) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)#,
           #show.legend = FALSE
             ) +
  scale_fill_manual(values = c('tan3', 'turquoise2')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  #coord_flip() +
  labs(#title = 'Publications per collaboration type over time',
       x = '**Year**',
       y = '**Proportion of total publications**',
       fill = '**Collaboration type**' 
         ) +
  theme(legend.position = 'none') +
 # md_theme_minimal(base_size = 16) #For Powerpoint
  md_theme_minimal(base_size = 12)

plot(fmd_collab_proportions_plot)

```

### Absolute number of publications per year by collaboration type

```{r}
#' Count the number of each author affiliation type per year
fmd_collabs_per_year <- fmd_data %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

collabs_per_year_plot <- ggplot(data = fmd_collabs_per_year, 
                                    aes(x = year_aggreg, 
                                        y = total_publications
                                        ), 
                                    color = 'black') + 
  geom_bar(aes(fill = collab_type), 
           stat = 'identity', 
           position = 'dodge',
           width = 0.5
           ) + 
  scale_fill_manual(values = c('tan3', 'turquoise2')) + 
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  labs(x = '**Year**', 
       y = '**Total publications**', 
       fill = 'Collaboration type'
       ) +
  md_theme_minimal(base_size = 12) +
  #md_theme_minimal(base_size = 16) +
  theme(legend.position = 'bottom')

plot(collabs_per_year_plot)

```

## Collaboration patterns in geographic space

### Geographic connection of authors to the studied locations

#### Top 5 countries
```{r}
author_connectedness_to_location_fmd <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
  group_by(collab_type) %>% 
  count(country_studied, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  slice(n = 1:5) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no')) %>% 
  ungroup()

knitr::kable(author_connectedness_to_location_fmd)
```
#### Aggregated

```{r}
author_connectedness_to_location_aggreg <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(country_studied %nin% c('other', 'none', 'northern_hemisphere', 'west_africa', 'global')) %>% 
  count(collab_type, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(author_connectedness_to_location_aggreg)
```

## Interventions

### Types of interventions

```{r}
intervention_categories_fmd <- fmd_data %>% 
  count(collab_type, intervention_type, sort = TRUE) %>% 
  pivot_wider(names_from = intervention_type, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no_vax', 'vax_alongside', 'vax_single')) %>% 
  adorn_pct_formatting(... = c('no_vax', 'vax_alongside', 'vax_single')) %>% 
  adorn_ns(... = c('no_vax', 'vax_alongside', 'vax_single'))

knitr::kable(intervention_categories_fmd)
```

### Impact of vaccination

```{r}
fmd_vax_impact_conclusions <- fmd_data %>% 
  filter(intervention_type == 'vax_single') %>% 
  count(collab_type, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  replace_na(list(no = 0, inconclusive = 0, yes = 0)) %>% 
  rename('inconclusive' = 'the_outcomes_were_inconclusive') %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(fmd_vax_impact_conclusions, 
             caption = 'Conclusions about impact of vaccination'
             )
```

# Secondary objectives of review

## Modelling objectives

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_fmd <- fmd_data %>% 
  count(collab_type, objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('future', 'past')) %>% 
  adorn_pct_formatting(... = c('future', 'past')) %>% 
  adorn_ns(... = c('future', 'past'))

#Print a table of the results
knitr::kable(modelling_objectives_fmd,
             caption = 'Study objectives by collaboration type'
             )
```

## Outbreak types

```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type_fmd <- fmd_data %>% 
  count(collab_type, outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_pct_formatting(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_ns(... = c('hypothetical_outbreak', 'real_outbreak'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type_fmd,
             caption = 'Study objectives by collaboration type'
             )
```

## Modelling objectives and outbreak type by collaboration types

```{r echo=FALSE}
objectives_and_outbreak_types_fmd <- fmd_data %>% 
  count(objectives, outbreak_type, name = 'num_of_studies', sort = TRUE) %>% 
  pivot_wider(names_from = objectives, 
              values_from = num_of_studies
              ) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('future', 'past')) %>% 
  adorn_pct_formatting(... = c('future', 'past')) %>% 
  adorn_ns(... = c('future', 'past'))

knitr::kable(objectives_and_outbreak_types_fmd, 
             caption = 'Number of studies per objective and outbreak type'
             ) 
```

## Model characteristics

### Individual heterogeneity: agent-based versus compartmental models

```{r}
individual_representation_fmd <- fmd_data %>% 
  count(collab_type, individual_representation, sort = TRUE) %>% 
  pivot_wider(names_from = individual_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('compartments', 'agents')) %>% 
  adorn_pct_formatting(... = c('compartments', 'agents')) %>% 
  adorn_ns(... = c('compartments', 'agents'))

knitr::kable(individual_representation_fmd)
```

### Spatial heterogeneity

```{r}
space_representation_fmd <- fmd_data %>% 
  count(collab_type, space_representation, sort = TRUE) %>% 
  pivot_wider(names_from = space_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(space_representation_fmd, caption = 'Spatial models')
```

### Model dynamics: deterministic vs stochastic

```{r}
stochastic_vs_deterministic_fmd <- fmd_data %>% 
  count(collab_type, model_structure) %>% 
  pivot_wider(names_from = model_structure, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('deterministic', 'stochastic')) %>% 
  adorn_pct_formatting(... = c('deterministic', 'stochastic')) %>% 
  adorn_ns(... = c('deterministic', 'stochastic'))

knitr::kable(stochastic_vs_deterministic_fmd, 
             caption = 'Model dynamics (deterministic versus stochastic)'
             )
```

## Modelling methods
### Outcomes measured
```{r}
outcomes_tally_by_collab_type <- fmd_data %>% 
  mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  ungroup(outcome_measured) %>% 
  filter(outcome_measured %nin% c('outcome_other', '')) %>% 
  group_by(collab_type) %>% 
  count(outcome_measured) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  slice(1:6) %>% 
  ungroup()

collab_type_labels <- c('purely_academic' = 'Purely academic', 
                        'mixed' = 'Mixed'
                        )

outcomes_tally_plot <- ggplot(data = outcomes_tally_by_collab_type, 
                              aes(label = outcome_measured, 
                                  size = n
                                  )
                              ) + 
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 7) + 
  facet_wrap(~ collab_type, 
             labeller = as_labeller(collab_type_labels),
             strip.position = 'top'
             ) +
  #theme_minimal() + 
  theme(strip.text = element_text(size = 15, face = 'bold'))

#save the plot
ggsave(filename = 'outcomes_tally_fmd_plot.png',
       plot = outcomes_tally_plot, 
       device = 'png',
       path = save_fig_here
       )

print(outcomes_tally_plot)
```

### Parametrization methods

```{r}
#wide data for table
model_parametrization_fmd_wide <- fmd_data %>% 
  count(collab_type, parametrization, sort = TRUE) %>% 
  pivot_wider(names_from = parametrization, values_from = n) %>% 
  replace_na(list('expert_opinion and fitted' = 0, 'literature and fitted' = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type')) 
#table
knitr::kable(model_parametrization_fmd_wide, 
             caption = 'How model parameters are obtained'
             )
```

### Validation methods

```{r}
#wide data for table
model_validation_fmd_wide <- fmd_data %>% 
  count(collab_type, validation, sort = TRUE) %>% 
  pivot_wider(names_from = validation, values_from = n) %>% 
  replace_na(list(none = 0, 
                  data = 0, 
                  another_model = 0, 
                  data_and_another_model = 0
                  )
             ) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type'))

#table
knitr::kable(model_validation_fmd_wide, 
             caption = 'How the model\'s performance is assessed'
             )
```

### Sensitivity analysis

```{r}
sensitivity_analysis_fmd <- fmd_data %>% 
  count(collab_type, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(sensitivity_analysis_fmd)

```

### Data use and data availability

```{r}
data_use_and_availability_fmd <- fmd_data %>% 
  filter(data_used == 'yes') %>%
  count(collab_type, data_available, sort = TRUE) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(data_use_and_availability_fmd, 
             caption = 'How often is the data used available for public access')

```

### Code availability

```{r}

code_availability_fmd <- fmd_data %>% 
  count(collab_type, simulation_code_available) %>% 
  pivot_wider(names_from = simulation_code_available, values_from = n) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(code_availability_fmd)
```