### 2023-06-26 analyses/scripts/S2_File.Rmd
Removed unused packages

---
title: "Outbreak response models of vaccine-preventable diseases in humans (1970-2019) - A systematic review"
subtitle: "Supplementary file (analysis for main manuscript)"
author: "James M. Azam, Xiaoxi Pang, Elisha B. Are, Juliet R.C. Pulliam, Matthew J. Ferrari"
output:
  pdf_document:
    toc: yes
    fig_crop: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      dev = 'png', 
                      out.height = "\\textheight",  
                      out.width = "\\textwidth", 
                      message = FALSE, 
                      warning = FALSE
                      )
```

```{r}
#' load packages
suppressPackageStartupMessages({
  library('tidyverse')
  library('ggthemes')
  library('scales')
  library('janitor')
  library('countrycode')
  library('bib2df')
  library('conflicted')
  library('here')
  })

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```

```{r}
# if the figs folder doesn't exist, create it for storing the figures
gh_figs_dir <- "../figs/"
dropbox_figs_dir <- "~/Dropbox/My Academic Repository/Manuscripts/First author/In review/2022 - Systematic review of ORV models/Epidemics/submission2/manuscript_writeup/figs/"

if (!dir.exists(here::here(gh_figs_dir))) {
  dir.create(here::here(gh_figs_dir),
    showWarnings = TRUE
  )
}
```

```{r}
args <- c(
  here::here("data", "review_data_long_cleaned.rds"),
  here::here("data", "review_data_compact_cleaned.rds"),
  here::here("data", "compact_data_with_citation_keys_cleaned.rds")
  )

# Load the cleaned data 
long_data <- readRDS(args[1])

#The cleaned data without citations
compact_data_init <- readRDS(args[2])

# The data with citations 
compact_data_with_citation_keys <- readRDS(args[3])

# Keep the human disease studies (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd')

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% 
  filter(disease != 'fmd')

# Filter the FMD data
fmd_data <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```
# General analyses

## Number of studies included

The review involved `r nrow(compact_data_with_citation_keys)` published between
`r min(compact_data_with_citation_keys$publication_year)` and `r max(compact_data_with_citation_keys$publication_year)`.

## Breakdown of studies before and after 2005
```{r}
before_2005 <- compact_data_with_citation_keys_no_fmd %>% 
  select(publication_year) %>% 
  mutate(before_2005 = ifelse(publication_year <= 2005, "yes", "no")) %>% 
  count(before_2005, name = "num_of_studies") %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front")

knitr::kable(before_2005)
```

## Publications per year
```{r}
publications_per_year <- compact_data_with_citation_keys_no_fmd %>%
  count(publication_year)

publications_per_year_plot <- ggplot(data = publications_per_year) +
  geom_bar(aes(x = publication_year,
               y = n
               ),
           stat = 'identity',
           color = 'black',
           fill = 'tomato3'
           ) +
  scale_y_continuous(breaks = seq(0, 30, 2),
                     labels = seq(0, 30, 2),
                     expand = c(0, 0)
                     ) +
    scale_x_continuous(breaks = seq(1970, 2020, 5),
                     labels = seq(1970, 2020, 5)
                     ) +
  labs(x = "Year of publication", y = "Number of papers") +
  theme_minimal(base_size = 12)

plot(publications_per_year_plot)

# Save to github
ggsave(filename = "publications_per_year.tiff",
       plot = publications_per_year_plot,
       path = gh_figs_dir
       )

# Save to dropbox folder
ggsave(filename = "fig2_publications_per_year.png",
       plot = publications_per_year_plot,
       path = dropbox_figs_dir
       )
```

# Human vaccine-preventable diseases

## Analysis of collaboration type compositions

### Aggregated
```{r}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- compact_data_with_citation_keys_no_fmd %>%
    count(collab_type, sort = TRUE) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(studies_per_collab_type, 
             caption = 'Number of studies per collaboration type (Human VPDs)',
             format = "latex", 
             booktabs = TRUE
             ) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Number of studies per year and broken down by collaboration types

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, collab_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = collab_type, values_from = num_of_studies) %>% 
  clean_names() %>% 
  adorn_totals(... = c(mixed, purely_academic), where = c("row", "col"))

knitr::kable(collap_types_by_year,
             caption = "Breakdown of studies per year by collaboration types")
```

```{r echo=FALSE}
#' Count by collaboration types and make a bar plot
collab_types_count <- compact_data_with_citation_keys_no_fmd %>%
  count(author_affiliation_type, sort = TRUE, name = "num_of_studies") %>% 
  mutate(author_affiliation_type = as_factor(author_affiliation_type)) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals() %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front")

knitr::kable(collab_types_count)
```

### Number of publications per year by collaboration type

```{r} 
 #' Count the number of each author affiliation type per year
 collabs_per_year <- compact_data_with_citation_keys_no_fmd %>%
   count(year_aggreg,
     collab_type,
     name = "total_publications"
   )
 x_axis_labels <- as.character(seq(2005, 2020, 2))

 collabs_per_year_plot <- ggplot(
   data = collabs_per_year,
   aes(
     x = year_aggreg,
     y = total_publications
   ),
   color = "black"
 ) +
   geom_bar(aes(fill = collab_type),
     stat = "identity",
     position = "dodge"
   ) +
   scale_fill_manual(values = c("tan3", "turquoise2")) +
   scale_x_continuous(
     breaks = seq(2005, 2020, 2),
     labels = x_axis_labels,
     expand = c(0, 0)
   ) +
   scale_y_continuous(
     breaks = seq(0, 20, 2),
     labels = seq(0, 20, 2)
   ) +
   labs(
     x = "Year",
     y = "Total publications",
     fill = "Collaboration type"
   ) +
   theme_minimal(base_size = 12) +
   theme(legend.position = "bottom")

 plot(collabs_per_year_plot)

 ggsave(
   filename = "total_collabs_per_year_plot.tiff",
   plot = collabs_per_year_plot,
   path = gh_figs_dir
 )

 ggsave(
   filename = "total_collabs_per_year_plot.tiff",
   plot = collabs_per_year_plot,
   path = dropbox_figs_dir
 )
```

Even though non-academic collaborations increased over time, the share of 
total publications per year over time decreased.

### Proportion of the total publications per year by collaboration type

```{r echo=FALSE, fig.width=8} 
#' Count the number of each author affiliation type per year (including FMD) 
collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>%
  count(year_aggreg,
    collab_type,
    name = "total_publications"
  )

# The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- "1970-2005"

collab_proportions_plot <- ggplot(data = collap_types_by_year) +
  geom_bar(
    aes(
      x = year_aggreg,
      y = total_publications,
      fill = collab_type
    ),
    color = "black",
    stat = "identity",
    position = position_fill(reverse = TRUE) # ,
    # show.legend = FALSE
  ) +
  scale_fill_manual(values = c("tan3", "turquoise2")) +
  scale_x_continuous(
    breaks = seq(2005, 2020, 2),
    labels = x_axis_labels,
    expand = c(0, 0)
  ) +
  labs(
    x = "Year",
    y = "Proportion of total publications",
    fill = "Collaboration type"
  ) +
  theme(legend.position = "none") +
  theme_minimal(base_size = 12)

plot(collab_proportions_plot)

ggsave(
  filename = "collab_proportions.tiff",
  plot = collab_proportions_plot,
  path = gh_figs_dir
)

ggsave(
  filename = "collab_proportions.tiff",
  plot = collab_proportions_plot,
  path = dropbox_figs_dir
)
```

### Mixed collaborations with and without academics

```{r echo=FALSE}
academic_collabs_versus_others <- compact_data_with_citation_keys_no_fmd %>%
  filter(author_affiliation_type != "academic_institutions") %>%
  mutate(
    year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year)),
    collab_type = as_factor(if_else(str_detect(
      author_affiliation_type,
      "academic_institutions"
    ),
    "including_academics",
    "excluding_academics"
    ))
  ) %>%
  count(year_aggreg, collab_type, name = "num_of_studies")

# The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- "1970-2005"

# plot the result
academic_collabs_trend <- ggplot(data = academic_collabs_versus_others) +
  geom_bar(
    aes(
      x = year_aggreg,
      y = num_of_studies,
      fill = collab_type
    ),
    color = "black",
    stat = "identity",
    position = position_fill(reverse = TRUE)
  ) +
  scale_fill_manual(values = c("tan1", "turquoise1")) +
  scale_x_continuous(
    breaks = seq(2005, 2020, 2),
    labels = x_axis_labels
  ) +
  coord_flip() +
  labs(
    title = "Proportion of mixed collaborations per year",
    x = "Year",
    y = "Proportion of publications",
    fill = "Mixed collaboration type"
  ) +
  theme_minimal(base_size = 12)

plot(academic_collabs_trend)
```

## Geographic analyses

### Location studied

#### Real versus hypothetical

```{r}
location_types_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    mutate(location_type = ifelse(country_studied == 'none', 'none', 'real')) %>% 
    count(location_type) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(location_types_tally, 
             caption = 'Number of studies per location type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

#### Larger than one country

```{r}
larger_geographies_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    filter(country_studied %in% c('west_africa', 'global', 
                                  'northern_hemisphere', 'southeast_asia', 
                                  'who_southeast_asia_region')
           ) %>% 
    count(country_studied) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(larger_geographies_tally, 
             caption = 'Number of studies per location larger than one country', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

#### Countries studied

```{r}
countries_studied_by_count <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>% 
    count(collab_type, country_studied, name = 'num_of_studies', sort = T)

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  group_by(country_studied) %>%
  summarise(num_of_studies = sum(num_of_studies),
            .groups = 'drop'
            ) %>% 
  filter(!country_studied %in% c('none', 'west_africa', 'global', 
                                 'northern_hemisphere', 'southeast_asia', 
                                 'who_southeast_asia_region')
         ) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  ungroup()

#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             caption = 'Number of studies per country', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 
                                                    'HOLD_position'), 
                                  font_size = 7)
```

#### Continents studied

```{r}
continents_studied_by_count <- countries_studied_by_count %>% 
  filter(!country_studied %in% c('none', 'northern_hemisphere', 'global')) %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 
                                                  'asia') == TRUE ~ 'asia', 
                                         str_detect(continent_studied, 
                                                    'africa') == TRUE ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            .groups = 'drop'
            ) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

#Print the results
knitr::kable(continents_studied_by_count, 
             caption = 'Number of studies per continent', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Connection of authors to the studied locations

```{r}
author_connectedness_top_5_countries <- compact_data_with_citation_keys_no_fmd %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(!country_studied %in% c('none', 'west_africa', 'global', 
                                 'northern_hemisphere', 'southeast_asia', 
                                 'who_southeast_asia_region')) %>% 
  separate_rows(country_studied,  sep = ' ') %>% 
  group_by(collab_type) %>% 
  count(country_studied, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  arrange(collab_type, desc(yes)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no')) %>% 
  ungroup()

knitr::kable(author_connectedness_top_5_countries, 
             caption = 'Studies with at least one author affiliation in the location studied', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position', 
                                                    'repeat_header'), 
                                  font_size = 8)
```

### Author affiliation in the studied location (overall)

```{r}
author_connectedness_to_location_aggreg <- compact_data_with_citation_keys_no_fmd %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(!country_studied %in% c('other', 'none', 'northern_hemisphere', 
                                 'west_africa', 'global')) %>% 
  count(collab_type, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(author_connectedness_to_location_aggreg, 
             caption = 'At least one author with an affiliation in the studied location', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Diseases studied per collab type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_collab_type <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease,  sep = ' ') %>%
  distinct(title, disease, .keep_all = TRUE) %>%
  count(collab_type, disease, sort = TRUE) %>%
  pivot_wider(names_from = collab_type, values_from = n) %>%
  replace_na(list(purely_academic = 0, mixed = 0)) %>%
  adorn_totals(where = c('row', 'col')) %>%
  adorn_percentages(... = c('purely_academic', 'mixed')) %>%
  adorn_pct_formatting(... = c('purely_academic', 'mixed')) %>%
  adorn_ns(... = c('purely_academic', 'mixed')) %>%
  mutate(disease = str_to_sentence(str_replace(disease, '_', ' '))) %>%
  arrange(desc(Total))

disease_studied_by_collab_type_mod <- rbind(disease_studied_by_collab_type[-1, ], 
                   disease_studied_by_collab_type[1, ])

knitr::kable(disease_studied_by_collab_type_mod, 
             caption = 'Number of studies per disease and collaboration type', 
             format = "latex", booktabs = TRUE) %>%
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r}
#tally the diseases by collaboration type
disease_by_collab_type_long <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease,  sep = ' ') %>%
  count(collab_type, disease, sort = TRUE)

disease_by_collab_type_plot <- ggplot(disease_by_collab_type_long) +
  geom_bar(aes(x = reorder(disease, n),
               y = n,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)
           ) +
  labs(x = 'Disease', y = 'Number of studies') +
  coord_flip() +
  theme_minimal(base_size = 12)

print(disease_by_collab_type_plot)
```

## Interventions

### Types of interventions

```{r}
intervention_categories <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, intervention_type, sort = TRUE) %>% 
  pivot_wider(names_from = intervention_type, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no_vax', 
                            'vax_as_mixed_intervention', 
                            'vax_compared_to_others')
                    ) %>% 
  adorn_pct_formatting(... = c('no_vax', 
                               'vax_as_mixed_intervention', 
                               'vax_compared_to_others')
                       ) %>% 
  adorn_ns(... = c('no_vax', 
                   'vax_as_mixed_intervention', 
                   'vax_compared_to_others'
                   )
           )

knitr::kable(intervention_categories, 
             caption = 'Number of studies per intervention categories', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Diseases and interventions

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup()

#interventions are disease specific
interventions_per_disease <- interventions_unique_entries %>% 
  mutate(intervention_modelled = str_trim(intervention_modelled)) %>% 
  filter(intervention_modelled != 'e.g.') %>% #remove false rows
  filter(intervention_modelled != 'various') %>% #remove false rows
  count(disease, intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  ungroup()

knitr::kable(interventions_per_disease)
```
### Diseases and interventions (top 2 per disease)

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_per_disease_top2 <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup() %>% 
  count(disease, intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  group_by(disease) %>% 
  slice_head(n = 2) %>% #keep only top 2 per group 
  ungroup()

knitr::kable(interventions_per_disease_top2,
             caption = "Top 2 most studied intervention per disease")
```

### Impact of vaccination

```{r}
vax_impact_conclusions <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_compared_to_others') %>% 
  count(collab_type, is_vax_effective, sort = TRUE) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
    replace_na(list(yes = 0, n0 = 0, the_outcomes_were_inconclusive = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(vax_impact_conclusions, 
             caption = 'Conclusions about impact of vaccination', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position')
                                  )
```

```{r eval=FALSE}
no_vax_modelled <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(disease,  sep = ' ') %>% 
    filter(vax_modelled == 'no') %>% 
    count(disease, sort = TRUE, name = 'num_of_studies') %>% 
     adorn_totals(where = 'row') %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(no_vax_modelled, 
             caption = sprintf("%s %s", 
                               'Number of studies per disease for',
                               'which vaccination was not modelled'
                               ), 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
vax_impact_by_disease <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_compared_to_others') %>% 
  count(disease, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
    replace_na(list(yes = 0, no = 0, inconclusive = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(vax_impact_by_disease, 
             caption = 'Conclusions about impact of vaccination per disease', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
vax_as_single_intervention_studies <- compact_data_with_citation_keys_no_fmd %>% 
  filter(intervention_type == 'vax_compared_to_others') %>% 
  select(title, author, publication_year, disease, paper_doi, 
         intervention_modelled)

#Print the results
# knitr::kable(no_vax_impact_by_disease_citation_table, 
#              caption = 'Studies that did not find vaccination as most impactful'
#              )

writexl::write_xlsx(vax_as_single_intervention_studies, 
          path = file.path("citation_tables", 
                           "studies_modelling_vax_for_comparison.xlsx"
                           )
           )
```

## Modelling objectives 
<!-- We grouped studies into whether they assessed the impact of a future 
intervention strategy (prospective), a past intervention, or both (dual).  -->

<!-- In general, there were more (73.2%; 167/228) studies about prospective 
intervention impact assessments than retrospective assessments (24.6%; 56/228), 
and the dual (2.2%; 5/228). A similar pattern was observed across the two 
collaboration types except that purely academic collaborations were more 
likely to perform the dual assessments.   -->

<!-- In terms of the outbreak types, there was an approximately equal number 
of papers about real and hypothetical outbreaks overall, and there was no 
clear difference between the two collaborations.   -->

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_count <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('prospective_analysis', 'retrospective_analysis')) %>% 
  adorn_pct_formatting(... = c('prospective_analysis', 'retrospective_analysis')) %>% 
  adorn_ns(... = c('prospective_analysis', 'retrospective_analysis'),
           position = "front")

#Print a table of the results
knitr::kable(modelling_objectives_count,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Outbreak types

```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_pct_formatting(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_ns(... = c('hypothetical_outbreak', 'real_outbreak'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type,
             caption = 'outbreak types by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
objectives_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  count(objectives, collab_type, name = 'num_of_studies', sort = TRUE) %>% 
  pivot_wider(names_from = objectives, 
              values_from = num_of_studies
              ) %>% 
  replace_na(list(prospective_analysis = 0, retrospective_analysis = 0, both = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('prospective_analysis', 'retrospective_analysis', 'both')) %>% 
  adorn_pct_formatting(... = c('prospective_analysis', 'retrospective_analysis', 'both')) %>% 
  adorn_ns(... = c('prospective_analysis', 'retrospective_analysis', 'both'))

knitr::kable(objectives_by_collab_type, 
             caption = 'Number of studies per objective by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Model characteristics  -->

## Individual heterogeneity: agent-based versus compartmental models

```{r}
individual_representation <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, individual_representation, sort = TRUE) %>% 
  pivot_wider(names_from = individual_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('compartments', 'agents', 
                            'individuals_representation_other')
                    ) %>% 
  adorn_pct_formatting(... = c('compartments', 'agents', 
                               'individuals_representation_other')
                       ) %>% 
  adorn_ns(... = c('compartments', 'agents', 
                   'individuals_representation_other')
           )

knitr::kable(individual_representation, 
             caption = 'How individuals were represented', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```
<!-- ### Individual representation by model structure -->
<!-- ```{r} -->
<!-- mod_structure_vs_individual_representation <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   group_by(collab_type) %>%  -->
<!--   count(model_structure, individual_representation, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = model_structure, values_from = n) %>%  -->
<!--   adorn_totals(where = c('col', 'row')) -->

<!-- knitr::kable(mod_structure_vs_individual_representation, caption = 'Model structure versus individual heterogeneity') -->
<!-- ``` -->

## Spatial heterogeneity

```{r}
space_representation <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, space_representation, sort = TRUE) %>% 
  pivot_wider(names_from = space_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(space_representation, 
             caption = 'Was space explicitly represented in the model',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ### Spatial models by model structure -->
<!-- ```{r} -->
<!-- mod_structure_vs_space_representation <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   group_by(collab_type) %>%  -->
<!--   count(model_structure, space_representation, sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = model_structure, values_from = n) %>%  -->
<!--   adorn_totals(where = c('col', 'row')) -->

<!-- knitr::kable(mod_structure_vs_space_representation, caption = 'Model structure versus spatial heterogeneity') -->
<!-- ``` -->

## Model dynamics: deterministic vs stochastic

```{r}
stochastic_vs_deterministic <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, model_structure) %>% 
  pivot_wider(names_from = model_structure, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('both', 'deterministic', 'stochastic')) %>% 
  adorn_pct_formatting(... = c('both', 'deterministic', 'stochastic')) %>% 
  adorn_ns(... = c('both', 'deterministic', 'stochastic'))

knitr::kable(stochastic_vs_deterministic, 
             caption = 'Model dynamics (deterministic versus stochastic)', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Outcomes measured

<!-- We tallied the overall top five outcomes most used between the 
two collaboration types.  -->

<!-- The top five most commonly used outcomes, irrespective of collaboration 
type, were: final epidemic size, attack rate, timing of epidemic peak, 
cases averted, and outbreak duration. -->

<!-- There was no difference in the two collaboration types as they were 
commonly used by both groups. -->

```{r}
outcomes_measured_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(collab_type, outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  filter(!outcome_measured %in% c('outcome_other', '', 's')) %>% 
  count(collab_type, outcome_measured, name = 'num_of_studies') %>% 
  ungroup(outcome_measured) %>% 
  arrange(desc(num_of_studies), .by_group = TRUE)

knitr::kable(outcomes_measured_by_collab_type, 
             caption = 'Outcomes studied by the collaboration types', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 
                                                    'HOLD_position', 
                                                    'repeat_header')
                                  )
```

### Top 6 outcomes measured by collaboration type
```{r}
outcomes_measured_by_collab_type_top6 <- outcomes_measured_by_collab_type %>% 
  group_map(~ {
    .x %>% 
      adorn_totals() %>% 
      adorn_percentages(denominator = "col") %>%
      adorn_pct_formatting() %>%
      adorn_ns()},
    .keep = TRUE) %>% 
  map(head, n = 6) %>% 
  list_rbind()

knitr::kable(outcomes_measured_by_collab_type_top6, 
             caption = 'Top 6 outcomes studied by the collaboration types', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 
                                                    'HOLD_position', 
                                                    'repeat_header')
                                  )
```

## Parametrization methods

```{r}
#wide data for table
model_parametrization_wide <- compact_data_with_citation_keys_no_fmd %>%
  count(collab_type, parametrization, sort = TRUE) %>%
  pivot_wider(names_from = parametrization, values_from = n) %>%
  adorn_totals(where = c('row', 'col')) %>%
  adorn_percentages(... = -c('Total', 'collab_type')) %>%
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>%
  adorn_ns(... = -c('Total', 'collab_type'))

#table
knitr::kable(model_parametrization_wide, 
             caption = 'How model parameters are obtained',
             format = "latex", booktabs = TRUE) %>%
        kableExtra::kable_styling(latex_options = c('striped', 
                                                    'scale_down', 
                                                    'HOLD_position')
                                  )
```

## Validation methods

```{r}
#wide data for table
model_validation_wide <- compact_data_with_citation_keys_no_fmd %>%
  count(collab_type, validation, sort = TRUE) %>%
  pivot_wider(names_from = validation, values_from = n) %>%
  replace_na(list(none = 0,
                  data = 0,
                  another_model = 0,
                  none_and_data = 0,
                  data_and_another_model = 0
                  )
             ) %>%
  adorn_totals(where = c('row', 'col')) %>%
  adorn_percentages(... = -c('Total', 'collab_type')) %>%
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>%
  adorn_ns(... = -c('Total', 'collab_type'))

#table
knitr::kable(model_validation_wide, 
             caption = 'How the model\'s performance is assessed',
             format = "latex", booktabs = TRUE) %>%
        kableExtra::kable_styling(latex_options = c('striped', 
                                                    'scale_down', 
                                                    'HOLD_position')
                                  )
```

## Sensitivity analysis

<!-- Less than half (46.5%; 106/228) of the total studies performed sensitivity analyses. There was no clear difference between the collaborations.  -->

```{r eval=FALSE}
mod_structure_vs_sensitivity_analysis <- compact_data_with_citation_keys_no_fmd %>% 
  count(model_structure, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(mod_structure_vs_sensitivity_analysis, 
             caption = 'Sensitivity analysis in relation to model structure', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

```{r}
sensitivity_analysis <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(sensitivity_analysis, caption = 'Sensitivity analysis', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Data use and data availability

```{r}
data_use_and_availability <- compact_data_with_citation_keys_no_fmd %>% 
  filter(data_used == 'yes') %>%
  count(collab_type, data_available, sort = TRUE) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(data_use_and_availability, 
             caption = 'Data accessibility', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Code availability

```{r}
code_availability <- compact_data_with_citation_keys_no_fmd %>% 
  count(collab_type, simulation_code_available) %>% 
  pivot_wider(names_from = simulation_code_available, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(code_availability, caption = 'Code availability', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

```{r eval=FALSE}
papers_with_code_available <- compact_data_with_citation_keys_no_fmd %>% 
    filter(simulation_code_available == 'yes') %>% 
  select(title, author, publication_year, paper_doi, publication_type, journal)

writexl::write_xlsx(papers_with_code_available, 
                     path = file.path("citation_tables", 
                                      "papers_with_code_available.xlsx")
                    )
```

# Foot and mouth disease (FMD)

<!-- ## Collaboration patterns in time -->

## Unique combinations of author affiliation types

```{r}
#' Count by collaboration types and make a bar plot
author_affiliation_type_combos <- fmd_data %>%
    count(author_affiliation_type, sort = TRUE) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type)) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(author_affiliation_type_combos, caption = 'Number of studies by author affiliation type combination',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Collaboration types (aggregated)

```{r}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- fmd_data %>%
    count(collab_type, sort = TRUE) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(studies_per_collab_type, caption = 'Number of studies per collaboration type',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Proportions of the total publications per year by collaboration type

```{r echo=FALSE, fig.width=8}
#' Count the number of each author affiliation type per year (including FMD)
fmd_collab_types_per_year <- fmd_data %>% 
  count(year_aggreg, collab_type, name = 'total_publications')

#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'

fmd_collab_proportions_plot <- ggplot(data = fmd_collab_types_per_year) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)#,
           #show.legend = FALSE
             ) +
  scale_fill_manual(values = c('tan3', 'turquoise2')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  labs(x = 'Year',
       y = 'Proportion of total publications',
       fill = '**Collaboration type**' 
         ) +
  theme(legend.position = 'none')

plot(fmd_collab_proportions_plot)

```

### Absolute number of publications per year by collaboration type

```{r}
#' Count the number of each author affiliation type per year
fmd_collabs_per_year <- fmd_data %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

collabs_per_year_plot <- ggplot(data = fmd_collabs_per_year, 
                                    aes(x = year_aggreg, 
                                        y = total_publications
                                        ), 
                                    color = 'black') + 
  geom_bar(aes(fill = collab_type), 
           stat = 'identity', 
           position = 'dodge',
           width = 0.5
           ) + 
  scale_fill_manual(values = c('tan3', 'turquoise2')) + 
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  labs(x = 'Year', 
       y = 'Total publications', 
       fill = 'Collaboration type'
       ) +
  theme(legend.position = 'bottom')

plot(collabs_per_year_plot)

```

## Geographic analyses

### Connection of authors to the studied locations 
```{r}
author_connectedness_to_location_fmd <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(!country_studied %in% c('none', 'west_africa', 'global', 
                                 'northern_hemisphere', 'southeast_asia',
                                 'who_southeast_asia_region')) %>% 
  separate_rows(country_studied,  sep = ' ') %>% 
  group_by(collab_type) %>% 
  count(country_studied, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  arrange(collab_type, desc(yes)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no')) %>% 
  ungroup()

knitr::kable(author_connectedness_to_location_fmd, 
             caption = 'Studies with at least one author affiliation in the location studied', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Aggregated

```{r}
author_connectedness_to_location_aggreg <- fmd_data %>% 
  filter(author_in_country_studied != 'not_applicable') %>% 
  filter(!country_studied %in% c('other', 'none', 'northern_hemisphere', 
                                 'west_africa', 'global')
         ) %>% 
  count(collab_type, author_in_country_studied, sort = TRUE) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(author_connectedness_to_location_aggreg, 
             caption = 'Studies with at least one author affiliation in the location studied', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Interventions

### Types of interventions

```{r}
intervention_categories_fmd <- fmd_data %>% 
  count(collab_type, intervention_type, sort = TRUE) %>% 
  pivot_wider(names_from = intervention_type, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no_vax', 'vax_as_mixed_intervention', 'vax_compared_to_others')) %>% 
  adorn_pct_formatting(... = c('no_vax', 'vax_as_mixed_intervention', 'vax_compared_to_others')) %>% 
  adorn_ns(... = c('no_vax', 'vax_as_mixed_intervention', 'vax_compared_to_others'))

knitr::kable(intervention_categories_fmd, caption = 'Number of studies per intervention type', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

### Impact of vaccination

```{r}
fmd_vax_impact_conclusions <- fmd_data %>% 
  filter(intervention_type == 'vax_compared_to_others') %>% 
  count(collab_type, is_vax_effective, sort = T) %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  replace_na(list(no = 0, inconclusive = 0, yes = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_pct_formatting(... = c('yes', 'no', 'inconclusive')) %>% 
  adorn_ns(... = c('yes', 'no', 'inconclusive'))

#Print the results
knitr::kable(fmd_vax_impact_conclusions, 
             caption = 'Conclusions about impact of vaccination', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Modelling objectives

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_fmd <- fmd_data %>% 
  count(collab_type, objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('prospective_analysis', 'retrospective_analysis')) %>% 
  adorn_pct_formatting(... = c('prospective_analysis', 'retrospective_analysis')) %>% 
  adorn_ns(... = c('prospective_analysis', 'retrospective_analysis'))

#Print a table of the results
knitr::kable(modelling_objectives_fmd,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ### Modelling objectives and outbreak type by collaboration types -->

<!-- ```{r echo=FALSE} -->
<!-- objectives_and_outbreak_types_fmd <- fmd_data %>%  -->
<!--   count(objectives, outbreak_type, name = 'num_of_studies', sort = TRUE) %>%  -->
<!--   pivot_wider(names_from = objectives,  -->
<!--               values_from = num_of_studies -->
<!--               ) %>%  -->
<!--   replace_na(list(prospective_analysis = 0, retrospective_analysis = 0, both = 0)) %>%  -->
<!--   adorn_totals(where = c('row', 'col')) %>%  -->
<!--   adorn_percentages(... = c('prospective_analysis', 'retrospective_analysis')) %>%  -->
<!--   adorn_pct_formatting(... = c('prospective_analysis', 'retrospective_analysis')) %>%  -->
<!--   adorn_ns(... = c('prospective_analysis', 'retrospective_analysis'))  -->

<!-- knitr::kable(objectives_and_outbreak_types_fmd,  -->
<!--              caption = 'Number of studies per objective and outbreak type',  -->
<!--              format = "latex", booktabs = TRUE) %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position')) -->
<!-- ``` -->

## Outbreak types

```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type_fmd <- fmd_data %>% 
  count(collab_type, outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_pct_formatting(... = c('hypothetical_outbreak', 'real_outbreak')) %>% 
  adorn_ns(... = c('hypothetical_outbreak', 'real_outbreak'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type_fmd,
             caption = 'Study objectives by collaboration type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

<!-- ## Model characteristics  -->

## Individual heterogeneity: agent-based versus compartmental models

```{r}
individual_representation_fmd <- fmd_data %>% 
  count(collab_type, individual_representation, sort = TRUE) %>% 
  pivot_wider(names_from = individual_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('compartments', 'agents')) %>% 
  adorn_pct_formatting(... = c('compartments', 'agents')) %>% 
  adorn_ns(... = c('compartments', 'agents'))

knitr::kable(individual_representation_fmd, 
             caption = 'How individuals are represented', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Spatial heterogeneity

```{r}
space_representation_fmd <- fmd_data %>% 
  count(collab_type, space_representation, sort = TRUE) %>% 
  pivot_wider(names_from = space_representation, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(space_representation_fmd, caption = 'Models with spatial structure', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Model dynamics: deterministic vs stochastic

```{r}
stochastic_vs_deterministic_fmd <- fmd_data %>% 
  count(collab_type, model_structure) %>% 
  pivot_wider(names_from = model_structure, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('deterministic', 'stochastic')) %>% 
  adorn_pct_formatting(... = c('deterministic', 'stochastic')) %>% 
  adorn_ns(... = c('deterministic', 'stochastic'))

knitr::kable(stochastic_vs_deterministic_fmd, 
             caption = 'Model dynamics (deterministic versus stochastic)', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

<!-- ## Modelling methods -->

## Outcomes measured
```{r}
outcomes_tally_by_collab_type <- fmd_data %>% 
  mutate(outcome_measured = str_trim(outcome_measured)) %>% 
  separate_rows(outcome_measured, sep = ',') %>% 
  group_by(collab_type, outcome_measured) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  filter(!outcome_measured %in% c('outcome_other', '')) %>% 
  count(collab_type, outcome_measured, name = 'num_of_studies') %>% 
  ungroup(outcome_measured) %>% 
  arrange(desc(num_of_studies), .by_group = TRUE) %>% 
  ungroup() %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(outcomes_tally_by_collab_type, 
             caption = 'Model outcomes by collaboration type', 
             format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position', 'repeat_header'))
```

```{r eval=FALSE}
collab_type_labels <- c('purely_academic' = 'Purely academic', 
                        'mixed' = 'Mixed'
                        )

outcomes_tally_plot <- ggplot(data = outcomes_tally_by_collab_type, 
                              aes(label = outcome_measured, 
                                  size = n
                                  )
                              ) + 
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 7) + 
  facet_wrap(~ collab_type, 
             labeller = as_labeller(collab_type_labels),
             strip.position = 'top'
             ) +
  #theme_minimal() + 
  theme(strip.text = element_text(size = 15, face = 'bold'))

#save the plot
ggsave(filename = 'outcomes_tally_fmd_plot.tiff',
       plot = outcomes_tally_plot, 
       device = 'tiff',
       path = figs_dir
       )
```

## Parametrization methods

```{r}
#wide data for table
model_parametrization_fmd_wide <- fmd_data %>% 
  count(collab_type, parametrization, sort = TRUE) %>% 
  pivot_wider(names_from = parametrization, values_from = n) %>% 
  replace_na(list('expert_opinion and fitted' = 0, 'literature and fitted' = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type')) 
#table
knitr::kable(model_parametrization_fmd_wide, 
             caption = 'Model parametrization methods', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'scale_down',  'HOLD_position'))
```

## Validation methods

```{r}
#wide data for table
model_validation_fmd_wide <- fmd_data %>% 
  count(collab_type, validation, sort = TRUE) %>% 
  pivot_wider(names_from = validation, values_from = n) %>% 
  replace_na(list(none = 0, 
                  data = 0, 
                  another_model = 0, 
                  data_and_another_model = 0
                  )
             ) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = -c('Total', 'collab_type')) %>% 
  adorn_pct_formatting(... = -c('Total', 'collab_type')) %>% 
  adorn_ns(... = -c('Total', 'collab_type'))

#table
knitr::kable(model_validation_fmd_wide, 
             caption = 'Model validation methods', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped',  'HOLD_position'))
```

## Sensitivity analysis

```{r}
sensitivity_analysis_fmd <- fmd_data %>% 
  count(collab_type, sensitivity_analysis, sort = TRUE) %>% 
  pivot_wider(names_from = sensitivity_analysis, values_from = n) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('no', 'yes')) %>% 
  adorn_pct_formatting(... = c('no', 'yes')) %>% 
  adorn_ns(... = c('no', 'yes'))

knitr::kable(sensitivity_analysis_fmd, caption = 'Sensitivity analysis', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped',  'HOLD_position'))

```

## Data use and data availability

```{r}
data_use_and_availability_fmd <- fmd_data %>% 
  filter(data_used == 'yes') %>%
  count(collab_type, data_available, sort = TRUE) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0)) %>% 
  adorn_totals(where = c('row', 'col')) %>% 
  adorn_percentages(... = c('yes', 'no')) %>% 
  adorn_pct_formatting(... = c('yes', 'no')) %>% 
  adorn_ns(... = c('yes', 'no'))

knitr::kable(data_use_and_availability_fmd, 
             caption = 'Data accessibility', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

## Code availability

```{r}

code_availability_fmd <- fmd_data %>% 
  count(collab_type, simulation_code_available) %>% 
  pivot_wider(names_from = simulation_code_available, values_from = n) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(code_availability_fmd, caption = 'Code availability', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```