### 2023-09-02 manuscript/submission2/submission_files/supplementary/S1 File.tex
Added v1 and v2 of submitted manuscripts and associated files

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{float}
\usepackage{rotating}
\usepackage{pdfpages}

% Title must be 250 characters or less.
\title{S1 File }

\begin{document}
\maketitle

\section{Human vaccine-preventable diseases considered}

\begin{table}[H]
	\centering
	\caption{The World Health Organization (WHO) list of diseases with an available vaccine as of Jan 1, 2020.}
	\begin{tabular}{ll}
		\toprule
		Cholera	& Mumps \\
		Dengue	&  Pertussis\\
		%	\hline
		Diphtheria	& Pneumococcal disease \\
		%	\hline
		Hepatitis A		&  Poliomyelitis\\
		%	\hline
		Hepatitis B		& Rabies \\
		%	\hline
		Hepatitis E	& 	Rotavirus \\
		%	\hline
		Haemophilus influenzae type b (Hib)		& Rubella \\
		%	\hline
		Human papillomavirus (HPV)	&  	Tetanus\\
		%	\hline
		Influenza	&  Tick-borne encephalitis	\\
		%	\hline
		Japanese encephalitis		&  Tuberculosis\\
		%	\hline
		Malaria		&  Typhoid\\
		%	\hline
		Measles		& Varicella \\
		%	\hline
		Meningococcal meningitis		&  Yellow Fever\\
		\bottomrule
	\end{tabular}
	\label{chap2-table: The World Health Organization (WHO) list of diseases with an available vaccine}
\end{table}

\section{Data items}
The following information will be extracted from each of the included studies:
\begin{enumerate}
		\item Disease(s) modelled
		\item Study's main objective(s)
		\item Location studied and whether the first or last author has an affiliation in the location
		\item Model dynamics: deterministic, stochastic, or a hybrid
		\item How individuals are represented in model: individual-bases or compartmental
		\item Whether spatial structure was included
		\item Interventions modelled: vaccination (ring, mass, targeted, prophylactic, pulse) or non-vaccination (movement restrictions, palliative care, quarantine, isolation, treatment, other(s) as indicated)
		\item Model parametrisation (We define model parametrization as the process of obtaining the model's parameter values either from literature or through some kind of mathematical or statistical technique)
		\item Method of model validation with some form of data. (We define model validation as any procedure used to evaluate the model’s suitability for the data and problem being studied)
		\item Outcome measured with model
		\item Authors’ conclusion on the predicted or potential impact of the intervention
		\item Timing of modelling practice: retrospective, real-time, other(s) as indicated
		\item Sensitivity analysis performed or not
		\item Data and code availability
	\end{enumerate}

\section{Questionnaire for data extraction}
	\subsection{Citation information}
	\begin{enumerate}
		\item What is the title of this paper?
		\item What is the year of publication? 
		\item What is the DOI of the paper? 
	\end{enumerate}

	\subsection{Collaboration }
	\begin{enumerate}[resume]
		\item Which of the following describe the nature of the author affiliations? 
		\begin{itemize}
			\item Academic institutions (universities, research institutions, etc)
			\item Government institutions (government ministries or departments, etc)
			\item Non-governmental institutions 
		\end{itemize}
		\item Which country is this study based on? [Dropdown menu of all countries including an option for hypothetical, and other, if multiple. Please specify]
		\item Are any of the author affiliations situated in the country being studied? 
		\begin{itemize}
			\item Yes 
			\item No
		\end{itemize}
	\end{enumerate}

	\subsection{Introduction}
	\subsubsection{Disease}
	\begin{enumerate}[resume]
		\item Which disease(s) was modelled in this study?
		\item Which of the following will describe the study’s broad objective(s)? 
		\begin{itemize}
			\item Assessing the potential impact/effectiveness of past interventions
			\item Assessing the potential impact/effectiveness of future/proposed interventions
			\item Other. Please indicate.
		\end{itemize}
	\end{enumerate}

\subsection{Methods}
	\subsubsection{Model structure }
	\begin{enumerate}[resume]
		\item How are individuals represented in the model? 
		\begin{itemize}
			\item Compartments, where identical individuals are grouped according to disease state
			\item Agent/Individual-based, where each agent is captured individually
		\end{itemize}
		\item Is space explicitly represented in the model? 
		\begin{itemize}
			\item Yes
			\item No
			\item Unclear. Please indicate why.
		\end{itemize}
		\item Which of the following describes the model structure?
		\begin{itemize}
			\item Deterministic
			\item Stochastic
			\item Other. Please indicate
		\end{itemize}
	\end{enumerate}

	\subsubsection{Calibration and validation}
	\begin{enumerate}[resume]
		\item How were the model’s parameter values determined? 
		\begin{itemize}
			\item Values were sourced from various literature sources, including, systematic reviews studies reporting on primary data
			\item Values were assumed/guessed based on a plausible range (expert opinion)
			\item Values were fitted to a time series 
		\end{itemize}
		\item Which of the following best describe the model validation process, that is, the process of evaluating the model’s ability to describe the dynamics of the disease?
		\begin{itemize}
			\item The model was not validated. (This includes where one set of the model’s runs were compared to another set of runs from the same model.)
			\item The model's output was compared to an independently observed dataset
			\item The model's output was compared to at least one other model's output
			\item Other. Please indicate
		\end{itemize}
	\end{enumerate}

	\subsubsection{Outbreak and intervention }
	\begin{enumerate}[resume]
		\item What type of outbreak was this study based on? 
		\begin{itemize}
			\item A real outbreak that occurred 
			\item A hypothetical outbreak
		\end{itemize}
		\item Which of the following would describe the type of intervention modelled? 
		\begin{multicols}{2}
			\begin{itemize}
				\item Antiviral prophylaxis
				\item Bednets
				\item Behaviour change
				\item Contact tracing
				\item Culling
				\item Education
				\item Face masks
				\item Handwashing
				\item Hospital bed capacity
				\item Improved hygiene
				\item Indoor/outdoor spraying
				\item Isolation
				\item Larvicides
				\item Media campaign
				\item Palliative care
				\item Protective clothing (PPE, other than face masks)
				\item Quarantine
				\item Safe burial
				\item Social/physical distancing
				\item School closure
				\item Testing/screening
				\item Travel restriction/border control
				\item Treatment
				\item Travel screening
				\item Vaccination
				\item Ventilation
				\item Workplace closure
				\item 	Other. Please indicate.
			\end{itemize}
		\end{multicols}
	\end{enumerate}

	\subsubsection{Impact of interventions}
	\begin{enumerate}[resume]
		\item Was vaccination modelled for comparison with other interventions? 
		\begin{itemize}
			\item Yes 
			\item No
		\end{itemize}
	\end{enumerate}

	\begin{enumerate}[resume]
		\item If yes, was vaccination found to be the most effective strategy in terms of the stated objective? 
		\begin{itemize}
			\item 	Yes
			\item 	No
			\item 	The outcomes were inconclusive
		\end{itemize}
		\item What was the timing of the modelling practice? 
		\begin{itemize}
			\item Prospective
			\item Retrospective
			\item In real-time during the outbreak
		\end{itemize}
		\item Which outcome was measured?
		\begin{multicols}{2}
			\begin{itemize}
				\item Cases averted
				\item Intervention coverage
				\item Lives saved
				\item Deaths prevented/averted
				\item Campaign duration
				\item Outbreak duration/timing
				\item Final epidemic size/Outbreak size/Total cases
				\item Cost (direct cost/cost-effectiveness/cost-benefit)
				\item Timing of peak
				\item Attack rate
				\item Hospitalizations
				\item QUALYS
				\item DALYs
				\item Case-fatality
				\item Other. Please indicate
			\end{itemize}
		\end{multicols}
	\end{enumerate}

	\subsubsection{Sensitivity analysis}
	\begin{enumerate}[resume]
		\item Was a sensitivity analysis performed? 
		\begin{itemize}
			\item Yes
			\item No
		\end{itemize}
	\end{enumerate}

	\subsubsection{Reproducibility/transparency}
	\textbf{Data}
	\begin{enumerate}[resume]
		\item	Was any independently observed data used in this modelling study? (Independently observed data implies data that was not simulated with the model being used for evaluation in this study.)
		\begin{itemize}
			\item No
			\item	Yes 
		\end{itemize}
		\item	If Yes, Is the independently observed data openly available? 
		\begin{itemize}
			\item	Yes
			\item	No
			\item	Nothing about this was mentioned
		\end{itemize}
	\end{enumerate}

	\subsubsection{Code}
	\begin{enumerate}[resume]
		\item	Is the simulation code openly available?
		\begin{itemize}
			\item Yes (supplementary materials, git, zenodo, google drive, etc)
			\item No, including studies in where the authors indicated they could be emailed for the simulation code
		\end{itemize}
	\end{enumerate}

\newpage	
\section{Search strings and results}
Below are the detailed search strategies and results for the three bibliographic databases used:

\begin{sidewaysfigure}
	\begin{table}[H]
		\centering
		\caption{Scopus search strings and results.}
		\begin{tabular}{lp{0.85\textwidth}l}
			\toprule
			\textbf{Search} & \textbf{Query} & \textbf{Records found} \\
			\midrule
			\#1 & TITLE-ABS-KEY ( epidemic  OR    outbreak  OR  emergency    OR  reactive  OR    crisis ) & 1,834,598 \\
			\#2 & TITLE-ABS-KEY ( respon*  OR    manage*  OR  control    OR  interven*  OR    strateg* ) & 19,366,062 \\
			\#3 & TITLE-ABS-KEY ( stochastic  OR    transmission  OR  computational  OR    mathematical  OR  mechanistic    OR  statistical  OR    simulation  OR  "In silico"  OR    dynamic* ) & 11\,890\,240 \\
			\#4 & TITLE-ABS-KEY ( model* ) & 12\,504\,649 \\
			\#5 & TITLE-ABS-KEY ( cholera  OR    dengue  OR  diphtheria    OR  ebola  OR    "Foot-and-mouth"    OR  "foot and mouth"  OR    fmd  OR  "Hepatitis A"  OR    "Hepatitis B"  OR  "Hepatitis E"  OR  "Haemophilus   influenzae type b"  OR  hib    OR  "Human   papillomavirus"  OR  hpv    OR  influenza ) )  OR   (   TITLE-ABS-KEY ( "Japanese encephalitis"  OR    malaria  OR  measles    OR  "Meningococcal   meningitis"   OR  mumps    OR  pertussis  OR    "Whooping cough"    OR  "Pneumococcal   disease"  OR  poliomyelitis  OR    polio  OR  rabies    OR  rotavirus  OR    rubella ) )  OR  ( TITLE-ABS-KEY ( tetanus  OR    "Tick-borne encephalitis"    OR  tuberculosis  OR    typhoid  OR  varicella    OR  chickenpox  OR    "Yellow Fever"    OR    "vaccine-preventable" ) )& 1\,123\,110 \\
			\#6 & TITLE-ABS-KEY ( epidemic  OR    outbreak  OR  emergency    OR  reactive  OR    crisis ) )  AND  ( TITLE-ABS-KEY ( respon*  OR    manage*  OR  control    OR  interven*  OR    strateg* ) )  AND   ( TITLE-ABS-KEY ( stochastic  OR    transmission  OR  computational  OR    mathematical  OR  mechanistic    OR  statistical  OR    simulation  OR  "In silico"  OR    dynamic* ) )  AND  ( TITLE-ABS-KEY ( model* ) )  AND    ( ( TITLE-ABS-KEY ( cholera    OR  dengue  OR    diphtheria  OR  ebola    OR    "Foot-and-mouth"    OR   "foot and   mouth"  OR  fmd    OR  "Hepatitis A"  OR    "Hepatitis B"  OR  "Hepatitis E"  OR    "Haemophilus influenzae type b"  OR    hib  OR  "Human papillomavirus"  OR    hpv  OR  influenza ) )  OR  (   TITLE-ABS-KEY ( "Japanese encephalitis"  OR    malaria  OR  measles    OR  "Meningococcal meningitis"  OR    mumps  OR  pertussis    OR  "Whooping   cough"  OR  "Pneumococcal disease"  OR    poliomyelitis  OR  polio    OR  rabies  OR    rotavirus  OR  rubella ) )    OR  ( TITLE-ABS-KEY ( tetanus  OR    "Tick-borne encephalitis"    OR  tuberculosis  OR    typhoid  OR  varicella     OR  chickenpox  OR    "Yellow Fever"    OR    "vaccine-preventable" ) ) ) & 5\,253 (5\, 095*) \\
			\multicolumn{3}{l}{*Scopus returned 5\,253 records but we could not retrieve 158 records after several attempts.}\\ \bottomrule
			\hline
		\end{tabular}
	\end{table}
\end{sidewaysfigure}

\begin{sidewaysfigure}
	\begin{table}[H]
		\centering
		\caption{PubMed search strings and results.}
		\begin{tabular}{lp{0.85\textwidth}l}
			\toprule
			\textbf{Search} & \textbf{Query} & \textbf{Records found} \\
			\midrule
			\#1 & Search (Epidemic OR Outbreak OR   Emergency OR Reactive OR Crisis) & 1\,021\,994 \\
			\#2 & Search (respon*  OR    manage*  OR  control    OR  interven*  OR    strateg*) & 9\,134\,363 \\
			\#3 & Search (Stochastic OR Transmission   OR Computational OR Mathematical OR Mechanistic OR Statistical OR Simulation   OR "In silico" OR Dynamic*) & 2\,563\,238 \\
			\#4 & Search (model*) & 2\,222\,906 \\
			\#5 & Search (Cholera OR Dengue OR   Diphtheria OR Ebola OR  "Foot-and-mouth" OR "foot and   mouth" OR FMD OR "Hepatitis A" OR "Hepatitis B" OR   "Hepatitis E" OR "Haemophilus influenzae type b" OR Hib   OR "Human papillomavirus" OR HPV OR Influenza OR "Japanese   encephalitis" OR Malaria OR Measles OR "Meningococcal   meningitis" OR Mumps OR Pertussis OR "Whooping cough" OR   "Pneumococcal disease" OR Poliomyelitis OR Polio OR Rabies OR Rotavirus   OR Rubella OR Tetanus OR "Tick-borne encephalitis" OR Tuberculosis   OR Typhoid OR Varicella OR Chickenpox OR "Yellow Fever" OR   "vaccine-preventable") & 887\,440 \\
			\#6 & Search ((((((Epidemic OR Outbreak   OR Emergency OR Reactive OR Crisis)))  AND ((Response OR Management OR Control   OR Intervention OR Strategies))) AND ((Stochastic OR Transmission OR   Computational OR Mathematical OR Mechanistic OR Statistical OR Simulation OR   "In silico" OR Dynamic*))) AND model*) AND ((Cholera OR Dengue OR   Diphtheria OR Ebola OR  "Foot-and-mouth" OR "foot and   mouth" OR FMD OR "Hepatitis A" OR  "Hepatitis B" OR   "Hepatitis E" OR "Haemophilus influenzae type b" OR Hib  OR "Human papillomavirus" OR HPV OR Influenza OR "Japanese   encephalitis" OR Malaria OR Measles OR "Meningococcal   meningitis" OR Mumps OR Pertussis OR "Whooping cough" OR   "Pneumococcal disease" OR Poliomyelitis OR Polio OR Rabies OR   Rotavirus OR Rubella OR Tetanus OR "Tick-borne encephalitis" OR   Tuberculosis OR Typhoid OR Varicella OR Chickenpox OR "Yellow Fever" OR "vaccine-preventable")) & 3\,088 \\
			\bottomrule
		\end{tabular}
	\end{table}
\end{sidewaysfigure}

\begin{sidewaysfigure}
	\begin{table}[H]
		\centering
		\caption{Web of Science search strings and results}
		\begin{tabular}{lp{0.85\textwidth}l}
			\toprule
			\textbf{Search} & \textbf{Query} & \textbf{Records found} \\
			\midrule
			\# 1 & TOPIC: (Epidemic OR Outbreak OR   Emergency OR Reactive OR Crisis) & 1\,241\,675 \\
			\# 2 & TOPIC: (respon* OR  manage*    OR  control  OR    interven*  OR  strateg*) & 13\,791\,565 \\
			\# 3 & TOPIC: (Stochastic OR Transmission   OR Computational OR Mathematical OR Mechanistic OR Statistical OR Simulat* OR    "In silico" OR Dynamic*) & 7\,869\,289 \\
			\# 4 & TOPIC: (model*) & 705\,715 \\
			\# 5 & TOPIC: (Cholera OR Dengue OR Diphtheria   OR Ebola OR "Foot-and-mouth" OR "foot and mouth" OR FMD   OR "Hepatitis A" OR "Hepatitis B" OR "Hepatitis   E" OR "Haemophilus influenzae type b" OR Hib OR "Human   papillomavirus" OR HPV OR Influenza OR "Japanese encephalitis" OR Malaria OR Measles OR "Meningococcal meningitis" OR Mumps OR  Pertussis OR "Whooping cough" OR "Pneumococcal disease"   OR Poliomyelitis OR Polio OR Rabies OR Rotavirus OR Rubella OR Tetanus OR   "Tick-borne encephalitis" OR Tuberculosis OR Typhoid OR Varicella   OR Chickenpox OR "Yellow Fever" OR "vaccine-preventable") & 8\,790\,133 \\
			\#6 & TOPIC: (Epidemic OR Outbreak OR   Emergency OR Reactive OR Crisis) AND TOPIC: (Respon* OR  Manage* OR  Control  OR    Interven*  OR  Strateg*)) AND TOPIC: (Stochastic OR   Transmission OR Computational OR Mathematical OR Mechanistic OR Statistical   OR Simulation OR In silico OR Dynamic*) AND TOPIC: (model*) AND TOPIC:   (Cholera OR Dengue OR Diphtheria OR Ebola OR  "Foot-and-mouth" OR   "foot and mouth" OR FMD OR "Hepatitis A" OR   "Hepatitis B" OR "Hepatitis E" OR "Haemophilus   influenzae type b" OR Hib OR "Human papillomavirus" OR HPV OR   Influenza OR "Japanese encephalitis" OR Malaria OR Measles OR   "Meningococcal meningitis" OR Mumps OR Pertussis OR "Whooping   cough" OR "Pneumococcal disease" OR Poliomyelitis OR Polio OR   Rabies OR Rotavirus OR Rubella OR Tetanus OR "Tick-borne   encephalitis" OR Tuberculosis OR Typhoid OR  Varicella OR Chickenpox OR   "Yellow Fever" OR "vaccine-preventable") & 4\,803 \\
			\bottomrule
		\end{tabular}
	\end{table}
\end{sidewaysfigure}

%\newpage		
%\section{Collaboration patterns in time}
%\begin{figure}[H]
%		\centering
%		\includegraphics[scale=0.75]{figs/collab_proportions.png}
%		\caption[Relative proportion (purely academic versus mixed collaborations) of studies per year.]{Relative proportion (purely academic versus mixed collaborations) of studies per year. Academic collaborations refer to those authored by only authors with academic institution affiliations whereas mixed collaborations include those authored by a mixture of authors with academic institution affiliations, government institutions, and NGOs. Mixed collaborations are represented by the turquoise bars and purely academic collaborations by the brown bars.} 
%	\end{figure}
\end{document}