### 2023-10-18 analyses/scripts/S4_Table.tex
Generate latex version of files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Supplementary material},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Supplementary material}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

The following table contains information on the data extracted from each
of the included studies.

Technical note: Due to the large number of rows, the table could not be
printed in one go and was instead printed a few columns at a time. In
actual fact, it is just a repetition of the title column along with the
presented columns.

\hypertarget{table-1}{%
\subsection{Table 1}\label{table-1}}

This section presents extracted data on the author affiliation type
(affiliation\_type), whether it is a purely academic collaboration
(academic\_collab), and the location that was studied.

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{4cm}l>{\raggedright\arraybackslash}p{3.5cm}l>{\raggedright\arraybackslash}p{1.5cm}}
\toprule
num & title & year & affiliation\_type & academic\_collab & location\\
\midrule
\endfirsthead
\multicolumn{6}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & year & affiliation\_type & academic\_collab & location\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{ZW}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & 2019 & academic institutions + government institutions & FALSE & DK FR IS IT NL NO GB\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{LY}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & 2019 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{TW}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & 2019 & academic institutions & TRUE & BR\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CN}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & 2019 & academic institutions + government institutions & FALSE & CN\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{MX US}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & 2019 & academic institutions & TRUE & none\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CN}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & 2019 & academic institutions & TRUE & west africa\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & 2019 & academic institutions & TRUE & ET\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{NG}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & 2019 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CN}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & 2019 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{west africa}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & 2019 & academic institutions & TRUE & HT\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GN LR SL}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & 2018 & academic institutions + NGO & FALSE & CN\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{IN}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & 2018 & academic institutions & TRUE & LR\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{IL}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & 2018 & academic institutions & TRUE & NO\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{AU CA NZ GB US}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & 2018 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{BF}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & 2018 & academic institutions & TRUE & none\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{GB}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & 2018 & academic institutions & TRUE & PH\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{BR}\\
34 & modeling the heterogeneity of dengue transmission in a city & 2018 & academic institutions + government institutions + NGO & FALSE & CN\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{AO}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & 2018 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CN}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & 2018 & academic institutions + government institutions & FALSE & BE\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{TD}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & 2018 & academic institutions & TRUE & US\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & 2017 & government institutions & FALSE & none\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{JP}\\
44 & analysis of an seriv epidemic model of human influenza & 2017 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & 2017 & NGO & FALSE & NG\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & 2017 & academic institutions & TRUE & CN\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GB}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & 2017 & academic institutions & TRUE & US\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & 2017 & academic institutions + government institutions & FALSE & CN\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{SE}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & 2017 & academic institutions + government institutions & FALSE & none\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{LR}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & 2017 & academic institutions & TRUE & SD\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CN}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & 2017 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{LR}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & 2017 & academic institutions & TRUE & west africa\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & 2016 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & 2016 & government institutions + NGO & FALSE & global\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & 2016 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HK}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & 2016 & academic institutions & TRUE & HK\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{AT}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & 2016 & academic institutions + government institutions & FALSE & US\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CA}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & 2016 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{NP NE YE ZM}\\
74 & metapopulation and non-proportional vaccination models overview & 2016 & academic institutions & TRUE & MX\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{GB}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & 2016 & academic institutions & TRUE & west africa\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
78 & modeling the impact of behavior change on the spread of ebola & 2016 & academic institutions & TRUE & LY\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{IL}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & 2016 & academic institutions & TRUE & CN\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{west africa}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & 2016 & academic institutions & TRUE & HK\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{west africa}\\
84 & projected impact of dengue vaccination in yucatán, mexico & 2016 & academic institutions + government institutions & FALSE & MX\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & 2016 & academic institutions & TRUE & MX\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{GN}\\
88 & the interplay of vaccination and vector control on small dengue networks & 2016 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & 2015 & academic institutions & TRUE & none\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{global}\\
92 & dengue in madeira island & 2015 & academic institutions & TRUE & PT\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{MX}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & 2015 & academic institutions & TRUE & CA\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{LR}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & 2015 & academic institutions + NGO & FALSE & SL\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CH}\\
98 & measuring the impact of ebola control measures in sierra leone & 2015 & academic institutions & TRUE & SL\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CN}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & 2015 & academic institutions + government institutions & FALSE & IL\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GN LR SL}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & 2015 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{west africa}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & 2015 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & 2015 & academic institutions & TRUE & LR\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HK}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & 2015 & academic institutions + NGO & FALSE & LR\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CA}\\
110 & the effects of media reports on disease spread and important public health measurements & 2015 & academic institutions + NGO & FALSE & CA\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{GN LR SL}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & 2015 & academic institutions & TRUE & west africa\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{west africa}\\
114 & vaccine prioritization for effective pandemic response & 2015 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & 2014 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{global}\\
118 & assessing the use of antiviral treatment to control influenza & 2014 & academic institutions & TRUE & CA\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{ZW}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & 2014 & academic institutions + government institutions & FALSE & none\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & 2014 & academic institutions + government institutions & FALSE & LR\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CN}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & 2014 & academic institutions + government institutions & FALSE & CH\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & 2014 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HT}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & 2014 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & 2014 & academic institutions & TRUE & US\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{NG}\\
132 & a perspective on multiple waves of influenza pandemics & 2013 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & 2013 & academic institutions + government institutions & FALSE & HT\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{DK}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & 2013 & academic institutions & TRUE & TW\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & 2013 & academic institutions & TRUE & DK\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CV}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & 2013 & academic institutions & TRUE & JP\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CA}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & 2013 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & 2013 & academic institutions & TRUE & IN\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{2013} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GB}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & 2013 & academic institutions + NGO & FALSE & CN\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{SG}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & 2012 & academic institutions + NGO & FALSE & CN\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{NO}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & 2012 & academic institutions + government institutions + NGO & FALSE & NO\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & 2012 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CN}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & 2012 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HK}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & 2012 & academic institutions + government institutions & FALSE & none\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & 2012 & academic institutions & TRUE & NL\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GR}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & 2011 & academic institutions & TRUE & MX\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{GB}\\
162 & a multilayer framework to assess influenza intervention policies & 2011 & academic institutions & TRUE & TW\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & 2011 & academic institutions & TRUE & CN\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CA}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & 2011 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{none}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & 2011 & academic institutions + government institutions & FALSE & MX\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HT}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & 2011 & academic institutions & TRUE & none\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & 2011 & academic institutions + government institutions & FALSE & none\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CN}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & 2011 & academic institutions & TRUE & HU\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CA}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & 2011 & academic institutions + government institutions + NGO & FALSE & US\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & 2011 & academic institutions + government institutions & FALSE & SG\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{global}\\
180 & toward effective vaccine deployment: a systematic study & 2011 & academic institutions & TRUE & HK\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CA}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & 2010 & academic institutions & TRUE & ES\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CN}\\
184 & effective real-time allocation of pandemic interventions & 2010 & NGO & FALSE & US\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & 2010 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions + government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{CA}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & 2010 & academic institutions + NGO & FALSE & none\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & 2010 & academic institutions + government institutions + NGO & FALSE & US\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & 2010 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{NL}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & 2010 & academic institutions + government institutions & FALSE & JP\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{MX}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & 2009 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & 2009 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{HK}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & 2009 & academic institutions + government institutions + NGO & FALSE & who southeast asia region\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{SE TZ}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & 2009 & academic institutions + government institutions & FALSE & JP\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{northern hemisphere}\\
204 & modelling foot and mouth disease & 2009 & academic institutions & TRUE & GB\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{MX}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & 2009 & academic institutions + government institutions & FALSE & none\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{global}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & 2009 & academic institutions & TRUE & HK\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{global}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & 2009 & academic institutions & TRUE & US\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{2009} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & 2008 & academic institutions & TRUE & HK\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{2008} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{NL}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & 2008 & academic institutions + government institutions & FALSE & CA\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{2008} & \cellcolor{gray!6}{academic institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & 2008 & academic institutions + government institutions & FALSE & NL\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{2008} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & 2008 & academic institutions + government institutions + NGO & FALSE & JP\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{2008} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{IT}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & 2008 & academic institutions + government institutions + NGO & FALSE & NE\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{global}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & 2007 & academic institutions & TRUE & NL GB US\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CO}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & 2007 & academic institutions & TRUE & NL\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{AU}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & 2007 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & 2007 & academic institutions & TRUE & US\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{BR}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & 2006 & academic institutions + NGO & FALSE & none\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{US}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & 2006 & academic institutions + government institutions & FALSE & UY\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
234 & simple models for containment of a pandemic & 2006 & academic institutions & TRUE & none\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{CD}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & 2005 & academic institutions + government institutions & FALSE & southeast asia\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{2003} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
238 & modelling vaccination strategies against foot-and-mouth disease & 2003 & academic institutions & TRUE & GB\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{2002} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{NL}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & 2002 & academic institutions & TRUE & NL\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{2001} & \cellcolor{gray!6}{academic institutions + government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{GB}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & 2000 & government institutions & FALSE & FR\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{2000} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{FR}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & 1999 & academic institutions + government institutions & FALSE & VE\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{1998} & \cellcolor{gray!6}{academic institutions} & \cellcolor{gray!6}{TRUE} & \cellcolor{gray!6}{BR}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & 1996 & academic institutions & TRUE & DK FO\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{1995} & \cellcolor{gray!6}{government institutions + NGO} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{AU}\\
248 & analysis of a measles epidemic & 1993 & academic institutions + government institutions & FALSE & US\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{1992} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{none}\\
250 & a discrete-time model with vaccination for a measles epidemic & 1991 & academic institutions & TRUE & US\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{1982} & \cellcolor{gray!6}{government institutions} & \cellcolor{gray!6}{FALSE} & \cellcolor{gray!6}{US}\\
252 & stochastic two-agent epidemic simulation models for a community of families & 1971 & academic institutions & TRUE & none\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-2}{%
\subsection{Table 2}\label{table-2}}

This section presents extracted data on the disease studied (disease),
type of objective studied (objectives), whether the paper had an author
in the location studied (author\_in\_country\_studied), and how
individuals were represented (individual\_representation).

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{3.3cm}l>{\raggedright\arraybackslash}p{3.3cm}>{\raggedright\arraybackslash}p{2cm}l}
\toprule
num & title & author\_in\_country\_studied & disease & objectives & individual\_representation\\
\midrule
\endfirsthead
\multicolumn{6}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & author\_in\_country\_studied & disease & objectives & individual\_representation\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & not applicable & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{hepatitis a} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & yes & tuberculosis & prospective analysis & compartments\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{dengue} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & yes & dengue & retrospective analysis & compartments\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & not applicable & influenza & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & no & tuberculosis & prospective analysis & compartments\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{meningococcal meningitis} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & no & cholera & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & no & ebola & retrospective analysis & compartments\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{poliomyelitis} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & yes & measles & retrospective analysis & compartments\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{meningococcal meningitis} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & not applicable & cholera ebola influenza measles & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & yes & tuberculosis & prospective analysis & compartments\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{measles} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
34 & modeling the heterogeneity of dengue transmission in a city & yes & dengue & prospective analysis & compartments\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yellow fever} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & no & ebola & prospective analysis & agents\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & yes & influenza & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & not applicable & influenza measles & prospective analysis & compartments\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
44 & analysis of an seriv epidemic model of human influenza & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & no & poliomyelitis & prospective analysis & compartments\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{individuals representation other}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & yes & dengue & retrospective analysis & compartments\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & yes & influenza & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{varicella} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & no & ebola & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & not applicable & poliomyelitis & prospective analysis & compartments\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{measles} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & yes & influenza & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{pertussis} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & not applicable & tuberculosis & prospective analysis & compartments\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{measles} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
74 & metapopulation and non-proportional vaccination models overview & no & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{pertussis} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
78 & modeling the impact of behavior change on the spread of ebola & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{poliomyelitis} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & yes & dengue & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
84 & projected impact of dengue vaccination in yucatán, mexico & yes & dengue & retrospective analysis & agents\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{dengue influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
88 & the interplay of vaccination and vector control on small dengue networks & not applicable & dengue & prospective analysis & compartments\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & not applicable & influenza & prospective analysis & individuals representation other\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
92 & dengue in madeira island & yes & dengue & prospective analysis & compartments\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
98 & measuring the impact of ebola control measures in sierra leone & no & ebola & retrospective analysis & compartments\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & yes & poliomyelitis & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & no & ebola & retrospective analysis & compartments\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & no & ebola & retrospective analysis & agents\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
110 & the effects of media reports on disease spread and important public health measurements & yes & influenza & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & no & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
114 & vaccine prioritization for effective pandemic response & yes & influenza & prospective analysis & individuals representation other\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{tuberculosis} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
118 & assessing the use of antiviral treatment to control influenza & no & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & not applicable & influenza & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & yes & ebola & prospective analysis & compartments\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza mumps varicella} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & yes & fmd & prospective analysis & agents\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{poliomyelitis} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & not applicable & typhoid & prospective analysis & compartments\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & yes & measles & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{meningococcal meningitis} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
132 & a perspective on multiple waves of influenza pandemics & yes & influenza & retrospective analysis & agents\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{dengue} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & no & cholera & prospective analysis & agents\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & yes & fmd & prospective analysis & agents\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{dengue} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & yes & influenza & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & yes & influenza & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & yes & fmd & prospective analysis & agents\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & yes & influenza & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
162 & a multilayer framework to assess influenza intervention policies & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{cholera} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & not applicable & influenza & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & not applicable & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
180 & toward effective vaccine deployment: a systematic study & yes & influenza & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & yes & fmd & prospective analysis & compartments\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
184 & effective real-time allocation of pandemic interventions & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & yes & influenza & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & yes & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & no & tuberculosis & prospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
204 & modelling foot and mouth disease & no & fmd & prospective analysis & compartments\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & yes & fmd & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & no & pneumococcal disease & retrospective analysis & compartments\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{hepatitis b} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza pneumococcal disease} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & yes & measles & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & yes & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{varicella} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{agents}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & yes & fmd & retrospective analysis & agents\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & not applicable & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & no & influenza & retrospective analysis & compartments\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yellow fever} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & not applicable & influenza & prospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & no & fmd & prospective analysis & agents\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{influenza} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
234 & simple models for containment of a pandemic & not applicable & influenza & prospective analysis & compartments\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{ebola} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & yes & influenza & prospective analysis & agents\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
238 & modelling vaccination strategies against foot-and-mouth disease & yes & fmd & prospective analysis & agents\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & yes & fmd & retrospective analysis & agents\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & yes & fmd & retrospective analysis & compartments\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{retrospective analysis} & \cellcolor{gray!6}{compartments}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & yes & malaria & retrospective analysis & compartments\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{rubella} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & no & varicella & retrospective analysis & compartments\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{fmd} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{agents}\\
248 & analysis of a measles epidemic & yes & measles & prospective analysis & compartments\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{not applicable} & \cellcolor{gray!6}{dengue} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
250 & a discrete-time model with vaccination for a measles epidemic & yes & measles & retrospective analysis & compartments\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{measles} & \cellcolor{gray!6}{prospective analysis} & \cellcolor{gray!6}{compartments}\\
252 & stochastic two-agent epidemic simulation models for a community of families & not applicable & poliomyelitis & prospective analysis & agents\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-3}{%
\subsection{Table 3}\label{table-3}}

This section presents extracted data on whether spatial heterogeneity
was represented (space\_representation), the type of model structure
(model\_structure), the type of parametrization and validation method
used.

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{3cm}l>{\raggedright\arraybackslash}p{3cm}ll}
\toprule
num & title & space\_representation & model\_structure & parametrization & validation\\
\midrule
\endfirsthead
\multicolumn{6}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & space\_representation & model\_structure & parametrization & validation\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & yes & deterministic & expert opinion & none\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & yes & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & no & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & no & deterministic & literature and expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & no & deterministic & literature & none\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & no & stochastic & literature and expert opinion & none\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & no & deterministic & literature and expert opinion & data\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & no & deterministic & literature and fitted & none\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{another model}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & no & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & no & stochastic & expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & no & deterministic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
34 & modeling the heterogeneity of dengue transmission in a city & no & deterministic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & no & stochastic & literature and expert opinion & none\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & yes & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & no & deterministic & expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
44 & analysis of an seriv epidemic model of human influenza & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{another model}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & no & stochastic & literature and expert opinion and fitted & none\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & no & deterministic & fitted & data\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{data}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & no & deterministic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & no & stochastic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & yes & both & literature and expert opinion & none\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & no & deterministic & literature & none\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & yes & deterministic & literature and expert opinion & data\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & yes & deterministic & literature & none\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
74 & metapopulation and non-proportional vaccination models overview & yes & deterministic & literature & none\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{another model}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & yes & stochastic & literature & none\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{data}\\
78 & modeling the impact of behavior change on the spread of ebola & no & deterministic & expert opinion and fitted & none\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & no & deterministic & literature & none\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & no & stochastic & literature and expert opinion and fitted & none\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
84 & projected impact of dengue vaccination in yucatán, mexico & no & deterministic & literature and fitted & none\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & no & stochastic & literature & none\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
88 & the interplay of vaccination and vector control on small dengue networks & yes & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & yes & deterministic & literature & none\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
92 & dengue in madeira island & no & deterministic & literature & none\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{data}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & no & stochastic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{data}\\
98 & measuring the impact of ebola control measures in sierra leone & yes & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & no & deterministic & literature & none\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & yes & stochastic & fitted & data\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & yes & both & literature and fitted & data\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & yes & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
110 & the effects of media reports on disease spread and important public health measurements & no & both & literature and expert opinion and fitted & data\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
114 & vaccine prioritization for effective pandemic response & no & both & literature and expert opinion & data and another model\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
118 & assessing the use of antiviral treatment to control influenza & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{data}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & no & deterministic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & no & stochastic & literature and fitted & data\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & yes & deterministic & fitted & none\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & no & stochastic & expert opinion & none\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & no & deterministic & literature & none\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
132 & a perspective on multiple waves of influenza pandemics & no & deterministic & literature and fitted & none\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & yes & deterministic & fitted & none\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & no & deterministic & expert opinion and fitted & none\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & yes & stochastic & literature and expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & yes & stochastic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & no & deterministic & literature & data\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & no & deterministic & literature and fitted & none\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & no & deterministic & literature and expert opinion and fitted & data\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & yes & deterministic & expert opinion and fitted & data\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & yes & stochastic & literature and fitted & none\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & no & deterministic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
162 & a multilayer framework to assess influenza intervention policies & yes & deterministic & expert opinion & none\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & no & stochastic & literature and expert opinion & none\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & no & deterministic & literature and fitted & none\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & yes & stochastic & fitted & none\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & no & both & literature and expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{none}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & no & deterministic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & no & both & literature and fitted & data\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & yes & stochastic & literature and expert opinion & none\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{data}\\
180 & toward effective vaccine deployment: a systematic study & no & deterministic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & yes & stochastic & literature and expert opinion and fitted & none\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
184 & effective real-time allocation of pandemic interventions & yes & stochastic & literature & none\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{data}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & yes & stochastic & literature and fitted & none\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & no & deterministic & literature & none\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & no & stochastic & literature and expert opinion & none\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & no & deterministic & literature and expert opinion and fitted & data\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & no & stochastic & literature and expert opinion and fitted & none\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
204 & modelling foot and mouth disease & yes & deterministic & literature and expert opinion & data\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion and fitted} & \cellcolor{gray!6}{none}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & no & deterministic & expert opinion & none\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & yes & deterministic & fitted & none\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{data}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & no & stochastic & fitted & none\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & no & deterministic & fitted & none\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & no & deterministic & literature & data\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & yes & stochastic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and expert opinion and fitted} & \cellcolor{gray!6}{none}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & yes & stochastic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & no & deterministic & literature & none\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible infected removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{both} & \cellcolor{gray!6}{literature and fitted} & \cellcolor{gray!6}{data}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & yes & stochastic & literature and expert opinion and fitted & none\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & no & stochastic & expert opinion & none\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{another model}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & yes & deterministic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & yes & stochastic & literature and expert opinion & data\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{another model}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & yes & deterministic & fitted & data\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
234 & simple models for containment of a pandemic & no & deterministic & literature & none\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{data}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & yes & stochastic & literature and expert opinion and fitted & none\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
238 & modelling vaccination strategies against foot-and-mouth disease & yes & stochastic & fitted & data\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & yes & stochastic & expert opinion & none\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{fitted} & \cellcolor{gray!6}{data}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & no & deterministic & literature and expert opinion & none\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & no & deterministic & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & no & deterministic & literature and fitted & data\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{literature} & \cellcolor{gray!6}{none}\\
248 & analysis of a measles epidemic & no & both & literature and expert opinion and fitted & data\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{deterministic} & \cellcolor{gray!6}{literature and expert opinion} & \cellcolor{gray!6}{none}\\
250 & a discrete-time model with vaccination for a measles epidemic & no & deterministic & literature and fitted & data\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{stochastic} & \cellcolor{gray!6}{expert opinion} & \cellcolor{gray!6}{none}\\
252 & stochastic two-agent epidemic simulation models for a community of families & no & stochastic & expert opinion & none\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-4}{%
\subsection{Table 4}\label{table-4}}

This section presents extracted data on the type of outbreak and the
interventions studied.

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{3cm}l>{\raggedright\arraybackslash}p{8cm}}
\toprule
num & title & outbreak\_type & interventions\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & outbreak\_type & interventions\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & hypothetical outbreak & travel restriction,vaccination\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{contact tracing,ppe other 1,quarantine,safe burial,distancing,social mobilisation}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & hypothetical outbreak & education\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & real outbreak & prison: entry screening,exit screening,secondary ipt,passive diagnosis active diagnosis and combined above\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{larvicides}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & real outbreak & isolation,adult vector control,larva control\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{media campaign}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & hypothetical outbreak & face masks\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{face masks,respiratory protective devices}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & real outbreak & treatment,vaccination\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{PPE other 1,vaccination}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{media campaign}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & real outbreak & school closure\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation,quarantine,treatment}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & real outbreak & hygiene,quarantine,treatment,vaccination\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{ebola treatment units}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & real outbreak & close live poultry market\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation,school closure,vaccination,stay home}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & real outbreak & education,isolation,quarantine,safe burial,distancing\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & hypothetical outbreak & sick leave strategy\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination,stomping out}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & real outbreak & vaccination\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & hypothetical outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{rewiring}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & real outbreak & distancing from infectious individuals,latent case finding,case holding,active case finding\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
34 & modeling the heterogeneity of dengue transmission in a city & real outbreak & vaccination,vector control\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & real outbreak & contact tracing,hospitalization\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,isolation,school closure,vaccination,antiviral treatment}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & real outbreak & school closure\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination,antibiotic prophylaxis,point-of-use water treatment}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & hypothetical outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & hypothetical outbreak & unclear\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{culling,vaccination,early detection}\\
44 & analysis of an seriv epidemic model of human influenza & real outbreak & media campaign,vaccination\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{disease control}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & real outbreak & reducing mosquito density\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{behavioural change}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & hypothetical outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & real outbreak & antiviral prophylaxis,isolation,school closure,vaccination,therapeutics\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{culling,vaccination,movement restriction,national standstill,protection zones,surveillance zones}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & hypothetical outbreak & behavioural change,vaccination\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,treatment,vaccination}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & hypothetical outbreak & education\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,quarantine,safe burial,treatment,vaccination}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & real outbreak & media campaign,capturing infected bats\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation,quarantine,safe burial}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & real outbreak & vaccination\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{travel restriction}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & hypothetical outbreak & media campaign\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{school closure}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & real outbreak & vaccination\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{culling,vaccination,ring culling,and combinations of above}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & real outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & hypothetical outbreak & treatment,vaccination\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
74 & metapopulation and non-proportional vaccination models overview & hypothetical outbreak & distancing,school closure,vaccination\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{school closure}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & real outbreak & vaccination\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{contact tracing,vaccination}\\
78 & modeling the impact of behavior change on the spread of ebola & real outbreak & behavioural change\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & hypothetical outbreak & larvicides\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{education,treatment,control the propagation,reduce the number of susceptible and infected and simultaneously increase recovered}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & hypothetical outbreak & face masks,isolation,school closure,vaccination,ventilation\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,safe burial,vaccination,protected sex with condoms,early detection and diagnosis,preventing importation}\\
84 & projected impact of dengue vaccination in yucatán, mexico & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{contact tracing,safe burial,vaccination,ebola treatment units}\\
88 & the interplay of vaccination and vector control on small dengue networks & hypothetical outbreak & vaccination,sterile insect technique\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{screening}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & hypothetical outbreak & antiviral prophylaxis,isolation\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{hygiene,isolation,hospitalization}\\
92 & dengue in madeira island & hypothetical outbreak & education,larvicides\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{quarantine,vaccination}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{ebola treatment units,community care centers}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & hypothetical outbreak & beds,ebola treatment units,community care centers\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{contact tracing,culling,travel restriction,vaccination,surveillance}\\
98 & measuring the impact of ebola control measures in sierra leone & real outbreak & beds\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{media campaign}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & real outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,media campaign,safe burial,vaccination}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & hypothetical outbreak & school closure\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & real outbreak & school closure,vaccination\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{reducing transmission}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & real outbreak & safe burial,effective hospitalization\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{school closure}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & real outbreak & beds,PPE other 1,safe burial\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
110 & the effects of media reports on disease spread and important public health measurements & real outbreak & media campaign,distancing,vaccination\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{behavioural change}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & real outbreak & several interventions to reduce transmission\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{treatment}\\
114 & vaccine prioritization for effective pandemic response & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{treatment}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & hypothetical outbreak & antiviral prophylaxis,isolation\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{travel restriction}\\
118 & assessing the use of antiviral treatment to control influenza & hypothetical outbreak & antiviral prophylaxis,vaccination\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{water treatment}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & hypothetical outbreak & behavioural change\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{behavioural change,vaccination}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & real outbreak & contact tracing,beds,household protective kits\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{school absenteeism surveillance}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & hypothetical outbreak & screening\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & hypothetical outbreak & direct measures to reduce infectiousness,e.g.,quarantine,prophylactic measures to mitigate outbreak seasonality and morbidity\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & hypothetical outbreak & behavioural change,contact tracing,vaccination\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
132 & a perspective on multiple waves of influenza pandemics & real outbreak & travel restriction,vaccination\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{treatment,mosquito repellent}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & real outbreak & vaccination\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination,depopulation of detected herds,protection,surveillance zones,movement tracing,national standstill}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & real outbreak & face masks,handwashing,isolation,vaccination,ventilation,high efficiency particulate air filtration,ultraviolet germicidal irradiation\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,distancing,school closure,travel restriction}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{larvicides,vaccination}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & hypothetical outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,distancing}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & real outbreak & vaccination\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{paid sick days,stay-at-home behaviour}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & real outbreak & school closure\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & real outbreak & quarantine\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{school closure}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & real outbreak & school closure\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{school closure}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & real outbreak & antiviral prophylaxis,vaccination\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation,travel restriction}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & hypothetical outbreak & unclear\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,quarantine}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & real outbreak & vaccination\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{travel restriction,hospitalization,antiviral treatment}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & hypothetical outbreak & school closure\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{school closure}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & hypothetical outbreak & culling,vaccination\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & real outbreak & isolation\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
162 & a multilayer framework to assess influenza intervention policies & real outbreak & treatment,hospitalization\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{distancing,antiviral treatment}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & hypothetical outbreak & school closure\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{face masks,pallative care,treatment}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & hypothetical outbreak & treatment\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{antiviral prophylaxis}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & real outbreak & distancing,school closure\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{hygiene,vaccination}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & hypothetical outbreak & antiviral prophylaxis,treatment\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{distancing,treatment}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & hypothetical outbreak & isolation,school closure,vaccination\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & real outbreak & vaccination\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & real outbreak & vaccination\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & hypothetical outbreak & school closure\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
180 & toward effective vaccine deployment: a systematic study & hypothetical outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & hypothetical outbreak & vaccination,preventive depopulation\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{quarantine,school closure,hygiene precaution}\\
184 & effective real-time allocation of pandemic interventions & hypothetical outbreak & vaccination\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & hypothetical outbreak & school closure,vaccination,antiviral treatment\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{mass immunization program}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & hypothetical outbreak & isolation,antiviral treatment\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & real outbreak & school closure\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{face masks}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & hypothetical outbreak & antiviral prophylaxis\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{culling}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & real outbreak & isolation,onboard quarantine inspection\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & hypothetical outbreak & antiviral prophylaxis\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,distancing,vaccination}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & hypothetical outbreak & antiviral prophylaxis,vaccination\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{behavioural change,travel restriction}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & real outbreak & treatment,vaccination,diagnostics\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & hypothetical outbreak & antiviral prophylaxis,isolation,school closure,vaccination\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,vaccination}\\
204 & modelling foot and mouth disease & hypothetical outbreak & culling,vaccination,movement restriction,surveillance\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{handwashing,sanitary measures}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & real outbreak & treatment,post-exposure prophylaxis\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & hypothetical outbreak & contact tracing,isolation,quarantine\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{treatment}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & hypothetical outbreak & early detection\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{vaccination}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & real outbreak & treatment\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & hypothetical outbreak & antiviral prophylaxis,face masks,hygiene,school closure,vaccination,work closure,antiviral treatment\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,quarantine,distancing,school closure,screening,work closure}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & real outbreak & vaccination\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,school closure,vaccination}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & real outbreak & school closure,vaccination,prohibition of traffic\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,distancing,school closure,vaccination,work closure,closure of public places}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & real outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{travel restriction}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & hypothetical outbreak & antiviral prophylaxis,treatment,vaccination,reduction of hospital transmission,reduction of community transmission\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible infected removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{isolation,vaccination}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & real outbreak & culling,vaccination,surveillance,preventive slaughter of dangerous contacts,movement restriction\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{travel restriction}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & hypothetical outbreak & distancing\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{treatment}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & real outbreak & face masks,hygiene,isolation,quarantine,distancing,school closure,work closure,prevention of mass gatherings,several combinations of interventions\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & hypothetical outbreak & antiviral prophylaxis,school closure,treatment,vaccination,work closure,stay-at-home\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,vaccination}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & real outbreak & vaccination\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{antiviral prophylaxis,distancing,school closure,travel restriction,vaccination}\\
234 & simple models for containment of a pandemic & hypothetical outbreak & antiviral prophylaxis,vaccination\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{interventions to reduce transmission}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & hypothetical outbreak & antiviral prophylaxis,quarantine,distancing,school closure,work closure\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{culling,vaccination}\\
238 & modelling vaccination strategies against foot-and-mouth disease & real outbreak & vaccination\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{culling,vaccination,stomping out,movement control}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & hypothetical outbreak & culling,vaccination\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{real outbreak} & \cellcolor{gray!6}{culling,vaccination}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & real outbreak & vaccination,stamping-out and slaughter of contact herds\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{stamping out}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & real outbreak & treatment,drugs\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & real outbreak & vaccination\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{vaccination,stamping out,dangerous contact slaughter}\\
248 & analysis of a measles epidemic & real outbreak & vaccination\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{ultra-low volume (ulv) insecticide applications}\\
250 & a discrete-time model with vaccination for a measles epidemic & real outbreak & vaccination\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{hypothetical outbreak} & \cellcolor{gray!6}{isolation,quarantine}\\
252 & stochastic two-agent epidemic simulation models for a community of families & hypothetical outbreak & school closure,vaccination\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-5}{%
\subsection{Table 5}\label{table-5}}

This section presents extracted data on whether vaccination was modelled
as a sole intervention for comparison against other interventions
(vax\_modelled\_with\_non\_vax) and the studies' conclusion on the
impact of vaccination (is\_vax\_effective).

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{4cm}l>{\raggedright\arraybackslash}p{4cm}}
\toprule
num & title & vax\_modelled\_with\_non\_vax & is\_vax\_effective\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & vax\_modelled\_with\_non\_vax & is\_vax\_effective\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & no & not applicable\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & no & not applicable\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & no & not applicable\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & no & not applicable\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & no & not applicable\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & no & not applicable\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & no & not applicable\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & no & not applicable\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & yes & yes\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & no & not applicable\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & no & not applicable\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & no & not applicable\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & no & not applicable\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & no & not applicable\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
34 & modeling the heterogeneity of dengue transmission in a city & no & not applicable\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & no & not applicable\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & no & not applicable\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & no & not applicable\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
44 & analysis of an seriv epidemic model of human influenza & no & not applicable\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & no & not applicable\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & no & not applicable\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & yes & yes\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & yes & yes\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & no & not applicable\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & no & not applicable\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & no & not applicable\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & no & not applicable\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & no & not applicable\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & no & not applicable\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{inconclusive}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & no & not applicable\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
74 & metapopulation and non-proportional vaccination models overview & no & not applicable\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & no & not applicable\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
78 & modeling the impact of behavior change on the spread of ebola & no & not applicable\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & yes & no\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
84 & projected impact of dengue vaccination in yucatán, mexico & no & not applicable\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & no & not applicable\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
88 & the interplay of vaccination and vector control on small dengue networks & no & not applicable\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
92 & dengue in madeira island & no & not applicable\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{inconclusive}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & no & not applicable\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & no & not applicable\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
98 & measuring the impact of ebola control measures in sierra leone & no & not applicable\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & no & not applicable\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & no & not applicable\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & no & not applicable\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & no & not applicable\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
110 & the effects of media reports on disease spread and important public health measurements & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & no & not applicable\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
114 & vaccine prioritization for effective pandemic response & no & not applicable\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & no & not applicable\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
118 & assessing the use of antiviral treatment to control influenza & yes & inconclusive\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & no & not applicable\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & no & not applicable\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & no & not applicable\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & no & not applicable\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
132 & a perspective on multiple waves of influenza pandemics & no & not applicable\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & no & not applicable\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{inconclusive}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & yes & no\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & no & not applicable\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & no & not applicable\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & no & not applicable\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & no & not applicable\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & no & not applicable\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & yes & no\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & no & not applicable\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & no & not applicable\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & no & not applicable\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & yes & no\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
162 & a multilayer framework to assess influenza intervention policies & no & not applicable\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & no & not applicable\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & no & not applicable\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & no & not applicable\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & no & not applicable\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & no & not applicable\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & yes & yes\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & no & not applicable\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
180 & toward effective vaccine deployment: a systematic study & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & yes & no\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
184 & effective real-time allocation of pandemic interventions & no & not applicable\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & no & not applicable\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & no & not applicable\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & no & not applicable\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & no & not applicable\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & no & not applicable\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & yes & yes\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & yes & no\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
204 & modelling foot and mouth disease & no & not applicable\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & no & not applicable\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & no & not applicable\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & no & not applicable\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & yes & no\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & no & not applicable\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & yes & yes\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & yes & inconclusive\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & no & not applicable\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & no & not applicable\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & no & not applicable\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & no & not applicable\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
234 & simple models for containment of a pandemic & no & not applicable\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & no & not applicable\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{inconclusive}\\
238 & modelling vaccination strategies against foot-and-mouth disease & no & not applicable\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & yes & inconclusive\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & yes & no\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & no & not applicable\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & no & not applicable\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
248 & analysis of a measles epidemic & no & not applicable\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
250 & a discrete-time model with vaccination for a measles epidemic & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
252 & stochastic two-agent epidemic simulation models for a community of families & no & not applicable\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-6}{%
\subsection{Table 6}\label{table-6}}

This section presents extracted data on the outcomes measured and
whether a sensitivity analysis was conducted.

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{6cm}l}
\toprule
num & title & outcome\_measured & sensitivity\_analysis\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & outcome\_measured & sensitivity\_analysis\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{cases averted,final epidemic size,cost,case fatality} & \cellcolor{gray!6}{no}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & final epidemic size & yes\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{yes}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & final epidemic size,cost & no\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{yes}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & cases averted & yes\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{cases averted,final epidemic size} & \cellcolor{gray!6}{yes}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & attack rate & yes\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{yes}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & final epidemic size,control reproduction number, & yes\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{intervention coverage,outbreak duration and timing,final epidemic size,attack rate} & \cellcolor{gray!6}{yes}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & final epidemic size & no\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{attack rate,prevalence of infection} & \cellcolor{gray!6}{yes}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & incidence & yes\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{cases averted,cost} & \cellcolor{gray!6}{yes}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & outbreak duration and timing,final epidemic size,cost & no\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{final epidemic size,peak magnitude} & \cellcolor{gray!6}{yes}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & cases averted,timing of peak,cumulative attack rate & yes\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{number of individuals in s,number of susceptible individuals,number of latently infected individuals,isolated or quarantined individuals,and the individuals that have recovered and are immune to evd} & \cellcolor{gray!6}{yes}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & final epidemic size,cost & no\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{funerals,cumulative infections} & \cellcolor{gray!6}{yes}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & cases averted,outbreak duration and timing,final epidemic size & no\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{attack rate} & \cellcolor{gray!6}{no}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & cumulative deaths,cumulative cases & yes\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & cost & no\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{outbreak duration and timing,total number of infected premises} & \cellcolor{gray!6}{no}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & outbreak duration and timing,final epidemic size,instantaneous reproduction number,case reproduction number & yes\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{incidence} & \cellcolor{gray!6}{yes}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & intervention coverage & no\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & cost,number of exposed and infectious individuals & yes\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{immunization threshold} & \cellcolor{gray!6}{yes}\\
34 & modeling the heterogeneity of dengue transmission in a city & final epidemic size,timing of peak,peak size & no\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{deaths averted,final epidemic size,attack rate,case fatality} & \cellcolor{gray!6}{yes}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & attack rate & no\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{outbreak duration and timing,attack rate} & \cellcolor{gray!6}{no}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & timing of peak,peak difference,incidence & no\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{cases averted,outbreak duration and timing,incidence} & \cellcolor{gray!6}{yes}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & intervention coverage & no\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & cumulative infected & yes\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,cost} & \cellcolor{gray!6}{no}\\
44 & analysis of an seriv epidemic model of human influenza & timing of peak,peak size,transmission & yes\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & cost,dalys,incremental cost effectiveness ratio,incremental net benefits & yes\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{final epidemic size,cost,total deaths,years of life lost} & \cellcolor{gray!6}{yes}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & outbreak duration and timing,final epidemic size & no\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & cost,attack rate,total deaths,risk of death,net benefits,return on investment & yes\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & outbreak duration and timing,attack rate & yes\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,total number of herds surveilled,total number of herds culled,total number of herds vaccinated} & \cellcolor{gray!6}{yes}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & final epidemic size,timing of peak & no\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{infection rate} & \cellcolor{gray!6}{no}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & final epidemic size,cumulative incidence & yes\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{cases averted,intervention coverage,final epidemic size,peak size} & \cellcolor{gray!6}{no}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & final epidemic size,cost & no\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{effective reproduction number} & \cellcolor{gray!6}{no}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & timing of peak,cumulative cases,cumulative deaths,peak magnitude & yes\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{r0} & \cellcolor{gray!6}{yes}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & hospitalizations,case fatality & no\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & paralytic incidence,average expected number of cases & yes\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{final epidemic size,vaccine doses,peak magnitude} & \cellcolor{gray!6}{no}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & intervention coverage & no\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{cost,attack rate} & \cellcolor{gray!6}{yes}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & final epidemic size,attack rate & no\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{intervention coverage,outbreak duration and timing,final epidemic size} & \cellcolor{gray!6}{yes}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & intervention coverage,final epidemic size & no\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{cases averted,cumulative incidence} & \cellcolor{gray!6}{yes}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & incidence,basic reproduction number & no\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{yes}\\
74 & metapopulation and non-proportional vaccination models overview & outbreak duration and timing,final epidemic size & no\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{timing of peak,attack rate,peak incidence,average number of weeks lost} & \cellcolor{gray!6}{yes}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & outbreak duration and timing,final epidemic size & no\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,number of contacts traced} & \cellcolor{gray!6}{yes}\\
78 & modeling the impact of behavior change on the spread of ebola & outbreak duration and timing,final epidemic size,timing of peak & yes\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{outbreak duration and timing,cumulative incidence} & \cellcolor{gray!6}{yes}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & final epidemic size,attack rate & yes\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{s,i and r curve} & \cellcolor{gray!6}{no}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & timing of peak,peak infection rate & yes\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{final epidemic size,total deaths} & \cellcolor{gray!6}{yes}\\
84 & projected impact of dengue vaccination in yucatán, mexico & intervention coverage,effectiveness of vaccination strategies & no\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & outbreak duration and timing & no\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{cumulative cases,incidence} & \cellcolor{gray!6}{yes}\\
88 & the interplay of vaccination and vector control on small dengue networks & geometric mean number of infected hosts,basic reproduction number & yes\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{basic reproduction number} & \cellcolor{gray!6}{yes}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & household reproduction number & no\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{cumulative incidence,cumulative deaths} & \cellcolor{gray!6}{yes}\\
92 & dengue in madeira island & final epidemic size & no\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{cases averted,final epidemic size} & \cellcolor{gray!6}{no}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & attack rate & no\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{cases averted,cumulative incidence} & \cellcolor{gray!6}{no}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & final epidemic size & no\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{outbreak duration and timing,number of animals culled,number of premises culled} & \cellcolor{gray!6}{yes}\\
98 & measuring the impact of ebola control measures in sierra leone & cases averted & yes\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{timing of peak,hospital notifications} & \cellcolor{gray!6}{yes}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & intervention coverage,final epidemic size & no\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{final epidemic size,basic reproduction number} & \cellcolor{gray!6}{yes}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & timing of peak & yes\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{attack rate} & \cellcolor{gray!6}{no}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & attack rate & yes\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{cost,total fraction of infected and exposed} & \cellcolor{gray!6}{no}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & outbreak duration and timing,final epidemic size & yes\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,timing of peak,attack rate} & \cellcolor{gray!6}{yes}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & cases averted,final epidemic size,number of deaths & yes\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{attack rate,hospitalizations} & \cellcolor{gray!6}{no}\\
110 & the effects of media reports on disease spread and important public health measurements & outbreak duration and timing,final epidemic size,timing of peak,peak magnitude,total vaccinated & yes\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{r0} & \cellcolor{gray!6}{yes}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & final epidemic size & yes\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{case fatality} & \cellcolor{gray!6}{no}\\
114 & vaccine prioritization for effective pandemic response & attack rate,mortality rate & yes\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{control reproduction number} & \cellcolor{gray!6}{no}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & control reproduction number & yes\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{risk of case importation} & \cellcolor{gray!6}{yes}\\
118 & assessing the use of antiviral treatment to control influenza & intervention coverage,attack rate & no\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{incidence} & \cellcolor{gray!6}{yes}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & r0 & no\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{final epidemic size,timing of peak,number of simulations with epidemic outbreak,peak infections} & \cellcolor{gray!6}{no}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & cases averted,deaths averted & no\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{attack rate,extra protective rate} & \cellcolor{gray!6}{no}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & outbreak duration and timing & no\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{incidence,vaccine-derived virus prevalence,population immunity} & \cellcolor{gray!6}{no}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & final epidemic size & no\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & campaign duration,final epidemic size,cost,peak magnitude,number of times countermeasures are started,proportion of time that infected size is above a threshold number of infecteds & no\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & intervention coverage,final epidemic size & yes\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{cases averted,outbreak duration and timing} & \cellcolor{gray!6}{yes}\\
132 & a perspective on multiple waves of influenza pandemics & outbreak duration and timing,final epidemic size,attack rate & no\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{yes}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & intervention coverage,attack rate & yes\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{outbreak duration and timing,cost,number of infected herds,number of depopulated herds,export losses} & \cellcolor{gray!6}{yes}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & cost,basic reproduction number & no\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{cost} & \cellcolor{gray!6}{yes}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & outbreak duration and timing,final epidemic size,cost & yes\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{campaign duration,final epidemic size,timing of peak} & \cellcolor{gray!6}{no}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & attack rate & no\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{cases averted,intervention coverage,attack rate} & \cellcolor{gray!6}{no}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & cases averted & no\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{final epidemic size,attack rate} & \cellcolor{gray!6}{yes}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & reproduction number & no\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{intervention coverage} & \cellcolor{gray!6}{yes}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & outbreak duration and timing,cost,timing of peak,cumulative incidence & no\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{attack rate,peak incidence,peak day} & \cellcolor{gray!6}{no}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & hospitalizations & yes\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{attack rate,qualys} & \cellcolor{gray!6}{yes}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & final epidemic size & yes\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{first arrival time} & \cellcolor{gray!6}{no}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & cumulative attack rate & no\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{attack rate} & \cellcolor{gray!6}{no}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & cases averted & no\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{timing of peak,attack rate} & \cellcolor{gray!6}{yes}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & cost,qualys,incremental cost effectiveness ratio (icer),cumulative attack rate & yes\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{final epidemic size,timing of peak} & \cellcolor{gray!6}{no}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & outbreak duration and timing,final epidemic size & no\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{yes}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & timing of peak & no\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{campaign duration,hospitalizations,case fatality} & \cellcolor{gray!6}{yes}\\
162 & a multilayer framework to assess influenza intervention policies & timing of peak,peak magnitude & no\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{final epidemic size,cost,incidence} & \cellcolor{gray!6}{yes}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & timing of peak,attack rate,peak daily infection & no\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{yes}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & final epidemic size,timing of peak,peak size & no\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{final epidemic size,attack rate} & \cellcolor{gray!6}{no}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & cases averted,attack rate & no\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,timing of peak,cumulative hospital cases,date of first reported cases} & \cellcolor{gray!6}{no}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & attack rate & no\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{final epidemic size,cost} & \cellcolor{gray!6}{no}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & intervention coverage,outbreak duration and timing,attack rate,case fatality & no\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & final epidemic size,attack rate,total deaths & yes\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{final epidemic size,hospitalizations,case fatality} & \cellcolor{gray!6}{no}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & cases averted,final epidemic size,timing of peak,hospitalizations & no\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{deaths averted,attack rate,hospitalizations} & \cellcolor{gray!6}{yes}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & timing of peak,attack rate,peak incidence & no\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{attack rate,prevalence} & \cellcolor{gray!6}{yes}\\
180 & toward effective vaccine deployment: a systematic study & cumulative infections & no\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{campaign duration,attack rate,case fatality} & \cellcolor{gray!6}{yes}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & outbreak duration and timing,number of infected premises,number of depopulated premises,number of quarantined premises & yes\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{timing of peak} & \cellcolor{gray!6}{no}\\
184 & effective real-time allocation of pandemic interventions & morbidity,indirect protection & yes\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{yes}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & final epidemic size,timing of peak,cumulative attack rate,peak magnitude & no\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{cost,hospitalizations,qualys} & \cellcolor{gray!6}{no}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & attack rate,cumulative incidence & yes\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{attack rate,hospitalizations,case fatality} & \cellcolor{gray!6}{yes}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & attack rate,incidence & yes\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{attack rate,r0} & \cellcolor{gray!6}{no}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & final epidemic size,attack rate & no\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{effective reproduction number} & \cellcolor{gray!6}{yes}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & time of detection,maximum number of symptomatic cases per day & no\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{timing of peak,hospitalizations,deaths} & \cellcolor{gray!6}{no}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & attack rate & no\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{outbreak duration and timing,cost} & \cellcolor{gray!6}{yes}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & cases averted,deaths averted,cost,qualys,incremental cost effectiveness ratio & yes\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{final epidemic size} & \cellcolor{gray!6}{no}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & incidence,cumulative cases,cumulative deaths & yes\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{household reproduction number for} & \cellcolor{gray!6}{yes}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & final epidemic size & no\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{final epidemic size,attack rate,peak size} & \cellcolor{gray!6}{yes}\\
204 & modelling foot and mouth disease & number of infected farms,basic reproduction number & no\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{attack rate} & \cellcolor{gray!6}{no}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & final epidemic size,total deaths & yes\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{final epidemic size,attack rate,mortality rate} & \cellcolor{gray!6}{no}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & attack rate & no\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{final epidemic size,timing of peak,incidence} & \cellcolor{gray!6}{yes}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & lives saved,outbreak duration and timing & no\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{attack rate,r0} & \cellcolor{gray!6}{yes}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & attack rate & no\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & final epidemic size,hospitalizations,cumulative deaths & no\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{cases averted,attack rate} & \cellcolor{gray!6}{yes}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & cases averted,hospitalizations & yes\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{final epidemic size,timing of peak,attack rate,case fatality} & \cellcolor{gray!6}{no}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & final epidemic size,timing of peak,peak magnitude & no\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{outbreak duration and timing,timing of peak,cumulative attack rate} & \cellcolor{gray!6}{yes}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & cases averted & no\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{timing of peak,peak magnitude} & \cellcolor{gray!6}{no}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & final epidemic size,hospitalizations,deaths & no\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{cases averted} & \cellcolor{gray!6}{no}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & outbreak duration and timing,number of infected farms,the number of pre-emptively slaughtered farms,number of farms in control zones & yes\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{delay between epidemics} & \cellcolor{gray!6}{yes}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & final epidemic size,timing of peak,peak magnitude & no\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{final epidemic size,attack rate,resistant cases} & \cellcolor{gray!6}{no}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & final epidemic size,mortality & no\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{final epidemic size,r0,vaccination coverage} & \cellcolor{gray!6}{no}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & outbreak duration and timing,final epidemic size,attack rate,cumulative incidence & yes\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{attack rate} & \cellcolor{gray!6}{yes}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & final epidemic size & no\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{attack rate,cumulative incidence} & \cellcolor{gray!6}{yes}\\
234 & simple models for containment of a pandemic & final epidemic size & no\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size} & \cellcolor{gray!6}{no}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & probability of preventing a large outbreak,number of courses of drug required to achieve containment & yes\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{outbreak duration and timing,cost,number of herds slaughtered,number of herds vaccinated} & \cellcolor{gray!6}{no}\\
238 & modelling vaccination strategies against foot-and-mouth disease & outbreak duration and timing,final epidemic size & yes\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{final epidemic size,cost,export losses} & \cellcolor{gray!6}{no}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & outbreak duration and timing,final epidemic size,attack rate & no\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,total number of animals culled} & \cellcolor{gray!6}{no}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & outbreak duration and timing,final epidemic size & yes\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{cost} & \cellcolor{gray!6}{no}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & duration of infection & yes\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{force of infection,proportion of susceptible individuals} & \cellcolor{gray!6}{yes}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & mortality rate,incidence & no\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{outbreak duration and timing,final epidemic size,cost,compensation,control costs,output effects,income effects,employment effects} & \cellcolor{gray!6}{no}\\
248 & analysis of a measles epidemic & cumulative cases & no\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{cases averted,timing of peak,peak magnitude,population immunity} & \cellcolor{gray!6}{yes}\\
250 & a discrete-time model with vaccination for a measles epidemic & cumulative incidence & no\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{outbreak duration and timing} & \cellcolor{gray!6}{no}\\
252 & stochastic two-agent epidemic simulation models for a community of families & intervention coverage,outbreak duration and timing,timing of peak & no\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-7}{%
\subsection{Table 7}\label{table-7}}

This section presents extracted data on whether observed data was used
for model validation (data\_used), and whether the data was made
available (data\_available).

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{9cm}ll}
\toprule
num & title & data\_used & data\_available\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & data\_used & data\_available\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & no & not applicable\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & no & not applicable\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & yes & yes\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & yes & no\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & no & not applicable\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & yes & no\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & no & not applicable\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & yes & yes\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & yes & no\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & no & not applicable\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & yes & yes\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & no & not applicable\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & yes & no\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & yes & yes\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
34 & modeling the heterogeneity of dengue transmission in a city & yes & no\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & no & not applicable\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & yes & no\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & no & not applicable\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
44 & analysis of an seriv epidemic model of human influenza & no & not applicable\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & yes & no\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & yes & no\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & yes & yes\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & yes & yes\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & no & not applicable\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & yes & yes\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & no & not applicable\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & yes & yes\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & yes & yes\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & no & not applicable\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & no & not applicable\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & yes & yes\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & no & not applicable\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
74 & metapopulation and non-proportional vaccination models overview & yes & yes\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & no & not applicable\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
78 & modeling the impact of behavior change on the spread of ebola & yes & yes\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & yes & yes\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & yes & yes\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
84 & projected impact of dengue vaccination in yucatán, mexico & no & not applicable\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & no & not applicable\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
88 & the interplay of vaccination and vector control on small dengue networks & no & not applicable\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
92 & dengue in madeira island & no & not applicable\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & yes & yes\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & yes & no\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
98 & measuring the impact of ebola control measures in sierra leone & yes & yes\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & yes & yes\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & yes & yes\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & yes & yes\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & yes & yes\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
110 & the effects of media reports on disease spread and important public health measurements & yes & no\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & yes & yes\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
114 & vaccine prioritization for effective pandemic response & yes & yes\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & no & not applicable\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
118 & assessing the use of antiviral treatment to control influenza & no & not applicable\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & yes & yes\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & no & not applicable\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & no & not applicable\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & no & not applicable\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
132 & a perspective on multiple waves of influenza pandemics & yes & yes\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & yes & yes\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & yes & yes\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & no & not applicable\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & yes & yes\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & yes & yes\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & yes & no\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & yes & no\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & yes & yes\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & yes & no\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & no & not applicable\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & yes & yes\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & yes & yes\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & yes & no\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & yes & no\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
162 & a multilayer framework to assess influenza intervention policies & no & not applicable\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & yes & yes\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & no & not applicable\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & yes & yes\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & no & not applicable\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & yes & yes\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & yes & yes\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & yes & no\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
180 & toward effective vaccine deployment: a systematic study & yes & yes\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & yes & yes\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
184 & effective real-time allocation of pandemic interventions & no & not applicable\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & no & not applicable\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & no & not applicable\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & yes & yes\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & no & not applicable\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & no & not applicable\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & no & not applicable\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & yes & yes\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & yes & yes\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & yes & yes\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
204 & modelling foot and mouth disease & yes & yes\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & no & not applicable\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & no & not applicable\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & yes & yes\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & no & not applicable\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & no & not applicable\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & yes & no\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & yes & no\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & no & not applicable\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & yes & no\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & no & not applicable\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & yes & yes\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & yes & no\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & yes & no\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
234 & simple models for containment of a pandemic & no & not applicable\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & yes & no\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
238 & modelling vaccination strategies against foot-and-mouth disease & yes & no\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & no & not applicable\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & no & not applicable\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & yes & no\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & yes & yes\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
248 & analysis of a measles epidemic & yes & no\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
250 & a discrete-time model with vaccination for a measles epidemic & yes & yes\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{not applicable}\\
252 & stochastic two-agent epidemic simulation models for a community of families & no & not applicable\\*
\end{longtable}
\endgroup{}
\end{landscape}

\hypertarget{table-8}{%
\subsection{Table 8}\label{table-8}}

This section presents extracted data on whether the simulation code was
made available (simulation\_code\_available) and whether vaccination was
modelled (vax\_modelled).

\begin{landscape}\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}{l>{\raggedright\arraybackslash}p{4cm}l>{\raggedright\arraybackslash}p{4cm}}
\toprule
num & title & simulation\_code\_available & vax\_modelled\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
num & title & simulation\_code\_available & vax\_modelled\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{a cholera metapopulation model interlinking migration with intervention strategies - a case study of zimbabwe (2008-2009)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
2 & a theoretical single-parameter model for urbanisation to study infectious disease spread and interventions & no & yes\\
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{assessing the effects of modeling the spectrum of clinical symptoms on the dynamics and control of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
4 & dynamical analysis and control strategies in modelling ebola virus disease & no & no\\
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{effect of a hepatitis a vaccination campaign during a hepatitis a outbreak in taiwan, 2015–2017: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
6 & evaluating strategies for control of tuberculosis in prisons and prevention of spillover into communities: an observational and modeling study from brazil & no & no\\
\cellcolor{gray!6}{7} & \cellcolor{gray!6}{evaluating the effects of control interventions and estimating the inapparent infections for dengue outbreak in hangzhou, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
8 & incidence dynamics and investigation of key interventions in a dengue outbreak in ningbo city, china & no & no\\
\cellcolor{gray!6}{9} & \cellcolor{gray!6}{incorporating media data into a model of infectious disease transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
10 & influence of asymptomatic infections for the effectiveness of facemasks during pandemic influenza & no & no\\
\addlinespace
\cellcolor{gray!6}{11} & \cellcolor{gray!6}{mathematical analysis of a cholera infection model with vaccination strategy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
12 & modeling and stability analysis of epidemic expansion disease ebola virus with implications prevention in population & no & yes\\
\cellcolor{gray!6}{13} & \cellcolor{gray!6}{modeling the effectiveness of respiratory protective devices in reducing influenza outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
14 & modeling the effects of vaccination and treatment on tuberculosis transmission dynamics & no & yes\\
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{optimal control and cost-effective analysis of the 2017 meningitis outbreak in nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
16 & optimal control strategies for a two-group epidemic model with vaccination-resource constraints & no & yes\\
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{optimal media reporting intensity on mitigating spread of an emerging infectious disease} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
18 & school dismissal as a pandemic influenza response: when, where and for how long? & no & no\\
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{study on modeling simulation and optimal control method for the transmission risk of the ebola virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
20 & transmission dynamics of cholera with hyperinfectious and hypoinfectious vibrios: mathematical modelling and control strategies & no & yes\\
\addlinespace
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{a new epidemics–logistics model: insights into controlling the ebola virus disease in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
22 & contact reductions from live poultry market closures limit the epidemic of human infections with h7n9 influenza & no & no\\
\cellcolor{gray!6}{23} & \cellcolor{gray!6}{disparities in spread and control of influenza in slums of delhi: findings from an agent-based modelling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
24 & efficacy of control measures in the control of ebola, liberia 2014–2015 & no & no\\
\cellcolor{gray!6}{25} & \cellcolor{gray!6}{epidemiology of the silent polio outbreak in rahat, israel, based on modeling of environmental surveillance data} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
26 & evaluating costs and health consequences of sick leave strategies against pandemic and seasonal influenza in norway using a dynamic model & no & no\\
\cellcolor{gray!6}{27} & \cellcolor{gray!6}{evaluating vaccination strategies to control foot-and-mouth disease: a country comparison study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
28 & impact of public health responses during a measles outbreak in an amish community in ohio: modeling the dynamics of transmission & no & yes\\
\cellcolor{gray!6}{29} & \cellcolor{gray!6}{initial validation of a simulation model for estimating the impact of serogroup a neisseria meningitidis vaccination in the african meningitis belt} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
30 & logistical constraints lead to an intermediate optimum in outbreak response vaccination & no & yes\\
\addlinespace
\cellcolor{gray!6}{31} & \cellcolor{gray!6}{manipulation of contact network structure and the impact on foot-and-mouth disease transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
32 & mathematical model and intervention strategies for mitigating tuberculosis in the philippines & no & no\\
\cellcolor{gray!6}{33} & \cellcolor{gray!6}{measles outbreak response decision-making under uncertainty: a retrospective analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
34 & modeling the heterogeneity of dengue transmission in a city & no & yes\\
\cellcolor{gray!6}{35} & \cellcolor{gray!6}{modelling the large-scale yellow fever outbreak in luanda, angola, and the impact of vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
36 & quantifying the impact of early-stage contact tracing on controlling ebola diffusion & no & no\\
\cellcolor{gray!6}{37} & \cellcolor{gray!6}{simulation of key interventions for seasonal influenza outbreak control at school in changsha, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
38 & the impact of regular school closure on seasonal influenza epidemics: a data-driven spatial transmission model for belgium & no & no\\
\cellcolor{gray!6}{39} & \cellcolor{gray!6}{the potential impact of case-area targeted interventions in response to cholera outbreaks: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
40 & value of inventory information in allocating a limited supply of influenza vaccine during a pandemic & no & yes\\
\addlinespace
\cellcolor{gray!6}{41} & \cellcolor{gray!6}{an age-structured model for cholera control with vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
42 & an approach to and web-based tool for infectious disease outbreak intervention analysis & no & no\\
\cellcolor{gray!6}{43} & \cellcolor{gray!6}{an economic assessment of foot and mouth disease in japan} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
44 & analysis of an seriv epidemic model of human influenza & no & yes\\
\cellcolor{gray!6}{45} & \cellcolor{gray!6}{control strategies of avian influenza pandemic model with time delay} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
46 & costs and benefits of including inactivated in addition to oral poliovirus vaccine in outbreak response after cessation of oral poliovirus vaccine use & no & yes\\
\cellcolor{gray!6}{47} & \cellcolor{gray!6}{deriving effective vaccine allocation strategies for pandemic influenza: comparison of an agent-based simulation and a compartmental model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
48 & early rigorous control interventions can largely reduce dengue outbreak magnitude: experience from chaozhou, china & no & no\\
\cellcolor{gray!6}{49} & \cellcolor{gray!6}{effects of reactive social distancing on the 1918 influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
50 & epidemiological and economic impact of pandemic influenza in chicago: priorities for vaccine interventions & no & yes\\
\addlinespace
\cellcolor{gray!6}{51} & \cellcolor{gray!6}{estimating direct and indirect protective effect of influenza vaccination in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
52 & evaluating the effects of common control measures for influenza a (h1n1) outbreak at school in china: a modeling study & no & yes\\
\cellcolor{gray!6}{53} & \cellcolor{gray!6}{evaluation of strategies to control a potential outbreak of foot-and-mouth disease in sweden} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
54 & impact of a waning vaccine and altered behavior on the spread of influenza & no & yes\\
\cellcolor{gray!6}{55} & \cellcolor{gray!6}{mathematical models for devising the optimal ebola virus disease eradication} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
56 & modeling the role of public health education in ebola virus disease outbreaks in sudan & no & no\\
\cellcolor{gray!6}{57} & \cellcolor{gray!6}{modelling the transmission and control strategies of varicella among school children in shenzhen, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
58 & optimal and adaptive control of an epidemic model of influenza with unknown parameters & no & yes\\
\cellcolor{gray!6}{59} & \cellcolor{gray!6}{simulating endogenous dynamics of intervention-capacity deployment: ebola outbreak in liberia} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
60 & simulating the potential role of media coverage and infected bats in the 2014 ebola outbreak & no & no\\
\addlinespace
\cellcolor{gray!6}{61} & \cellcolor{gray!6}{a mathematical model of ebola virus disease: using sensitivity analysis to determine effective intervention targets} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
62 & a multi-agent model for adaptive vaccination during infectious disease outbreaks & no & yes\\
\cellcolor{gray!6}{63} & \cellcolor{gray!6}{assessing the efficiency of movement restriction as a control strategy of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
64 & characterization of outbreak response strategies and potential vaccine stockpile needs for the polio endgame & no & yes\\
\cellcolor{gray!6}{65} & \cellcolor{gray!6}{designing public health policies to mitigate the adverse consequences of rural-urban migration via meta-population modeling} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
66 & dynamic model of influenza with age-structured and media coverage & no & no\\
\cellcolor{gray!6}{67} & \cellcolor{gray!6}{economic evaluation of individual school closure strategies: the hong kong 2009 h1n1 pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
68 & efficient vaccine distribution based on a hybrid compartmental model & no & yes\\
\cellcolor{gray!6}{69} & \cellcolor{gray!6}{epidemiological evaluation of different fmd control strategies in two selected regions in austria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
70 & evaluating targeted interventions via meta-population models with multi-level mixing & no & yes\\
\addlinespace
\cellcolor{gray!6}{71} & \cellcolor{gray!6}{evaluation of outbreak response immunization in the control of pertussis using agent-based modeling} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
72 & global stability and optimal control for a tuberculosis model with vaccination and treatment & no & yes\\
\cellcolor{gray!6}{73} & \cellcolor{gray!6}{impact on epidemic measles of vaccination campaigns triggered by disease outbreaks or serosurveys: a modeling study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
74 & metapopulation and non-proportional vaccination models overview & no & yes\\
\cellcolor{gray!6}{75} & \cellcolor{gray!6}{model-based comprehensive analysis of school closure policies for mitigating influenza epidemics and pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
76 & modeling ring-vaccination strategies to control ebola virus disease epidemics & no & yes\\
\cellcolor{gray!6}{77} & \cellcolor{gray!6}{modeling the effect of public health resources and alerting on the dynamics of pertussis spread*} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
78 & modeling the impact of behavior change on the spread of ebola & no & no\\
\cellcolor{gray!6}{79} & \cellcolor{gray!6}{modeling the spread of polio in an ipv-vaccinated population: lessons learned from the 2013 silent outbreak in southern israel} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
80 & modelling weekly vector control against dengue in the guangdong province of china & no & no\\
\addlinespace
\cellcolor{gray!6}{81} & \cellcolor{gray!6}{optimal control strategies for the spread of ebola in west africa} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
82 & potential impact of a ventilation intervention for influenza in the context of a dense indoor contact network in hong kong & no & yes\\
\cellcolor{gray!6}{83} & \cellcolor{gray!6}{predicting and evaluating the epidemic trend of ebola virus disease in the 2014-2015 outbreak and the effects of intervention measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
84 & projected impact of dengue vaccination in yucatán, mexico & no & yes\\
\cellcolor{gray!6}{85} & \cellcolor{gray!6}{seasonality and the effectiveness of mass vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
86 & simulation optimization for an age-based vaccines distribution strategy against the spread of influenza epidemic & no & yes\\
\cellcolor{gray!6}{87} & \cellcolor{gray!6}{spatiotemporal dynamics of the ebola epidemic in guinea and implications for vaccination and disease elimination: a computational modeling analysis} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
88 & the interplay of vaccination and vector control on small dengue networks & no & yes\\
\cellcolor{gray!6}{89} & \cellcolor{gray!6}{an epidemic patchy model with entry–exit screening} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
90 & assessing the role of voluntary self-isolation in the control of pandemic influenza using a household epidemic model & no & no\\
\addlinespace
\cellcolor{gray!6}{91} & \cellcolor{gray!6}{be-codis: a mathematical model to predict the risk of human diseases spread between countries—validation and application to the 2014–2015 ebola virus disease epidemic} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{no}\\
92 & dengue in madeira island & no & no\\
\cellcolor{gray!6}{93} & \cellcolor{gray!6}{dynamic simulation of a seiqr-v epidemic model based on cellular automata} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
94 & early vaccine availability represents an important public health advance for the control of pandemic influenza & no & yes\\
\cellcolor{gray!6}{95} & \cellcolor{gray!6}{effectiveness of ebola treatment units and community care centers — liberia, september 23–october 31, 2014} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
96 & evaluation of the benefts and risks of introducing ebola community care centers, sierra leone & no & no\\
\cellcolor{gray!6}{97} & \cellcolor{gray!6}{evidence for emergency vaccination having played a crucial role to control the 1965/66 foot-and-mouth disease outbreak in switzerland} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
98 & measuring the impact of ebola control measures in sierra leone & no & no\\
\cellcolor{gray!6}{99} & \cellcolor{gray!6}{media impact switching surface during an infectious disease outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
100 & modeling options to manage type 1 wild poliovirus imported into israel in 2013 & no & yes\\
\addlinespace
\cellcolor{gray!6}{101} & \cellcolor{gray!6}{modeling the effect of comprehensive interventions on ebola virus transmission} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
102 & modeling the effect of school closures in a pandemic scenario: exploring two different contact matrices & yes & no\\
\cellcolor{gray!6}{103} & \cellcolor{gray!6}{modelling the effect of early detection of ebola} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
104 & multi-scale modeling for the transmission of influenza and the evaluation of interventions toward it & no & yes\\
\cellcolor{gray!6}{105} & \cellcolor{gray!6}{optimal intervention strategies for a seir control model of ebola epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
106 & predicting the extinction of ebola spreading in liberia due to mitigation strategies & no & no\\
\cellcolor{gray!6}{107} & \cellcolor{gray!6}{school closure strategies for the 2009 hong kong hini influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
108 & spatiotemporal spread of the 2014 outbreak of ebola virus disease in liberia and the effectiveness of non-pharmaceutical interventions: a computational modelling analysis & no & no\\
\cellcolor{gray!6}{109} & \cellcolor{gray!6}{strategies for early vaccination during novel influenza outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
110 & the effects of media reports on disease spread and important public health measurements & no & yes\\
\addlinespace
\cellcolor{gray!6}{111} & \cellcolor{gray!6}{the impact of human behavioral changes in 2014 west africa ebola outbreak} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
112 & transmission dynamics and final epidemic size of ebola virus disease outbreaks with varying interventions & no & no\\
\cellcolor{gray!6}{113} & \cellcolor{gray!6}{treatment–donation-stockpile dynamics in ebola convalescent blood transfusion therapy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
114 & vaccine prioritization for effective pandemic response & no & yes\\
\cellcolor{gray!6}{115} & \cellcolor{gray!6}{a mathematical study of a tb model with treatment interruptions and two latent periods} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
116 & antiviral prophylaxis and isolation for the control of pandemic influenza & no & no\\
\cellcolor{gray!6}{117} & \cellcolor{gray!6}{assessing the impact of travel restrictions on international spread of the 2014 west african ebola epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
118 & assessing the use of antiviral treatment to control influenza & no & yes\\
\cellcolor{gray!6}{119} & \cellcolor{gray!6}{behavior in the time of cholera: evidence from the 2008-2009 cholera outbreak in zimbabwe} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
120 & behavioural response in educated young adults towards influenza a(h1n1)pdm09 & no & no\\
\addlinespace
\cellcolor{gray!6}{121} & \cellcolor{gray!6}{cover your cough! quantifying the benefits of a localized healthy behavior intervention on flu epidemics in washington dc} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
122 & dynamics and control of ebola virus transmission in montserrado, liberia: a mathematical modelling analysis & no & no\\
\cellcolor{gray!6}{123} & \cellcolor{gray!6}{estimating the effectiveness of early control measures through school absenteeism surveillance in observed outbreaks at rural schools in hubei, china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
124 & evaluation of the benefit of emergency vaccination in a foot-and-mouth disease free country with low livestock density & no & yes\\
\cellcolor{gray!6}{125} & \cellcolor{gray!6}{modeling the dynamics of oral poliovirus vaccine cessation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
126 & modeling the impact of optimal screening on typhoid dynamics & no & no\\
\cellcolor{gray!6}{127} & \cellcolor{gray!6}{modeling the influence of environment and intervention oncholera in haiti} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
128 & sequential bayesian inference in hidden markov stochastic kinetic models with application to detection and response to seasonal epidemics & no & no\\
\cellcolor{gray!6}{129} & \cellcolor{gray!6}{simulating optimal vaccination times during cholera outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
130 & the role of vaccination coverage, individual behaviors, and the public health response in the control of measles epidemics: an agent-based simulation for california & no & yes\\
\addlinespace
\cellcolor{gray!6}{131} & \cellcolor{gray!6}{time is (still) of the essence: quantifying the impact of emergency meningitis vaccination response in katsina state, nigeria} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
132 & a perspective on multiple waves of influenza pandemics & no & yes\\
\cellcolor{gray!6}{133} & \cellcolor{gray!6}{an optimal control problem arising from a dengue disease transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
134 & cholera in haiti: reproductive numbers and vaccination coverage estimates & no & yes\\
\cellcolor{gray!6}{135} & \cellcolor{gray!6}{comparing control strategies against foot-and-mouth disease: will vaccination be cost-effective in denmark?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
136 & cost-effectiveness of influenza control measures: a dynamic transmission model-based analysis & no & yes\\
\cellcolor{gray!6}{137} & \cellcolor{gray!6}{cost-effectiveness of workplace closure and travel restriction for mitigating influenza outbreaks: a network-based simulation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
138 & decisions on control of foot-and-mouth disease informed using model predictions & no & yes\\
\cellcolor{gray!6}{139} & \cellcolor{gray!6}{dengue in cape verde: vector control and vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
140 & enhancement of collective immunity in tokyo metropolitan area by selective vaccination against an emerging influenza pandemic & no & yes\\
\addlinespace
\cellcolor{gray!6}{141} & \cellcolor{gray!6}{estimated impact of aggressive empirical antiviral treatment in containing an outbreak of pandemic influenza h1n1 in an isolated first nations community} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
142 & multiregional dynamic vaccine allocation during an influenza epidemic & no & yes\\
\cellcolor{gray!6}{143} & \cellcolor{gray!6}{policies to reduce influenza in the workplace: impact assessments using an agent-based model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
144 & transmission dynamics of the 2009 influenza a (h1n1) pandemic in india: the impact of holiday-related school closure & no & no\\
\cellcolor{gray!6}{145} & \cellcolor{gray!6}{vaccination against foot-and-mouth disease: do initial conditions affect its benefit?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
146 & was mandatory quarantine necessary in china for controlling the 2009 h1n1 pandemic? & no & no\\
\cellcolor{gray!6}{147} & \cellcolor{gray!6}{a contact-network-based simulation model for evaluating interventions under 'what-if' scenarios in epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
148 & campus quarantine (fengxiao) for curbing emergent infectious diseases: lessons from mitigating a/h1n1 in xi'an, china & no & no\\
\cellcolor{gray!6}{149} & \cellcolor{gray!6}{dynamic modelling of costs and health consequences of school closure during an influenza pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
150 & effect of vaccines and antivirals during the major 2009 a(h1n1) pandemic wave in norway - and the influence of vaccination timing & no & yes\\
\addlinespace
\cellcolor{gray!6}{151} & \cellcolor{gray!6}{estimating the value of containment strategies in delaying the arrival time of an influenza pandemic: a case study of travel restriction and patient isolation} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
152 & how to minimize the attack rate during multiple influenza outbreaks in a heterogeneous population & no & no\\
\cellcolor{gray!6}{153} & \cellcolor{gray!6}{modeling strategies for controlling h1n1 outbreaks in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
154 & modeling the effects of h1n1 influenza vaccine distribution in the united states & no & yes\\
\cellcolor{gray!6}{155} & \cellcolor{gray!6}{modeling the impact of air, sea, and land travel restrictions supplemented by other interventions on the emergence of a new influenza pandemic virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
156 & simulating school closure policies for cost effective pandemic decision making & no & no\\
\cellcolor{gray!6}{157} & \cellcolor{gray!6}{the impact of school closures on pandemic influenza: assessing potential repercussions using a seasonal sir model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
158 & vaccination against foot-and-mouth disease i: epidemiological consequences & no & yes\\
\cellcolor{gray!6}{159} & \cellcolor{gray!6}{a deterministic resource scheduling model in epidemic control: a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
160 & a model for the a(h1n1) epidemic in mexico, including social isolation & no & no\\
\addlinespace
\cellcolor{gray!6}{161} & \cellcolor{gray!6}{a model to evaluate mass vaccination against pneumococcus as a countermeasure against pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
162 & a multilayer framework to assess influenza intervention policies & no & no\\
\cellcolor{gray!6}{163} & \cellcolor{gray!6}{a note on the use of optimal control on a discrete time model of influenza dynamics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
164 & a simulation-based assessment of school closure strategies during an influenza epidemic & no & no\\
\cellcolor{gray!6}{165} & \cellcolor{gray!6}{agent-based modeling of the spread of influenza-like illness in an emergency department: a simulation study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
166 & antiviral treatment for pandemic influenza: assessing potential repercussions using a seasonally forced sir model & no & no\\
\cellcolor{gray!6}{167} & \cellcolor{gray!6}{can antiviral drugs contain pandemic influenza transmission?} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
168 & characterizing the epidemiology of the 2009 influenza a/h1n1 pandemic in mexico & no & no\\
\cellcolor{gray!6}{169} & \cellcolor{gray!6}{cholera epidemic in haiti, 2010: using a transmission model to explain spatial spread of disease and identify optimal control interventions} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
170 & controlling an influenza outbreak in deterministic and stochastic models with the emergence and evolution of drug resistance (model 1) & no & no\\
\addlinespace
\cellcolor{gray!6}{171} & \cellcolor{gray!6}{modeling control strategies for concurrent epidemics of seasonal and pandemic h1n1 influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
172 & modeling the impact of behavior changes on the spread of pandemic influenza & no & yes\\
\cellcolor{gray!6}{173} & \cellcolor{gray!6}{modelling and analysis of influenza a (h1n1) on networks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
174 & modelling the strategies for age specific vaccination scheduling during influenza pandemic outbreaks & no & yes\\
\cellcolor{gray!6}{175} & \cellcolor{gray!6}{modelling the transmission dynamics and control of the novel 2009 swine influenza (h1n1) pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
176 & planning for the control of pandemic influenza a (h1n1) in los angeles county and the united states & yes & yes\\
\cellcolor{gray!6}{177} & \cellcolor{gray!6}{prioritization of delayed vaccination for pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
178 & temporal factors in school closure policy for mitigating the spread of influenza & no & no\\
\cellcolor{gray!6}{179} & \cellcolor{gray!6}{the global transmission and control of influenza} & \cellcolor{gray!6}{yes} & \cellcolor{gray!6}{yes}\\
180 & toward effective vaccine deployment: a systematic study & no & yes\\
\addlinespace
\cellcolor{gray!6}{181} & \cellcolor{gray!6}{vaccination against 2009 pandemic h1n1 in a population dynamical model of vancouver, canada: timing is everything} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
182 & a simulation model for the potential spread of foot-and-mouth disease in the castile and leon region of spain & no & yes\\
\cellcolor{gray!6}{183} & \cellcolor{gray!6}{community-based measures for mitigating the 2009 h1n1 pandemic in china} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
184 & effective real-time allocation of pandemic interventions & no & yes\\
\cellcolor{gray!6}{185} & \cellcolor{gray!6}{efficient mitigation strategies for epidemics in rural regions} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
186 & influenza pandemic waves under various mitigation strategies with 2009 h1n1 as a case study & no & yes\\
\cellcolor{gray!6}{187} & \cellcolor{gray!6}{is a mass immunization program for pandemic (h1n1) 2009 good value for money? early evidence from the canadian experience} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
188 & optimal control for pandemic influenza: the role of limited antiviral treatment and isolation & no & no\\
\cellcolor{gray!6}{189} & \cellcolor{gray!6}{optimizing vaccine allocation at different points in time during an epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
190 & simulating school closure strategies to mitigate an influenza epidemic & no & no\\
\addlinespace
\cellcolor{gray!6}{191} & \cellcolor{gray!6}{the effect of mask use on the spread of influenza during a pandemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
192 & transmission dynamics of an influenza model with age of infection and antiviral treatment & no & no\\
\cellcolor{gray!6}{193} & \cellcolor{gray!6}{transmission risks and control of foot-and-mouth disease in the netherlands: spatial patterns} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
194 & when should we intervene to control the 2009 influenza a(h1n1) pandemic? & no & no\\
\cellcolor{gray!6}{195} & \cellcolor{gray!6}{adaptive vaccination strategies to mitigate pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
196 & antiviral resistance and the control of pandemic influenza: the roles of stochasticity, evolution and model details & no & no\\
\cellcolor{gray!6}{197} & \cellcolor{gray!6}{economic evaluation of influenza pandemic mitigation strategies in the united states using a stochastic microsimulation transmission model} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
198 & effectiveness and cost-effectiveness of expanded antiviral prophylaxis and adjuvanted vaccination strategies for an influenza a (h5n1) pandemic & no & yes\\
\cellcolor{gray!6}{199} & \cellcolor{gray!6}{engineering responses to pandemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
200 & epidemiological benefits of more-effective tuberculosis vaccines, drugs, and diagnostics & no & yes\\
\addlinespace
\cellcolor{gray!6}{201} & \cellcolor{gray!6}{household epidemics: modelling effects of early stage vaccination} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
202 & measures against transmission of pandemic h1n1 influenza in japan in 2009: simulation model & no & yes\\
\cellcolor{gray!6}{203} & \cellcolor{gray!6}{modeling vaccination campaigns and the fall/winter 2009 activity of the new a(h1n1) influenza in the northern hemisphere} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
204 & modelling foot and mouth disease & no & yes\\
\cellcolor{gray!6}{205} & \cellcolor{gray!6}{modelling of the influenza a(h1n1)v outbreak in mexico city, april-may 2009, with control sanitary measures} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
206 & post-exposure prophylaxis during pandemic outbreaks & no & no\\
\cellcolor{gray!6}{207} & \cellcolor{gray!6}{potential for a global dynamic of influenza a (h1n1)} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
208 & reducing the impact of the next influenza pandemic using householdbased public health interventions & no & no\\
\cellcolor{gray!6}{209} & \cellcolor{gray!6}{seasonal transmission potential and activity peaks of the new influenza a(h1n1): a monte carlo likelihood analysis based on human mobility} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
210 & simulation of foot-and-mouth disease spread within an integrated livestock system in texas, usa & no & no\\
\addlinespace
\cellcolor{gray!6}{211} & \cellcolor{gray!6}{the transmissibility and control of pandemic influenza a (h1n1) virus} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
212 & bayesian analysis of severe acute respiratory syndrome: the 2003 hong kong epidemic & no & no\\
\cellcolor{gray!6}{213} & \cellcolor{gray!6}{hepatitis b vaccination and changes in sexual risk behaviour among men who have sex with men in amsterdam} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
214 & mathematical assessment of canada's pandemic influenza preparedness plan & no & yes\\
\cellcolor{gray!6}{215} & \cellcolor{gray!6}{modeling targeted layered containment of an influenza pandemic in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
216 & optimal allocation of pandemic influenza vaccine depends on age, risk and timing & no & yes\\
\cellcolor{gray!6}{217} & \cellcolor{gray!6}{pandemic simulation of antivirals plus school closures: buying time until strain-specific vaccine is available} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
218 & preparedness for the spread of influenza: prohibition of traffic, school closure, and vaccination of children in the commuter towns of tokyo & no & yes\\
\cellcolor{gray!6}{219} & \cellcolor{gray!6}{scenarios of diffusion and control of an influenza pandemic in italy} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
220 & time is of the essence: exploring a measles outbreak response vaccination in niamey, niger & no & yes\\
\addlinespace
\cellcolor{gray!6}{221} & \cellcolor{gray!6}{a stochastic equation-based model of the value of international air-travel restrictions for controlling pandemic flu} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
222 & assessing the role of basic control measures, antivirals and vaccine in curtailing pandemic influenza: scenarios for the us, uk and the netherlands & no & yes\\
\cellcolor{gray!6}{223} & \cellcolor{gray!6}{deterministic sir (susceptible–infected–removed) models applied to varicella outbreaks} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
224 & effectiveness of movement-prevention regulations to reduce the spread of foot-and-mouth disease in the netherlands & no & yes\\
\cellcolor{gray!6}{225} & \cellcolor{gray!6}{effects of internal border control on spread of pandemic influenza} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
226 & simple models of influenza progression within a heterogeneous population & no & no\\
\cellcolor{gray!6}{227} & \cellcolor{gray!6}{stochastic model of an influenza epidemic with drug resistance} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
228 & the effect of public health measures on the 1918 influenza pandemic in u.s. cities & no & no\\
\cellcolor{gray!6}{229} & \cellcolor{gray!6}{vaccinating in disease-free regions: a vaccine model with application to yellow fever} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
230 & a 'small-world-like' model for comparing interventions aimed at preventing and controlling influenza pandemics & no & yes\\
\addlinespace
\cellcolor{gray!6}{231} & \cellcolor{gray!6}{containing pandemic influenza with antiviral agents: analytical commentary} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
232 & critical response to post-outbreak vaccination against foot-and-mouth disease & no & yes\\
\cellcolor{gray!6}{233} & \cellcolor{gray!6}{mitigation strategies for pandemic influenza in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
234 & simple models for containment of a pandemic & no & yes\\
\cellcolor{gray!6}{235} & \cellcolor{gray!6}{statistical inference in a stochastic epidemic seir model with control intervention: ebola as a case study} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
\addlinespace
236 & strategies for containing an emerging influenza pandemic in southeast asia & no & no\\
\cellcolor{gray!6}{237} & \cellcolor{gray!6}{modeling alternative mitigation strategies for a hypothetical outbreak of foot-and-mouth disease in the united states} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
238 & modelling vaccination strategies against foot-and-mouth disease & no & yes\\
\cellcolor{gray!6}{239} & \cellcolor{gray!6}{a decision-tree to optimise control measures during the early stage of a foot-and-mouth disease epidemic} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
240 & effect of control measures on the course of simulated foot and mouth disease epidemics that started on different farm types in various dutch areas & no & yes\\
\addlinespace
\cellcolor{gray!6}{241} & \cellcolor{gray!6}{dynamics of the 2001 uk foot and mouth epidemic: stochastic dispersal in a heterogeneous landscape} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
242 & an extended state-transition model for foot-and-mouth disease epidemics in france & no & yes\\
\cellcolor{gray!6}{243} & \cellcolor{gray!6}{simulated economic consequences of foot-and-mouth disease epidemics and their public control in france} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
244 & effect of chemotherapy on malaria transmission among yanomami amerindians: simulated consequences of placebo treatment & no & no\\
\cellcolor{gray!6}{245} & \cellcolor{gray!6}{modelling vaccination strategy against directly transmitted diseases using a series of pulses} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
\addlinespace
246 & mass vaccination to control chickenpox: the influence of zoster & no & yes\\
\cellcolor{gray!6}{247} & \cellcolor{gray!6}{an evaluation of alternate control strategies for foot-and-mouth disease in australia: a regional approach} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{yes}\\
248 & analysis of a measles epidemic & no & yes\\
\cellcolor{gray!6}{249} & \cellcolor{gray!6}{a model of the transmission of dengue fever with an evaluation of the impact of ultra-low volume (ulv) insecticide applications on dengue epidemics} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
250 & a discrete-time model with vaccination for a measles epidemic & no & yes\\
\addlinespace
\cellcolor{gray!6}{251} & \cellcolor{gray!6}{a stochastic model of epidemics in military recruits} & \cellcolor{gray!6}{no} & \cellcolor{gray!6}{no}\\
252 & stochastic two-agent epidemic simulation models for a community of families & no & yes\\*
\end{longtable}
\endgroup{}
\end{landscape}

\end{document}